---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | Using AI for Social Research
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/chatgpt.pdf")
```
\vspace{0.2cm}
\begin{center}
\textbf{\url{https://www.youtube.com/watch?v=oc6RV5c1yd0}}
\end{center}

# How ChatGPT and Bing can help (Hendriksen 2023)

\begin{itemize}
	\item Identifying and adjusting research ideas/questions and relevant literatures (e.g., what has been done in this area?)
	\vspace{0.6cm}
	\item Locating relevant co-authors and key researchers in the field (e.g., who has worked on this topic?)
	\vspace{0.6cm}
	\item Exposing ideas to simulated and targeted critical readers (e.g., how will my research be received)
	\vspace{0.6cm}
	\item Producing research applications (e.g., develop an outline for research proposal; coding and programming for specified tasks)
\end{itemize}

# Research ethics and integrity 

\begin{itemize}
	\item Plagiarism
	\vspace{0.6cm}
	\item Transparency
	\vspace{0.6cm}
	\item Replicability
	\vspace{0.6cm}
	\item Bias
	\vspace{0.6cm}
	\item .... what else?
\end{itemize}

# AI contributions to research content (CUP 2023)

\begin{itemize}
	\item AI use must be \textbf{declared and clearly explained} in publications such as research papers, just as we expect scholars to do with other software, tools and methodologies.
	\vspace{0.6cm}
	\item \textbf{AI does not meet the Cambridge requirements for authorship}, given the need for accountability. AI and LLM tools may not be listed as an author on any scholarly work published by Cambridge.
	\vspace{0.6cm}
	\item \textbf{Authors are accountable for the accuracy, integrity and originality} of their research papers, including for any use of AI.
	\vspace{0.6cm}
	\item Any use of AI must not breach Cambridge's \textbf{plagiarism} policy. Scholarly works must be the author's own, and not present others' ideas, data, words or other material without adequate citation and transparent referencing.
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/survey.pdf")
```
\vspace{0.2cm}
\begin{center}
\textbf{\url{https://dx.doi.org/10.2139/ssrn.4310154}}
\end{center}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/ball.pdf")
```
\vspace{0.2cm}
\begin{center}
\textbf{\url{https://www.pnas.org/doi/10.1073/pnas.2314021121}}
\end{center}

# Exercise: Using AI for/in your research

\begin{itemize}
	\item What computational skills would you like to learn?
	\vspace{0.4cm}
	\item List the methods you will or plan to use in your dissertation project.
	\vspace{0.4cm}
	\item Without using ChatGPT and Bing, imagine how you can integrate AI into your research practice and workflows.
	\vspace{0.4cm}
	\item Now ask ChatGPT and Bing to see what they can do for your research. Are these suggestions at the risk of violating the College's policy on AI and academic integrity?
	\vspace{0.4cm}
	\item You can also ask ChatGPT and Bing in general what they can do for qualitative and/or quantitative (social) researchers.
	\vspace{0.4cm}
	\item How comfortable are you with using AI for research purposes? Why?
\end{itemize}
