---
title: |
  |
  |
  | CP$^2$ Week 10: Using Numbers to Study China
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Plan for today 

\begin{itemize}
  \item Bubble tea meet-up
  \vspace{0.2cm}
  \item Assessment
  \vspace{0.2cm}
  \item Reports on the "Two Sessions:" Asia Society v Chatham (v MERICS)
  \vspace{0.2cm}
  \item Lecture: Using numbers to study China
  \vspace{0.1cm}
  \begin{itemize}
    \item Why numbers matter: A paradigm shift?
    \item Where data comes from
    \item How to work with data
    \item Examples from past and current research
    \item Concluding remarks
  \end{itemize}
\end{itemize}

# Assessment reminders

\begin{itemize}
  \item Optional formative assessment (March 31) -- email me directly
  \vspace{0.1cm}
  \begin{itemize}
    \item For essay: A reference list (no more than 20 refs)
    \item For literature review: A topic statement (100 words) and a reference list (no more than 20 refs)
  \end{itemize}
  \vspace{0.3cm}
  \item Learning journal (April 7)
  \vspace{0.3cm}
  \item Essay or literature review (April 23)
\end{itemize}

# Why numbers matter: A paradigm shift?

\begin{itemize}
\small
  \item The field of China studies has changed significantly as a result of \textbf{methodological/technological advances}, (new) \textbf{data availability} and \textbf{geopolitics}
  \vspace{0.2cm}
  \begin{itemize}
  \vspace{0.1cm}
    \item Quant techniques are essential to \textbf{establish patterns} for theory building and policy impact evaluation
    \item Using numbers also encourage us to rethink the purpose of China studies -- area studies or comparative politics?
    \item The use of computational and quantitative data/methods by no means undermine the importance of traditional approaches -- \textbf{multi-method/data triangulation} is now the norm
  \end{itemize}
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="35%", fig.align="center"}
knitr::include_graphics("Figs/book_crop.pdf")
```
\vspace{0.2cm}
\begin{center}
\scriptsize
\textbf{Manion, Melanie (Duke), Kenneth Lieberthal (Michigan), Mary E. Gallagher (Michigan), and Allen Carlson (Cornell). 2010. \textit{Contemporary Chinese Politics: New Sources, Methods, and Field Strategies}. New York, NY: Cambridge University Press.}
\end{center}

# Where data comes from

\begin{itemize}
\small
  \item Surveys, such as
  \begin{itemize}
    \item Chinese General Social Survey (CGSS)
    \item China Family Panel Studies (CFPS) -- Prof Yu XIE (Princeton)
    \item Asianbarometer or World Values Survey (with caveats)
  \end{itemize}
  \vspace{0.2cm}
  \item Admin data, including socioeconomic surveys and census (e.g., statistical yearbooks 统计年鉴)
  \vspace{0.2cm}
  \item Archives and documents, such as
  \begin{itemize}
    \item Gazetteers (通志)
    \item Yearbooks (年鉴)
    \item Government work reports (政府工作报告)
    \item Culture and history materials (文史资料)
  \end{itemize}
  \vspace{0.2cm}
  \item Digital and "big" data (e.g., Weibo and Baidu Index -- something like Google Trends)
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/truex_cover.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/truex_table.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/truex_res.pdf")
```

---

```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/Examples/hou_cover.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/lu_cover.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/lu_fig1.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/lu_fig2.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/lu_res.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/molly_cover.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/molly_res.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/molly_counter.pdf")
```

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/Examples/yan_cover.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/yan_table.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/Examples/yan_res.pdf")
```

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pkuverse_crop.pdf")
```
\vspace{0.3cm}
\begin{center}
\textbf{\url{https://opendata.pku.edu.cn/}}
\end{center}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pkulaw_crop.pdf")
```
\vspace{0.3cm}
\begin{center}
\textbf{\url{https://www.pkulaw.net/}}
\end{center}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/cdo_crop.pdf")
```
\vspace{0.3cm}
\begin{center}
\textbf{\url{https://www.china-data-online.com/}}
\end{center}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/baidu_crop.pdf")
```
\vspace{0.3cm}
\begin{center}
\textbf{\url{https://index.baidu.com/}}
\end{center}

# Work with data

\begin{itemize}
  \item Training for different \textbf{research methods}
  \vspace{1cm}
  \item A grasp of potential \textbf{data sources}
  \vspace{1cm}
  \item Solid and deep \textbf{field experiences} and/or \textbf{subject knowledge}
\end{itemize}

---

\begin{itemize}
  \item Training for different \textbf{research methods}
  \vspace{0.1cm}
  \begin{itemize}
    \item Qualitative: Case study (Bayesian and machine learning), ethnography and (elite) interviews
    \item Quantitative: Econometrics and regression (frequentist and Bayesian), experiments and causal inference
    \item Computational: Text-as-data, social network analysis, geographic data science and ChatGPT-assisted research
  \end{itemize}
  \vspace{0.6cm}
  \item A grasp of potential \textbf{data sources}
  \vspace{0.6cm}
  \item Solid and deep \textbf{field experiences} and/or \textbf{subject knowledge}
\end{itemize}

---

\begin{itemize}
  \item Training for different \textbf{research methods}
  \vspace{0.6cm}
  \item A grasp of potential \textbf{data sources}
  \vspace{0.1cm}
  \begin{itemize}
    \item Databases
    \item Libraries and archives
    \item Survey firms and research centres
    \item API and web scraping (i.e., coding and programming)
    \item Replication packs
  \end{itemize}
  \vspace{0.6cm}
  \item Solid and deep \textbf{field experiences} and/or \textbf{subject knowledge}
\end{itemize}

---

\begin{itemize}
  \item Training for different \textbf{research methods}
  \vspace{1cm}
  \item A grasp of potential \textbf{data sources}
  \vspace{1cm}
  \item Solid and deep \textbf{field experiences} (e.g., politics of data generation and release) and/or \textbf{subject knowledge} (e.g., political selection and government finances)
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="35%", fig.align="center"}
knitr::include_graphics("Figs/book_crop.pdf")
```
\vspace{0.2cm}
\begin{center}
\scriptsize
\textbf{Manion, Melanie (Duke), Kenneth Lieberthal (Michigan), Mary E. Gallagher (Michigan), and Allen Carlson (Cornell). 2010. \textit{Contemporary Chinese Politics: New Sources, Methods, and Field Strategies}. New York, NY: Cambridge University Press.}
\end{center}

# Data generation process as a research question

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/books.pdf")
```

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/liu_cover.pdf")
```

# Past and current research

\begin{itemize}
  \item Governance and development in multiethnic China
  \vspace{0.1cm}
  \begin{itemize}
    \item Rethinking the politics of ethnic local autonomy (book project)
    \item Poverty alleviation and local state-building in peripheral provinces (\textit{JJPS} 2021)
    \item Ethnic empowerment and environmental governance
    \item Dualilty of ethnic descriptive representation
  \end{itemize}
  \vspace{0.2cm}
  \item (Re)mapping the information network of political elites using computational and AI approaches
  \vspace{0.2cm}
  \item Using court records to study administrative litigation, with \textbf{Haibo He} (Tsinghua) and \textbf{Chao Ma} (UIBE)
  \vspace{0.2cm}
  \item Multilevel agenda setting in times of crisis: A text-as-data approach, with \textbf{Tao Lin} (University of Washington)
\end{itemize}

# Concluding remarks

\begin{itemize}
  \item Find out \textbf{who you are} and \textbf{what you want to be and do} -- your decision may depend on where you are (or going back to history)
  \vspace{0.6cm}
  \item Curate the skills you need -- be \textbf{selectively thorough}
  \vspace{0.6cm}
  \item Find those who can support and collaborate with you -- engage in \textbf{interdisciplinary knowledge co-production}
  \vspace{0.6cm}
  \item Read broadly -- \textbf{go beyond political science and (mainland) China} to gain inspiration and a broader comparative perspective
\end{itemize}

---

\vspace{0.7cm}
Thank you!\
\vspace{0.7cm}
\texttt{c.cheng@bbk.ac.uk}

# Rethinking the politics of ethnic local autonomy

\begin{itemize}
  \item Why do the Communist Party of China grant \textbf{ethnic local autonomy}?
  \vspace{0.5cm}
  \item When and where will the Center designate \textbf{ethnic autonomous territories} (EATs)?
  \vspace{0.5cm}
  \item How can local (policy/administrative) decentralization contribute to nation-state building?
\end{itemize}

# Rethinking the politics of ethnic local autonomy

\begincols
\begincol{.43\textwidth}
  \begin{small}
  \begin{itemize}
  \item \textbf{Archival research} in the US, Hong Kong and China
  \vspace{0.35cm}
  \item \textbf{Elite interviews}: Conversations with Han and non-Han cadres at the central and local levels across different provinces
  \vspace{0.35cm}
  \item \textbf{Process-tracing and QCA}: Structured process tracing the designation of ethnic autonomous prefectures
  \end{itemize}
  \end{small}
  \endcol
  \begincol{.5\textwidth}
  \vspace{0.5cm}
  \begin{figure}
  \centering
  \includegraphics[scale=0.43]{Figs/sites_6}
  \end{figure}
  \endcol
\endcols

# Rethinking the politics of ethnic local autonomy

\begin{itemize}
  \item \textbf{Supervised machine learning} and \textbf{social network analysis}: Capturing the \textit{connectedness} and \textit{network structure} among/between central and local cadres based on an original biographical dataset of Chinese political elites (1949-2003)
  \begin{itemize}
    \item Between the pair of cadres, I create the indices of \textbf{connectedness} based on their \textbf{biographical similarity}
    \begin{figure}
    \centering
    \includegraphics[scale=0.52]{Figs/cpen}
    \end{figure}
    \item The same connectedness measure can be used to construct the information network of Politburo members and measure the degree of \textbf{central leadership fragmentation} by \textbf{community detection}
  \end{itemize}
  \vspace{0.3cm}
  \item \textbf{Multivariate regression}: Modeling the number of active EATs as a function of different network structural indicators
\end{itemize}

# Rethinking the politics of ethnic local autonomy

\begin{itemize}
  \item Ethnic local autonomy acts as an \textbf{institution of agent control}
\end{itemize}
\vspace{0.2cm}
\begin{figure}
  \center
  \includegraphics[scale=0.5]{Figs/findings}
\end{figure}

# Rethinking the politics of ethnic local autonomy

\begin{itemize}
  \item EATs act as \textbf{enclaves} that constrain the power of provincial elites
  \vspace{0.3cm}
  \begin{itemize}
    \item Sub-provincial EATs are \textbf{more} likely to pass and revise autonomous regulations when the provincial elites are potentially defiant
    \vspace{0.2cm}
    \item Ruling non-Han cadres in sub-provincial EATs are \textbf{less} likely to have worked at the provincial level prior to their appointments
  \end{itemize}
\end{itemize}

# Rethinking the politics of ethnic local autonomy

\begin{itemize}
  \item Going \textbf{cross-national}: Post-WWII authoritarian regimes are more likely to introduce regional autonomy to ethnic minorities when their ruling power is contested within the inner circle
  \vspace{0.6cm}
  \item Going back to \textbf{history}: Post-civil war \textit{decentralization} and \textit{state power consolidation} in imperial and Republican China
  \vspace{0.2cm}
  \begin{itemize}
    \item Western Han dynasty (202 BC–220 AD): Kingdoms and marquis states
    \item The Kuomintang (KMT) regime (1928-1949 AD): Countering recalcitrant provincial warlords through the county-level self-government movement
  \end{itemize}
\end{itemize}

# (Re)mapping the information network of political elites

\begin{itemize}
  \item The focus on "high-politics" in the literature on Communist regimes: The absence of independent, strong formal institutions has pushed researchers have to focus on prominent political elites in authoritarian regimes
  \vspace{0.3cm}
  \item \textbf{Factionalism} is one of the most researched topics in the Chinese politics literature (e.g., Nathan 1973; Tsou 1976; Pye 1981; Dittmer 2002; Huang 2006; Bo 2007; Shih et al 2012)
  \vspace{0.3cm}
  \item The power struggles and cooperation between different factions or "informal" groups have been considered as the main driving forces for policy and political changes in the country
\end{itemize}

# (Re)mapping the information network of political elites

\begin{itemize}
  \item Create the timeline of key life events of each cadre using public biographic information
  \vspace{0.1cm}
  \begin{itemize}
    \item Demographic characteristics (e.g., birth year and provincial origin)
    \item Educational background (e.g., colleges and universities)
    \item Political career (e.g., involvement in key events between 1921 and 1949; Party and government posts after 1949)
  \end{itemize}
  \vspace{0.2cm}
  \item Create the (decomposed) \textbf{\textit{attribute vector}} of a cadre based on her/her timeline and measure the distance/similarity between a pair of attribute vectors in the multidimensional space
  \vspace{0.1cm}
  \begin{itemize}
    \item Time: Are the cadres of interest both alive in a given year?
    \item Location: Are they both at the same location in a given year?
    \item Activity: Are they working in the same department/unit in a given year?
  \end{itemize}
\end{itemize}

# (Re)mapping the information network of political elites

\begin{figure}
\centering
\includegraphics[scale=0.26]{Figs/selection}
\end{figure}
\vspace{0.3cm}
\begin{itemize}
  \item Use \textbf{variable selection} (elastic net regularization and other methods) to identify the key biographical variables in CC of each year when measuring biographical similarity
  \vspace{0.1cm}
  \item Use \textbf{community detection} to assess the degree to which a Politburo network can be divided
\end{itemize}

# Using court records to study admin litigation

The release of \textbf{China Judgements Online} (\url{https://wenshu.court.gov.cn/}) made a crucial breakthrough in the empirical legal studies in China (Ma, Yu and He 2016; Liebman et al 2020).
\vspace{0.2cm}
```{r echo=FALSE, out.width="65%", fig.align="center"}
knitr::include_graphics("Figs/wenshu_crop.pdf")
```

# Using court records to study admin litigation

\begin{itemize}
  \item \textbf{Question}: Are the local courts more likely stand with the citizens following the reforms?
  \vspace{0.2cm}
  \begin{itemize}
    \item Railway transport courts (\textit{China Review} 2022)
    \item Case registration system
    \item Elevated and off-site trials
  \end{itemize}
  \vspace{0.6cm}
  \item \textbf{Extensions}
  \vspace{0.2cm}
  \begin{itemize}
    \item Who do the citizens go to the courts? A national \textbf{conjoint} experiment
    \item Local dynamics of judicial decision-making: A \textbf{text-as-data} approach
  \end{itemize}
\end{itemize}