---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 8: Managing}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/week8.pdf")
```

# Today's plan

\begin{itemize}
\small
  \item Quick go-around: Cases for final essay
  \vspace{0.2cm}
  \item Recap from Week 7: Parties and party system
  \begin{itemize}
    \item Parties in democracy and where parties come from
    \item Party nationalization (Scott Morgenstern 2017)
    \item Extra: Parties in authoritarian regimes
  \end{itemize}
  \vspace{0.2cm}
  \item Refresher lecture: Bureaucratic politics
  \vspace{0.1cm}
  \begin{itemize}
    \item Bureaucracies and politics of delegation (Kiewiet and McCubbins 1991)
    \item Corruption and ways to counter it (Treisman 2000; Olken 2007)
    \item Extra: "Bureaucracies" as Party-state agents in China (Yang 2004; Cai 2014)
    \item Reading exercise: "Corruption and Inefficiency in Local Public Infrastructure"
  \end{itemize}
\end{itemize}

# Looking back: What we have accomplished

\begin{itemize}
\small
  \item Varieties of constitutional democracy
  \vspace{0.1cm}
  \begin{itemize}
    \item Executive system: presidential, parliamentary and semi-presidential
    \item Legislative system: uni- or bi-cameral
    \item Electoral system: majoritarian/plurality, PR and mixed-member (and the resulting party system)
    \item Central-local relations: federal, unitary and decentralized
  \end{itemize}
  \vspace{0.3cm}
  \item Democratic constitutional design is a response to "\textbf{Madison's Dilemma}"
  \vspace{0.1cm}
  \begin{itemize}
    \item The goal is to create a government strong enough to govern while limited enough to protect civil liberties
    \item Two visions of democracy: Majoritarian Model (power concentration) versus Consensus Model (power dispersion)
  \end{itemize}
\end{itemize}

# Recap: Parties and party system

\begin{itemize}
\small
  \item Why parties: Political parties serve as the pillar of representative democracy in modern era
  \vspace{0.1cm}
  \begin{itemize}
    \item Recruit and train political elites (and as we discussed this may no longer hold)
    \item Mobilize the mass and political participation
    \item Build and coordinate the links between voters and elected representatives
  \end{itemize}
  \vspace{0.2cm}
  \item Where parties and party system come from
  \vspace{0.1cm}
  \begin{itemize}
    \item Institutional explanation: Duverger's law (electoral disproportionality and strategic voting)
    \item Sociological explanation: Cleavage theory of political parties
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/uk_greens.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
\url{https://www.bbc.co.uk/news/articles/cr453rvy6kvo}
\end{center}

# An ongoing partisan realignment in the UK?

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/uk_greens_2.pdf")
```

# "Party nationalization" (Scott Morgenstern 2017)

\begin{columns}
\column{0.38\linewidth}
  \centering
  \includegraphics[scale=0.30]{Figs/national}
\column{0.5\linewidth}
\hspace{0.15cm}
  \begin{itemize}
  \small
    \item The degree of homogeneity in the geographic distribution of a party's vote
    \vspace{0.4cm}
    \item The degree to which national events are reflected in the change in a party's electoral support in all regions of the country
    \end{itemize}
\end{columns} 

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/uk_green_cropped.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/uk_reform_cropped.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/germany_greens.pdf")
```

# Extra: Parties in authoritarian regimes

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/geddes.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/smith1.pdf")
```

---

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/smith2.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/smith_table.pdf")
```

# China: A "modernised" Leninist party managing to resolve the challenge of political succession

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/fewsmith.pdf")
```

# Refresher lecture: Bureaucracy and bureaucratic politics

\begin{itemize}
  \item Bureaucracy and politics of delegation (Kiewiet and McCubbins 1991)
  \vspace{0.1cm}
  \begin{itemize}
    \item Bureaucrats, elected politicians and technocrats
    \item Politics of delegation: Moral hazard, adverse selection and agency loss
  \end{itemize}
  \vspace{0.2cm}
  \item Corruption and ways to counter it (Treisman 2000; Olken 2007)
  \vspace{0.1cm}
  \begin{itemize}
    \item Defining and measuring corruption
    \item Petty versus grand corruption
    \item How to counter corruption
  \end{itemize}
  \vspace{0.2cm} 
  \item Extra: "Bureaucracies" as Party-state agents in China (Yang 2004; Cai 2014)
\end{itemize}

# Bureaucrats versus elected politicians

\begin{itemize}
\small
  \item Bureaucrats perform most of the tasks of government; the formulation of policies and their implementation cannot get away from bureaucrats
  \vspace{0.2cm}
  \item Elected politicians face term limits or are put under the pressure of reelection, thus being incentivized to be responsive to voters or perform well to be reelected
  \vspace{0.1cm}
  \begin{itemize}
    \item Policy-seeking politicians
    \item Office-seeking politicians
  \end{itemize}
  \vspace{0.2cm}
  \item What drives bureaucrats' behaviors at the absence of electoral accountability? An utopian idea: Technocrats in politics?
\end{itemize}

# Bureaucratic politics through the lens of delegation

\begin{itemize}
\small
  \item Delegation involves a dyad interaction between principal(s) and agent(s)
  \vspace{0.1cm}
  \begin{itemize}
    \item Principal selects agent(s) to complete certain tasks that require certain expertise
    \item Agent(s) have incentives to shirk from their assigned responsibilities for their self-interest
    \item Principal may not be aware of agent(s) shirking due to \textbf{information asymmetry}
    \item The \textbf{agency loss} incurred during delegation can be enormous
  \end{itemize}
  \vspace{0.2cm}
  \item Delegation is ubiquitous in democratic politics (see the next slide) for the sake of expertise (and implicitly), efficiency
  \vspace{0.1cm}
  \begin{itemize}
    \item Voters (principals) delegate governance to politicians (agents)
    \item Politicians (principals) delegate governance tasks to bureaucrats (agents)
  \end{itemize}
\end{itemize}

# Elected democratic government as "principaled" agents (of whom)?

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/besley_cropped.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/delegation.pdf")
```

# How delegation leads to agency loss due to information gap between principal and agent(s)

\begin{columns}
\column{0.38\linewidth}
  \centering
  \includegraphics[scale=0.25]{Figs/michigan}
\column{0.5\linewidth}
  \begin{itemize}
  \small
    \item \textbf{Adverse selection} occurs when the principal cannot fully observe the agent's "type" (e.g., motivation, ability and preferences) (Akerlof 1970)
    \vspace{0.5cm}
    \item \textbf{Moral hazard} occurs when the agent has the opportunity to take actions that are hidden from the principal
  \end{itemize}
\end{columns}

# Corruption as an agency loss

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/olken_indonesia.pdf")
```
\vspace{0.3cm}
\begin{itemize}
  \item "Police patrol" mechanism: A "top-down" inspection (e.g., central audits) imposed by the "principal" to curb or deter corruption
  \vspace{0.1cm}
  \item "Fire alarm" mechanism: Use "third-party" stakeholders (e.g., grassroots organizations) to monitor corruption 
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/olken_monitoring.pdf")
```
\vspace{0.4cm}
\begin{center}
Which one works better?
\end{center}


# Corruption Perceptions Index by Transparency International (2025)

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ti_2025.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
\url{https://www.transparency.org/en/cpi/2025}
\end{center}

# Measuring corruption

\begin{itemize}
  \item Using surveys
  \vspace{0.1cm}
  \begin{itemize}
    \item Experts versus citizens?
    \item Perceptions versus experiences?
  \end{itemize}
  \vspace{0.4cm}
  \item Using legal investigations and prosecutions (or other official records) (e.g., Yuhua Wang's China’s Corruption Investigations Dataset)
\end{itemize}

# Types of corruption: Decentralized versus centralized (William Easterly 2001)

\begin{columns}
\column{0.38\linewidth}
  \centering
  \includegraphics[scale=0.25]{Figs/easterly}
\column{0.5\linewidth}
  \begin{itemize}
  \small
    \item Under \textbf{decentralized} corruption, there are many bribe takers, and their imposition of bribes is not coordinated among them
    \vspace{0.3cm}
    \item Under \textbf{centralized} corruption, a government leader organizes all corruption activity in the economy and
determines the shares of each official in the ill-gotten proceeds
  \end{itemize}
\end{columns}

# Types of corruption

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/ang.pdf")
```
\vspace{0.1cm}
\begin{itemize}
\small
  \item Speed money: Petty bribes paid to low-level bureaucrats to circumvent hurdles or delays
  \item Access money: High-stakes rewards offered by elite capitalists to powerful officials in exchange for exclusive, lucrative privileges
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ang_compare.pdf")
```

# "Causes" of corruption (Treisman 2000)

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/corruption_cover.pdf")
```

---

\begin{itemize}
  \item Corruption is lower in
  \vspace{0.1cm}
  \begin{itemize}
    \item countries with common law systems
    \item Britain and its former colonies
    \item countries with a Protestant tradition
    \item democratic countries and those with a freer press and more vigorous civic associations
    \item more economically developed countries, where populations are more educated and literate, and where the normative separation between "public" and "private" is clearer
    \item countries with higher relative salaries in public office
    \item countries where political stability is greater
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/corruption_results.pdf")
```

# Bureaucrats as Party-state agents in China

\begin{columns}
\column{0.38\linewidth}
  \centering
  \includegraphics[scale=0.30]{Figs/yang_levi}
\column{0.5\linewidth}
\hspace{0.15cm}
  \begin{itemize}
  \small
    \item Govt officials are selected by the ruling party (redness/loyalty versus competence/expertise)
    \vspace{0.1cm}
    \item The ruling party provides career incentives for officials act upon (e.g., the promotion tournament model of bureaucratic behaviors); see Oi (1998)
    \vspace{0.1cm}
    \item China presents a case of "adverse" accountability, where all accountability measures such as inspections and audits come from above
    \vspace{0.1cm}
    \item Media censorship is strategically imposed to reveal local official corruption (Lorentzen 2014)
  \end{itemize}
\end{columns} 

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/glf.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/jiang_china.pdf")
```