---
title: |
  | 
  |
  | Doing Political and Social Research 
  |
  |
author: |
  | Week 3
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Assessment: Critical review essay

\begin{itemize}
  \item A \textbf{2000-word} critical review of a journal article that uses one of the methods covered in the module
  \vspace{0.2cm}
  \item A link to a list of suggested journal articles can be found on Moodle (or your own choice approved by the module convenor)
  \vspace{0.2cm}
  \item Consider the following in your review:
  \vspace{0.1cm}
  \begin{itemize}
    \item What is the research process: Research question, theory, concepts, measurement, approach to explanation?
    \item How appropriate was the method? How well was it implemented? What other methods could have been used?
  \end{itemize}
  \vspace{0.2cm}
  \item Similar to our weekly seminar discussions
\end{itemize}

---

\begin{itemize}
  \item Choose one of the articles listed
  \item Use the essay coversheet
  \item Enter your student ID in the field of Submission Title
  \item Submit your work by \textbf{Midday, 25 November 2024}
\end{itemize}
\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/turnitin.pdf")
```

# Strategic reading: The Berkeley style

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucb.pdf")
```
\vspace{0.5cm}
\begin{itemize}
  \item \textbf{Preview}: Get as much information about the reading before you actually read it
  \vspace{0.35cm}
  \item \textbf{Annotating}: Read with a pencil and making notes as you read
  \vspace{0.35cm}
  \item \textbf{Analyzing}: Break the reading apart to see how different parts relate to each other
  \vspace{0.35cm}
  \item \textbf{Responding}: Think again how the reading relates the topic of each week; come up with summary, comments and questions
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucb.pdf")
```
\vspace{0.5cm}
\begin{itemize}
  \item \textbf{Preview}: Get as much information about the reading before you actually read it
  \vspace{0.35cm}
  \begin{itemize}
    \item Why are you reading it (\textbf{hint: read session statement on Moodle})?
    \item What can the title tell you about the reading?
    \item How is the reading structured?
    \item What info is available from the introductory and concluding paragraphs?
    \item What is conveyed by the opening/topic sentences?
    \item What is the main thesis? What are the main points?
  \end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucb.pdf")
```
\vspace{0.5cm}
\begin{itemize}
  \item \textbf{Annotating}: Read with a pencil and making notes as you read
  \vspace{0.35cm}
  \begin{itemize}
    \item Mark the thesis and main points
    \item Write signposts to reorganize the reading
    \item Note the main evidence to support the argument
    \item Mark key terms and unfamiliar words and note confusing/challenging parts
    \item Write down your questions/comments and/or experiences
    \item Underline the sources (\textbf{hint: skim through the reference list and further readings on Moodle})
  \end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucb.pdf")
```
\vspace{0.5cm}
\begin{itemize}
  \item \textbf{Analyzing}: Break the reading apart to see how different parts relate to each other
  \vspace{0.35cm}
  \begin{itemize}
    \item Examine the evidence: Adequate? Relevant?
    \item Evaluate sources: Relevant? Credible? Current?
    \item Consider assumptions: What do we have to believe for the argument to be true?
    \item Watch out for author bias
  \end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucb.pdf")
```
\vspace{0.5cm}
\begin{itemize}
  \item \textbf{Responding}: Think again how the reading relates the topic of each week; come up with summary, comments and questions
  \vspace{0.35cm}
  \begin{itemize}
    \item Write a response to the reading
    \item Keep a notebook
    \item Discuss the reading with others
  \end{itemize}
\end{itemize}

# Epistemology in political and social research

\vspace{0.4cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/epis_barry.pdf")
```

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/epis_binary.pdf")
```

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/epis_spectrum.pdf")
```

# Find your research questions

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/mason.pdf")
```
\begin{center}
\begin{scriptsize}
Mason, J. 2002. \textit{Qualitative researching} (2nd). London: Sage, p.18.
\end{scriptsize}
\end{center}

---

\begin{itemize}
  \item List the relevant "scholarly" literature(s); arrange the literature(s) by topic/keyword if you'd like
  \vspace{0.2cm}
  \item Identify the gap, such as
  \begin{itemize}
    \item Theoretical/empirical tensions or contradictions
    \item Different or alternative theoretical perspectives
    \item Different methodological approach (e.g., build on more inductive work to propose some hypothesis-testing deductive approach)
    \item Extension to new context (e.g., different region, time period)
    \item Some conventional wisdom left untested or understudied (e.g., critical element not engaged with, need to consider other conditions)
  \end{itemize}
  \vspace{0.2cm}
  \item Use literature to think through your RQs (and supervisor preferences)
\end{itemize}