---
title: |
  | 
  |
  | CP$^2$ Week 8: Co-optation and Representation
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Recap from last week

\begin{itemize}
\small
  \item Dictator's dilemma: Finding balance between coercion and co-optation
  \vspace{0.2cm}
  \item Coercion and control: Preventing elite split and collective action
  \vspace{0.2cm}
  \item Co-optation and representation: Sharing the stake of the regime's survival
  \vspace{0.1cm}
  \begin{itemize}
    \item Local elections and legislatures
    \item Party membership and public employees
    \item Social policy and welfare state
  \end{itemize}
  \vspace{0.2cm}
  \item Recent "innovations" in Chinese authoritarian governance
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="45%", fig.align="center"}
knitr::include_graphics("Figs/doc.pdf")
```
\vspace{0.15cm}
\begin{center}
\small
"\textbf{I Want To Be A People's Representative}" by Zhitan Jia (2013)
\end{center}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/media.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/thinktanks.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/newsletters.pdf")
```
\vspace{0.6cm}
\begin{itemize}
  \item \url{https://www.neican.org/} (Australia-based China watcher)
  \item \url{https://www.prcleader.org/} (US-based scholars of Chinese politics)
  \item \url{https://www.chinafile.com/} (published by Asia Society)
  \item \url{https://www.pekingnology.com/} (Former Xinhua employee)
  \item \url{https://madeinchinajournal.com/} (oversea China scholars)
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/cdt.pdf")
```
\vspace{0.2cm}
\begin{center}
\url{https://chinadigitaltimes.net/}
\end{center}

---

```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/xin_talk.pdf")
```

# Week 10 discussion on "Two Sessions"

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/as_two.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
\url{https://asiasociety.org/policy-institute/key-takeaways-chinas-two-sessions-2025}
\end{center}

# Week 10 discussion on "Two Sessions"

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/ch_two.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
\url{https://www.chathamhouse.org/2025/03/chinas-two-sessions-what-did-we-learn-about-chinese-economy}
\end{center}

# Co-optation through natl and local legislatures

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.42]{Figs/manion}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item Representation in a Leninist party regime might take one of the following forms: \textbf{Mandate}, \textbf{Lennist} and \textbf{Trustee}
  \vspace{0.25cm}
  \item Chinese legislature as an institution of "authoritarian parochialism:" Deliever targeted public goods to geograpgic constituency through extra-legislative means
  \vspace{0.25cm}
  \item The People's Congress system provides a key solution to alleviate the classical "dictator's dilemma:" Information and loyalty
  \vspace{0.25cm}
  \item Responsive without being accountable?
\end{itemize}
  \endcol
\endcols

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/res_manion.pdf")
```
\vspace{0.1cm}
\begin{small}
Survey data shows that Chinese local congressmen and women view themselves and act as "delegates."
\end{small}

# Co-optation through local elections

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/election}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item Currently, the Party allows for village elections as well as local elections for the People's Congress delegates
  \vspace{0.4cm}
  \item While elections are not uncommon among authoritarian regimes, few have national and executive elections
  \end{itemize}
  \endcol
\endcols

# Co-optation through local elections

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/election}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item In the case of village elections, they were introduced in the 1990s to tackle peasant grienvances when it comes to fees collection (Berstein and Lu 2003)
  \vspace{0.4cm}
  \item Broadly speaking, they provided an opportunity to monitor local officials and hold them accountable (Mayling Birney 2010)
  \end{itemize}
  \endcol
\endcols

# Co-optation through Party

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/party}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item People who joined the party under Mao were more likely to report ideological reasons; those in the post-1989 cohort mostly joined for the career benefits, and not just to work for the party or the government (Dickson 2013)
  \vspace{0.3cm}
  \item Jiang Zemin's "Three Represents" opened the Party's door to entrepreneurs, and it appears as a good business opportunity especially when these entrepreneurs were "elected" into the People's Congress (Truex 2014)
  \vspace{0.3cm}
  \item The demographic composition of the Party Congress has changed a lot in the past two decades
  \end{itemize}
  \endcol
\endcols

# "Returns to Office in a Rubber Stamp Parliament" (Truex 2014)

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/truex_res.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/party_join.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/party_elite.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/hanzhang_both.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/liu_res.pdf")
```

# "Social Policy in Developing Countries" (Mares and Carnes 2009)

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/mares.pdf")
```

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/welfare.pdf")
```

# Three models of social protection in China (Solinger 2015)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.4]{Figs/solinger}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
    \item To extend assistance to accord with social citizenship rights (or to fulfill the Confucian concept of the rite of benevolence)
    \vspace{0.2cm}
    \item To offer subsidies to attain support or to pacify anger and silence demands from the poor
    \vspace{0.2cm}
    \item To grant benefits (education, health care) to enhance the nation's productivity
  \end{itemize}
  \endcol
\endcols

# "Innovations" in Chinese authoritarian governance

\begin{itemize}
  \item "Fragmented" and "responsive" authoritarianism: Controlled pluralism and accountability in the political system (Mertha 2009; Qiaoan and Teets 2020)
  \vspace{0.2cm}
  \item From Hukou to fine-grained, pervasive grassroots "control"
  \vspace{0.1cm}
  \begin{itemize}
    \item Internet and digital space (Yang 2015; Roberts 2018)
    \item Urban street-level grid (or celluar) management
    \item Rural household-level targetting poverty alleviation scheme
  \end{itemize}
  \vspace{0.2cm}
  \item "Campaign-style" governance (Zhou 2022)
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/books_internet.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/hukou.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/hukou_flow.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/urban.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/urban_tab.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/grid.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/xinjiang.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/zhou.pdf")
```

# Conclusion

\begin{itemize}
  \item Dictators rarely rule by force only, and they will try all means to induce compliance and loyalty
   \vspace{0.3cm}
  \item Authoritarian governance involves intense coalition building (within and outside the inner circle) and strikes the general equilibrium of coercion and co-optation
   \vspace{0.3cm}
  \item China is no exception, and yet a number of puzzles remain to be answered:
  \vspace{0.1cm}
  \begin{itemize}
    \item How can the Party sustain a system of responsiveness without complete electoral accountability?
    \item How can the Party strike the balance between the stick and carrot?
  \end{itemize}
\end{itemize}
