---
title: |
  |
  |
  | Investigating the Social World
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Masterlcass (2025 version) announced

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/masterclass.pdf")
```
\vspace{0.2cm}
\begin{center}
Invitation to audit will be sent to all PG students
\end{center}

# Before Reading Week: Varieties of social research

\begin{itemize}
\small
  \item ISW is a special module, as we aim to reflect on the production of "valid" knowledge about the social world
  \vspace{0.1cm}
  \begin{itemize}
    \item Know what you do and explain the objectives of your scholarly endeavors;
    \item Know the differences between various objectives and engage other researchers' work constructively and effectively
  \end{itemize}
  \vspace{0.2cm}
  \item Social research involves a series of careful thinking about epistemology, theory-building and methodology; hence \textbf{theory-empirics alignment} is key to rigour
  \vspace{0.1cm}
  \begin{itemize}
    \item \textbf{Reveal nuances of the real world}: Constructivism; (interpretative) theory as approach; interpretation; qualitative; inductive/abductive
    \item \textbf{Search for the "ultimate" truth}: Positivism; (explanatory) theory as paradigm; explanation; quantitative; deductive
  \end{itemize}
\end{itemize}

# Exceptions exist -- norms are constantly being contested and redefined

\pause
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/epis_bin.pdf")
```

# Thinking beyond your case to boost contribution

\vspace{0.2cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/lachlan.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/lachlan2.pdf")
```

# Weeks 6-10: Analytical (or investigative) frameworks

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/guest.pdf")
```
\vspace{0.5cm}
\begin{itemize}
\small
  \item Rational choice theory (Dr Chao-Yo Cheng)
  \item Psychosocial framing (Professor Silvia Posocco)
  \item Decolonizing social research (Dr Ivana Bevilacqua)
  \item Intersectional sensibilities (Ms Joy Brooks-Gilzeane)
  \item Situating lived experience (Dr Sarah Marks)
\end{itemize}

# Assessment II: Analytical essay (60%)

\begin{itemize}
  \item You will write an essay of 2,500 words in length
  \vspace{0.3cm}
  \item The idea is to choose \textbf{TWO} frameworks introduced after Reading Week to analyze an issue or problem of your choice
  \vspace{0.3cm}
  \item You are free to carry out any necessary outside/additional research; cite references properly (e.g., APA or MLA)
  \vspace{0.3cm}
  \item You can come to my office hours to discuss your preliminary responses with me; the goal is not to give you a mark before the deadline but to clear any confusion
  \vspace{0.3cm}
  \item Instructions and sample submissions available on Moodle now
\end{itemize}

---

\begin{center}
\textbf{Lecture 7: Rational choice}
\end{center}
\vspace{0.7cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/cover.pdf")
```
\vspace{0.1cm}
\pause
\begin{center}
\scriptsize
"Anyone fancies an end-of-term happy hour?"
\end{center}

# Plan for today

\begin{itemize}
  \item Rational choice and beyond
  \vspace{0.4cm}
  \item Workshop: "Analytical narratives"
  \vspace{0.1cm}
  \begin{itemize}
    \item Principles
    \item Critiques and responses
    \item Examples
    \item Exercise
  \end{itemize}
\end{itemize}

# Barbara's "verdict:" Breaking mis-perceptions

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/castles}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
    \item Rational choice does not assume people are "selfish"
    \vspace{0.3cm}
    \item Rational choice does not assume people are driven by material interests
    \vspace{0.3cm}
    \item Rational choice does not assume preferences are stable or fixed
    \vspace{0.3cm}
    \item Rational choice does not ignore/discard ideology and/or history
  \end{itemize}
  \endcol
\endcols

# Rational choice is an "approach," not much of a "theory" per se

```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/voice.pdf")
```
\vspace{0.2cm}
\begin{itemize}
\small
  \item Rational choice highlights the importance of "rationality" and yet how rationality should treated as a construct, not a given constant
  \item Rational choice does not dictate a single or universal explanation/prediction
  \item Rational choice provides a \textbf{micro-foundation} to research the social world
\end{itemize}

# John Levi Martin's take on rational choice (theory)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/martin}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
    \small
    \item A useful "modelling exercise" for deductive reasoning
    \vspace{0.3cm}
    \item Cautioning against using it as a "total" theory of society (which can grow into a form of intellectual imperialism)
    \vspace{0.3cm}
    \item \textbf{Caveats abound}, alas -- Do we have a "choice?" How about structures? In favor of parsimony at the sacrifice of complexities and nuances? Ad hoc? How do we handle unintended or unexpected outcomes?
    \vspace{0.3cm}
    \item \textbf{Question for all}: To what extend would he align with Geddes?
  \end{itemize}
  \endcol
\endcols

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/envy_book.pdf")
```

# Examples: "Envy in Politics" (McClendon 2018) 

\begin{itemize}
  \item \textbf{Paradox}: Why rich people support redistribution? Why do governments fail to provide and act responsively?
  \vspace{0.2cm}
  \item \textbf{Argument}: \textit{Status sentiment} matters -- envy, spite and pursuit of admiration will influence political behaviors 
  \vspace{0.2cm}
  \item \textbf{Literature}: Drawing on anthropology, social psychology, behavior economics and so on
  \vspace{0.2cm}
  \item \textbf{Empirics}: Case studies and quantitative analysis, especially surveys and election/admin data (see "Applications")
  \vspace{0.2cm}
  \item \textbf{Implications}: Inequality is the key driving force of various political issues we are observing, but may not manifest in a way we expect 
\end{itemize}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/envy_reg.pdf")
```

# Using rational choice to research the social world

\begin{itemize}
  \item Identify "paradox" to challenge the "conventional" wisdom (e.g., things taken for granted or so but turn out to be quite counterintuitive)
  \vspace{0.3cm}
  \item Define "actors" involved, laying down their different outcomes and actions
  \vspace{0.3cm}
  \item Set "utilities" to help the search for their "optimal" strategies; whether one action will be taken or not, can be an "interactive" and "repeated" process
  \vspace{0.3cm}
  \item Note that "utilities" do not come out of nowhere, and actors may be constrained by their "information" environment
\end{itemize}

# Example: Prisoner's dilemma

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/pd_1.pdf")
```
\vspace{0.2cm}
\begin{itemize}
\small
  \item Suppose A and B are arrested; the police puts them in two different interrogation rooms
  \item If both confess, both get 10 years in prison
  \item If none of them confesses, both get 5 years in prison
  \item For both A and B, they can get away if only the other side confesses, and the other side will get 20 years in prison
\end{itemize}

# Example: Prisoner's dilemma

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/pd_2.pdf")
```
\vspace{0.2cm}
\begin{itemize}
  \item A would not know what B will do (unless they make a deal beforehand)
  \item If B confesses, A should confess as well
  \item If B does not confess, A should confess, too
  \item So regardless what B will do, \textbf{A should confess anyway}
\end{itemize}

# Example: Prisoner's dilemma

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/pd_3.pdf")
```
\vspace{0.2cm}
\begin{itemize}
  \item Likewise, B would not know what A will do
  \item If A confesses, B should confess as well
  \item If A does not confess, B should confess, too
  \item So regardless what A will do, \textbf{B should confess anyway}
\end{itemize}

# Example: Prisoner's dilemma

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/pd_4.pdf")
```
\vspace{0.2cm}
\begin{itemize}
  \item As a result, both actors are stuck in the situation such that both will confess
  \item This is counterintuitive, as ideally none of them should
  \item What can we do to get away from this conundrum -- of course, we would not want to set suspects on loose, but let's put that aside
  \item Any other comments or questions?
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/analytic.pdf")
```
\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/analytic_people.pdf")
```

# Rationale and principle

\begin{itemize}
\small
  \item Using game theoretical models to explain particular political economy cases that may generate broad theoretical implications, such as
  \vspace{0.1cm}
  \begin{itemize}
    \item Political order as equilibria
    \item How domestic politics and intl relations shape each other
    \item Institutions matter for political economy of development
  \end{itemize}
  \vspace{0.3cm}
  \item The process is deductive, developing "thick" narratives as "guided" by rational-choice like argument (as some sort of empirical tests)
  \vspace{0.3cm}
  \item Inspires the EITM (empirical implications of theoretical models) school in the United States
  \vspace{0.3cm}
  \item Check out examples in the book for demonstration
\end{itemize}

# Critiques: A case of "excessive" ambition (Elster 2000)

\begin{itemize}
  \item Too ambitous -- is rational choice theory a "universal" theory? Theory- or problem-driven (Green and Shapiro 1994)?
  \vspace{0.3cm}
  \item The assumption for "rational" actors is too much, while rational choice is able to accommodate non-material or non-selfish interests
  \vspace{0.3cm}
  \item While the model is based on "individuals," aggregation of preferences may not be that straightforward (e.g., Arrow's impossibility theorem)
  \vspace{0.3cm}
  \item How about "beliefs" or "intentions?" And uncertainty, after all? The provided explanation is often ad hoc and unrealistic
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/eitm.pdf")
```
\vspace{0.3cm}
\begin{center}
\url{https://eitminstitute.org/}
\end{center}

# Exercise

What is the issue you'd like to deal with in your analytical essay? Share with your group members -- do you find the approach of "analytical narratives" useful, even with all the caveats associated with the rational choice framework?

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/books.pdf")
```

---

\vspace{0.5cm}
\begin{center}
Thank you!
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bbk_tv.pdf")
```
\vspace{0.3cm}
\begin{center}
\texttt{c.cheng@bbk.ac.uk}
\end{center}
