---
title: |
  |
  |
  | Investigating the Social World
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Masterlcass (2025 version) announced

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/masterclass.pdf")
```
\vspace{0.2cm}
\begin{center}
Invitation to audit will be sent to all PG students
\end{center}

# Before Reading Week: Varieties of social research

\begin{itemize}
\small
  \item ISW is a special module to reflect on the production of "valid" knowledge about the social world
  \vspace{0.1cm}
  \begin{itemize}
    \item Know what you do and explain the objectives of your scholarly inquiry;
    \item Know the differences between various objectives and engage other researchers' work constructively and effectively
  \end{itemize}
  \vspace{0.2cm}
  \item Social research involves a series of careful thinking about epistemology, theory-building, and methodology (hence theory-empirics alignment is key to rigour)
  \vspace{0.1cm}
  \begin{itemize}
    \item \textbf{Reveal nuances of the real world}: constructivism; (interpretative) theory as approach; interpretation; qualitative; inductive/abductive
    \item \textbf{Search for the "ultimate" truth}: positivism; (explanatory) theory as paradigm; explanation; quantitative; deductive
  \end{itemize}
\end{itemize}

# Exceptions exist -- norms are constantly being contested and redefined

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/epis_bin.pdf")
```

# Thinking beyond your case to boost contribution

\vspace{0.4cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/lachlan.pdf")
```

# Weeks 6-10: Analytical (or investigative) frameworks

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/guest.pdf")
```
\vspace{0.5cm}
\begin{itemize}
\small
  \item Rational choice theory (Dr Chao-Yo Cheng)
  \item Psychosocial framing (Professor Silvia Posocco)
  \item Decolonizing social research (Dr Ivana Bevilacqua)
  \item Intersectional sensibilities (Ms Joy Brooks-Gilzeane)
  \item Situating lived experience (Dr Sarah Marks)
\end{itemize}

# Assessment II: Analytical essay (60%)

\begin{itemize}
  \item You will write an essay of 2,500 words in length
  \vspace{0.3cm}
  \item The key idea is to choose \textbf{TWO} frameworks introduced after Reading Week to analyze an issue or problem of your choice
  \vspace{0.3cm}
  \item You are free to carry out any necessary outside/additional research; cite references properly (e.g., APA or MLA)
  \vspace{0.3cm}
  \item You can come to my office hours to discuss your preliminary responses with me; the goal is not to give you a mark before the deadline but to clear any confusion
  \vspace{0.3cm}
  \item Instructions and sample submissions available on Moodle now
\end{itemize}

---

\begin{center}
\textbf{Lecture 7: Rational choice}
\end{center}
\vspace{0.7cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/cover.pdf")
```
\vspace{0.1cm}
\pause
\begin{center}
\scriptsize
"Anyone fancies an end-of-term happy hour?"
\end{center}

# Plan for today

\begin{itemize}
  \item Rational choice and beyond
  \vspace{0.4cm}
  \item Workshop: "Analytical narratives"
  \vspace{0.1cm}
  \begin{itemize}
    \item Principles
    \item Critiques and responses
    \item Examples
    \item Exercise
  \end{itemize}
\end{itemize}

# Barbara's "verdict" on rational choice

\begin{itemize}
  \item There are many \textbf{misperceptions} about rational choice
  \begin{itemize}
    \item Rational choice does not assume people are "selfish"
    \item Rational choice does not assume people are driven by material interests
    \item Rational choice does not assume preferences are stable or fixed
    \item Rational choice does not ignore/discard ideology and/or history
  \end{itemize}
  \vspace{0.2cm}
  \item Rational choice is an \textbf{approach} (i.e., ways of seeing), not a theory
  \begin{itemize}
    \item Rational choice highlights the importance of "rationality" 
    \item Rational choice does not dictate a single or universal explanation/prediction
    \item Rational choice provides a micro-foundation for us to research the social world
  \end{itemize}
\end{itemize}

# Martin's take

TBA.

# Rational choice and rationality

# Ratonal choice as an analytical approach

# Example: Prisonor's dilemma

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/analytic.pdf")
```
\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/analytic_people.pdf")
```

# Rationale and principle

# Critque and responses

# Example

# Exercise

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/books.pdf")
```

---

\vspace{0.5cm}
\begin{center}
Thank you!
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bbk_tv.pdf")
```
\vspace{0.3cm}
\begin{center}
\texttt{c.cheng@bbk.ac.uk}
\end{center}
