---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 6: Choosing}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week6.pdf")
```

# Today's plan

\begin{itemize}
  \item Logistics: Start thinking about your final essay 
  \vspace{0.3cm}
  \item Recap from Week 5
  \begin{itemize}
    \item Executive structures and legislatures
    \item Supreme/constitutional courts and judicial review
  \end{itemize}
  \vspace{0.3cm}
  \item Refresher lecture: Elections and parties
  \vspace{0.1cm}
  \begin{itemize}
    \item Electoral system: Definitions and types
    \item Political consequences of electoral system
    \item Extra: Elections in authoritarian regimes
    \item Reading exercise: "Getting Good Politicians into Politics"
  \end{itemize}
\end{itemize}

# Recap: Presidential versus parliamentary democracies

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/type.pdf")
```

# Recap: (A)symmetry and (in)congruence bicameralism 

```{r echo=FALSE, out.width="55%", fig.align="center"}
knitr::include_graphics("Figs/legislature/bicameralism.pdf")
```
\vspace{0.2cm}
\begin{itemize}
\small
  \item (A)Symmetry: Relative powers between two chambers/houses
  \item (In)Congruence: Relative preferences (and interests) between two chambers/houses
\end{itemize}

---

\vspace{1cm}
"A presidential democracy with a symmetric and incongruent bicameral system is most prone to regime collapse."\
\pause
\vspace{0.5cm}
Why?

# Recap: Supreme/constitutional courts and judicial review

\begin{itemize}
  \item Judiciary in the constitution organizes relations between the executive and legislative branches of government:
  \vspace{0.2cm}
  \begin{itemize}
    \item \textbf{Parliamentary supremacy}: a principle according to which judges' decisions remain subordinate to decisions of the legislative majority (e.g., UK before 2009)
    \item \textbf{Judicial review}: the ability of a country's high court to invalidate laws the legislature has enacted by declaring them unconstitutional (e.g., USA, Germany and UK after 2009)
  \end{itemize}
  \vspace{0.3cm}
  \item What explains the adoption of either principle? What is the "politics" behind the operation of these principles?
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/trump.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://www.bbc.co.uk/news/articles/cd9g0e7zd8wo}
\end{center}

---

```{r echo=FALSE, out.width="62%", fig.align="center"}
knitr::include_graphics("Figs/trump_sc_vote.pdf")
```

# Refresher lecture

\begin{itemize}
  \item Electoral systems
  \vspace{0.1cm}
  \begin{itemize}
    \item Definition and types: Plurality/majoritarian, proportional representation and mixed-member systems
    \item Policy consequences of electoral systems: Income inequality and gender gap
  \end{itemize}
  \vspace{0.2cm}
  \item Duverger's law: Party system as the consequences of electoral systems
  \vspace{0.1cm}
  \begin{itemize}
      \item Disproportionality 
      \item Strategic voting
  \end{itemize}
  \vspace{0.2cm}
  \item Reading exercise: "Getting Good Politicians into Politics"
\end{itemize}

---

\begin{itemize}
\small
  \item Varieties of constitutional democracy
  \vspace{0.1cm}
  \begin{itemize}
    \item Executive system: presidential, parliamentary and semi-presidential
    \item Legislative system: uni- or bi-cameral
    \item \textbf{Electoral system}: majoritarian/plurality, PR and mixed-member (and the resulting party system)
    \item Central-local relations: federal, unitary and decentralized
  \end{itemize}
  \vspace{0.3cm}
  \item Democratic constitutional design is a response to "\textbf{Madison's Dilemma}"
  \vspace{0.1cm}
  \begin{itemize}
    \item The goal is to create a government strong enough to govern while limited enough to protect civil liberties
    \item Two visions of democracy: Majoritarian Model (power concentration) versus Consensus Model (power dispersion)
  \end{itemize}
\end{itemize}

# Number of elections in a year

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/elecs_calendar.pdf")
```

# Elections as instruments of democracy

\begin{columns}
  \column{0.38\linewidth}
  \centering
  \includegraphics[scale=0.25]{Figs/visions}
  \column{0.5\linewidth}
  "A country is democratic to the extent that its most powerful collective decision-makers are selected through \textbf{\textit{fair, honest, and periodic elections}} in which candidates freely compete for votes and in which virtually all the adult population is eligible to vote."
\end{columns}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/polyarchy_cropped.pdf")
```

# How do elections work?

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/cand_voting.pdf")
```
\vspace{0.3cm}
\begin{itemize}
\small
  \item For political parties and candidates: Winning thresholds; term limits; campaign finances
  \vspace{0.1cm}
  \item For voters: Mandatory turnout; secret ballot; absentee voting
  \vspace{0.1cm}
  \item ... and much more
\end{itemize}

# A typology of electoral systems

\begin{itemize}
  \item \textbf{Ballot structure}: Vote for party or individual candidate? Perhaps both?
  \vspace{0.3cm}
  \item \textbf{Electoral formula}: How are votes translated into seats? Should we search for relative vs. absolute majority (i.e., how "permissive" should the formula be)? 
  \vspace{0.3cm}
  \item \textbf{District magnitude}: How many seats should be elected from a district? Single-member or multi-member?
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/elecworld_tree.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/elecworld_map.pdf")
```

---

\begin{itemize}
  \item Plurality and majoritarian systems
  \vspace{0.3cm}
  \item Proportional representation systems
  \vspace{0.3cm}
  \item Mixed-member systems
\end{itemize}

# Plurality and majoritarian systems

\begin{itemize}
  \item Single-member districts
  \vspace{0.1cm}
  \begin{itemize}
    \item First-past-the-post (FPTP)
    \item Majority runoff system
    \item Alternative vote: Voters rank candidates; lowest vote-getter eliminated and his votes
transferred to the next one ranked; repeat until someone has a majority
  \end{itemize}
  \vspace{0.2cm}
  \item Multi-member districts, such as single non-transferrable voting system (or SNTV)
\end{itemize}

---

\begin{columns}
\column{0.38\linewidth}
  \centering
  \includegraphics[scale=0.35]{Figs/india1}
\column{0.5\linewidth}
  \begin{itemize}
  \item The largest (parliamentary) "democracy" in the world
  \vspace{0.2cm}
  \item FPTP for Lok Sabha election
  \vspace{0.2cm}
  \item At least one election for every 5 years
  \end{itemize}
\end{columns} 

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/india2.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
Electronic voting machine (EVM) in India
\end{center}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/india3.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
Govt officials carry EVM to a polling station in Himachal Pradesh (May 12, 2009)
\end{center}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/france1.pdf")
```
\vspace{0.1cm}
\begin{itemize}
  \item Fifth Republic (established by CDG) known for semi-presidential system
  \item Separation of origin for the executive and legislative branches; "cohabitation" is norm
  \item President popularly elected for a fixed term of 5 years
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/france2.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/nationalism.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
"Financial globalisation and Islamist globalisation are helping each other out. Those two ideologies want to bring France to its knees," said Marine Le Pen.
\end{center}

# Proportional representation (PR) systems

\begin{itemize}
  \item Voters cast votes for a particular party; voters are provided with a "list" of candidates from each political party
  \vspace{0.2cm}
  \item Seats are allocated to parties based on each party's vote share (with certain thresholds, such as Germany and Turkey)
  \vspace{0.2cm}
  \item A key component of "consociational" democracy; more "permissive" than plurality and majoritarian systems. Why?
\end{itemize}

---

\begin{itemize}
  \item Closed-list PR: Party determines the ranking of candidates on the party list
  \vspace{0.5cm}
  \item Open-list PR: Voters' preferences determine the ranking of candidates on the party list
\end{itemize}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/hungary.pdf")
```
\vspace{0.1cm}
\begin{center}
Closed-list PR voting ballot in Hungary
\end{center}

---

```{r echo=FALSE, out.width="45%", fig.align="center"}
knitr::include_graphics("Figs/denmark.pdf")
```
\vspace{0.1cm}
\begin{center}
Open-list PR voting ballot in Denmark
\end{center}

# Mixed-member system and parallel voting: Japan

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/japan1.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
In Japanese lower house elections, voters have two choices, but must write in the party name and the candidate name
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/japan2.pdf")
```

# Mixed-member system and parallel voting: Germany

```{r echo=FALSE, out.width="40%", fig.align="center"}
knitr::include_graphics("Figs/germany.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
Here, the voter enters two \textbf{X} marks -- one for a party on the left side, and the other for a district candidate on the right side
\end{center}

# Policy consequences of electoral systems: Income inequality

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/gini_system.pdf")
```

# Policy consequences of electoral systems: Gender gap

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/gender_system2.pdf")
```

# Other questions we ask about democratic elections

\begin{itemize}
\small
  \item Quality of elections: What are the metrics?
  \vspace{0.1cm}
  \begin{itemize}
    \item Candidate quality (e.g., education)
    \item Inclusiveness (e.g., voter registration, voter ID) 
    \item Integrity (e.g., vote buying and ballot box manipulation)
  \end{itemize}
  \vspace{0.2cm}
  \item Electoral violence -- why does it happen and how can we stop it?
  \vspace{0.2cm}
  \item Descriptive representation of marginalized social groups and discrimination
  \vspace{0.2cm}
  \item Why "good" elections do not lead to responsive governments?
\end{itemize}

# Duverger's law: "Plurality rule in single-member districts tends to produce a two-party system"

\begin{columns}
  \column{0.5\linewidth}
  \begin{itemize}
    \item Disproportionality: Largest two parties will get disproportionally more seats than other parties
    \vspace{0.4cm}
    \item Strategic voting: Voters do not wish to "waste" their votes on candidates with no realistic chance of winning
  \end{itemize}
  \column{0.46\linewidth}
  \centering
  \includegraphics[scale=0.5]{Figs/duverger}
\end{columns}

# Electoral disproportionality

\begin{equation}
  \text{Disproportionality} = \frac{1}{n}\sum_i^n |v_i - s_i|, 
\end{equation}
where $n$ refers to the number of parties contesting in the election. Also, $v_i$ and $s_i$ refer to the vote share and seat share of party $i$, respectively.
\vspace{0.2cm}
\begin{figure}
\centering
\includegraphics[scale=0.46]{Figs/dis}
\end{figure}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/uk2015.pdf")
```

# Concluding remarks

\begin{itemize}
  \item The choice of electoral systems reflects a country's preferred democratic vision between (majoritarian versus proportional)
  \vspace{0.3cm}
  \item Different electoral systems can generate different institutional, behavioral, and policy outcomes to democratic governance
  \vspace{0.3cm}
  \item Party system is the product of electoral systems (or is it)?
\end{itemize}

# Reading exercise

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/voxdev.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://voxdev.org/topic/institutions-political-economy/getting-good-politicians-politics}
\end{center}

---

\begin{itemize}
  \item According to Gulzar and Khan (2024), what does it mean to be a "good" politician? Do you agree with their definition?
  \vspace{0.3cm}
  \item How do they produce the empirical evidence to demonstrate "social" encouragement can lead to "better" politicians?
  \vspace{0.3cm} 
  \item Can you find comparable cases of local elections in Pakistan? Will the pattern hold in the case you choose?
\end{itemize}


# Extra: Elections in authoritarian regimes

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/dic_voting.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/elecs_change.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/dic_rep.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://doi.org/10.1016/j.electstud.2024.102791}
\end{center}