---
title: |
  | 
  |
  | CP$^2$ Week 7: Central-Local Relations and Local Governance
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Plan for today

\begin{itemize}
  \item Logistics: Upcoming weeks and assessment
  \vspace{0.5cm}
  \item Lectures and discussion
  \vspace{0.2cm}
  \begin{itemize}
    \item Informal politics and factionalism
    \item Central-local relations and local governance
  \end{itemize}
\end{itemize}

# Revised weekly syllabus

\begin{itemize}
  \item \textbf{Week 1}: State of field
  \vspace{0.1cm}
  \item \textbf{Weeks 2-7}: History and institutions
  \begin{itemize}
    \item From empire to nation-state
    \item Structures of the party-state
    \item Informal politics and factionalism
    \item Central-local relations and local governance
  \end{itemize}
  \vspace{0.1cm}
  \item \textbf{Weeks 8-10}: Governance and development
  \begin{itemize}
    \item Control and co-optation
    \item Ideology and propaganda (guest lecture)
    \item Social protection and redistribution
  \end{itemize}
  \vspace{0.1cm}
  \item \textbf{Week 11}: China as a method (Dim Sum party in Chinatown)
\end{itemize}

# Assessment: Essay (70%)

\begin{itemize}
\small
  \item The questions are currently under review and we expect to release them next week
  \vspace{0.2cm}
  \item You will choose one of the questions and write an original essay of 3,000 words
  \vspace{0.2cm}
  \item You are tasked to demonstrate an adequate understanding of the subject matter and relevant literature, which means
  \vspace{0.1cm}
  \begin{itemize}
    \item You will use any readings or additional references listed on Moodle and in lecture slides
    \item You will need to produce original insights based on the lectures and class discussions
  \end{itemize}
  \item Resources outside the classroom: 1:1 meetings with Chao-yo and Birkbeck's Study Skills workshops
  \vspace{0.2cm}
  \item Due on \textbf{18 April 2024} (midday or 12pm), about a month after the term ends
\end{itemize}

# Assessment: Learning journal (30%)

\begin{itemize}
  \item Briefly introduce the article you choose: Who are the authors? What is the main subeject matter (e.g., research questions, arguments and empirical strategies)? 
  \vspace{0.3cm}
  \item Briefly explain how you will answer the questions: What countries does the article attempt to compare with China? Why? Does the comparison make sense to you? 
  \vspace{0.3cm}
  \item Any questions you would like to ask?
\end{itemize}

# Documentary: "Datong: The Chinese Mayor" (2015)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.26]{Figs/datong_new}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
    \item Directed by Zhou Hao, a former journalist at Southern Weekly
    \vspace{0.4cm}
    \item Won the Best Documentary Prize at Golden Horse Film Festival, among others
    \vspace{0.4cm}
    \item Documented the early years of Geng Yanbo, the mayor of Taiyuan in the province of Shanxi
    \vspace{0.4cm}
    \item What has happened to Mayor Geng since then?
  \end{itemize}
  \endcol
\endcols

# Central-local relations and local governance

\begin{itemize}
  \item China: A federal state in disguise?
  \vspace{0.3cm}
  \item Understanding central-local relations in China: Three analytical approaches
  \vspace{0.3cm}
  \item How Chinese local officials behave and act: Three conceptual models
  \vspace{0.3cm}
  \item Concluding remarks: Local governance as the key of authoritarian resilience and durability
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/structure_new.pdf")
```

# China: A federal state in disguise? (Li, Linda Chelan 2010)

\begin{itemize}
  \item Centralization-decentralization (1950-80s)
  \vspace{0.4cm}
  \item State and market with adjusting roles (1980-90s)
  \vspace{0.4cm}
  \item Demarcating responsibilities over public service provision (2000s-)
\end{itemize}

# China: A federal state in disguise? (Li, Linda Chelan 2010)

\begin{itemize}
  \item Centralization-decentralization (1950-80s)
  \vspace{0.1cm}
  \begin{itemize}
    \item Administrative decentralization was favored, setting of a massive delegation of economic management and planning powers from the central government to the provinces in 1957-8
    \item The next decades till the late 1970s saw cycles of administrative decentralization and re-centralization, only leading to diminished return as the provincial governments gained more discretion
  \end{itemize}
  \vspace{0.4cm}
  \item State and market with adjusting roles (1980-90s)
  \vspace{0.4cm}
  \item Demarcating responsibilities over public service provision (2000s-)
\end{itemize}

# China: A federal state in disguise? (Li, Linda Chelan 2010)

\begin{itemize}
  \item Centralization-decentralization (1950-80s)
  \vspace{0.4cm}
  \item State and market with adjusting roles (1980-90s)
  \vspace{0.1cm}
  \begin{itemize}
    \item Further administrative and fiscal decentralization led to market expansion
    \item The tension between the central and local governments focused on powers to regulate rather than just over resources
    \item the tax-sharing reform in 1994 was conspicuously based on a "rule-based" principle 
  \end{itemize}
  \vspace{0.4cm}
  \item Demarcating responsibilities over public service provision (2000s-)
\end{itemize}

# China: A federal state in disguise? (Li, Linda Chelan 2010)

\begin{itemize}
  \item Centralization-decentralization (1950-80s)
  \vspace{0.4cm}
  \item State and market with adjusting roles (1980-90s)
  \vspace{0.4cm}
  \item Demarcating responsibilities over public service provision (2000s-)
  \vspace{0.1cm}
  \begin{itemize}
    \item The discussion of central-local relations moved from revenues to expenditure responsibilities, which remained decentralized after the 1994 reform
    \item More discussions on how different expenditure responsibilities should be shared between the central and local governments
    \item More key players other than the provinces
  \end{itemize}
\end{itemize}

# A structural approach to understand central-local relations (Zheng 2007)

\begin{itemize}
  \item \textbf{Totalitarian model}: Provincial and local governments have no agency at all and only carries out central policies; decentralization is at the will of the central government
  \vspace{0.3cm} 
  \item \textbf{Pluralist model}: Provincial and local governments pursue their interests and act like interest group in China to lobby and influence the central government
  \vspace{0.3cm}
  \item \textbf{Cellular model}: The central government, when interacting with provincial and local governments, do now follow a formal organizational formula, as complicated bargaining dynamics exist both between the central government and local governments and among local governments
\end{itemize}

# How local officials behave and act (more on this next week)

\begin{itemize}
  \item Promotion tournament system, proposed by Zhou LiAn (2004)
  \vspace{0.3cm}
  \item Pressure system, key scholars include Yang Xueguang (1998) and others
  \vspace{0.3cm}
  \item Project system, developed by Qu Jingdong (2012) and Zhou Feizhou (2012)
\end{itemize}

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/kung_new.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/jiang_new.pdf")
```

# Cai and Treisman (2006): "Does Decentralization Cause China's Economic Model"

\begin{itemize}
\small
  \item China's economy consisted of parallel, diversified provincial economies before the reform
  \vspace{0.2cm}
  \item Decentralization encouraged local reform experiments that allowed new ideas to percolate up from the grassroots and created political checks on central authorities
  \vspace{0.2cm}
  \item Fiscal decentralization gave local officials strong incentives to stimulate economic growth
  \vspace{0.2cm}
  \item Decentralization helped harden budget constraints on enterprises (maybe not SOEs), forcing them to restructure
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="30%", fig.align="center"}
knitr::include_graphics("Figs/susan_new.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
"In the past decade, China was able to carry out economic reform without political reform, while the Soviet Union attempted the opposite strategy. ... Susan Shirk shows that Chinese communist political institutions are more flexible and less centralized than their Soviet counterparts were."
\end{center}

# Concluding remarks

\begin{itemize}
  \item Central-local relations and local governance are key to authoritarian durability and resilience
  \vspace{0.3cm}
  \item China on paper is a unitary state but the picture in real life is more complicated -- how to balance between central control and subnational flexibility is a constant challenge
  \vspace{0.3cm}
  \item Following the Reform, more subnational actors have become political important, such as the prefectures and municipalities
\end{itemize}