---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | AI for Social Research: Computational and Quantitative
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/openai.pdf")
```

# Using LLMs to scale up "interviews" (Geiecke and Jaravel 2024)

\vspace{0.3cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/lse.pdf")
```

---

\begin{itemize}
\small
  \item Use "single" LLM agent (OpenAI's GPT-4o) to carry out interviews on different topics, such as
  \begin{itemize}
    \item Meaning of life
    \item Political preferences (over the 2024 French snap election)
    \item Educational and occupational choices
  \end{itemize}
  \item Evaluate LLM peformance 
  \begin{itemize}
    \item Compare AI-generated transcripts to human experts (in this case, sociologists)
    \item Solicit assessment from respondents, e.g. how well the content of the interview captures their views (or "empathetic") and whether they would prefer to participate in an interview with AI or human researchers
  \end{itemize}
  \item Why AI-led/assisted interviews might be a good idea
  \begin{itemize}
    \item Scale and cost
    \item Privacy and sensitivity bias
  \end{itemize}
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/lse_llm.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/lse_pref.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/lse_quality.pdf")
```

---

# LLMs and quantitative social research 

\begin{itemize}
  \item Survey and experiments: Using computational tools to design survey and simulate responses (see the NORC piece)
  \vspace{0.4cm}
  \item Natural language processing (NLP): Using computational tools to analyze unstructured text, image or audio data (more later in Masterclass)
  \vspace{0.4cm}
  \item Agent-based modeling (ABM): Using computational tools to model and similate the social world and human behaviors (e.g., segregation)
\end{itemize}

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/pa_llm.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/ai_game.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
Demo: \url{https://reverie.herokuapp.com/arXiv_Demo/}
\end{center}

# LLMs for coding and programming

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/logos.pdf")
```
\vspace{0.5cm}
\begin{itemize}
\small
  \item Claude Code (\url{https://www.anthropic.com/claude-code})
  \item Google Gemini Code Assist (\url{https://codeassist.google/})
  \item OpenAI Codex (\url{https://chatgpt.com/codex})
\end{itemize}

# LLM rankings (Openrouter)

```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/rankings.pdf")
```

# AI "agents"

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/llm_agent.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://code.visualstudio.com/api/extension-guides/language-model}
\end{center}

# Exercise: Simulating survey responses

\begin{itemize}
\small
  \item Form a group of three to four students so you can discuss findings among yourselves
  \vspace{0.15cm}
  \item Use LLMs to carry out a "mock" survey
  \begin{itemize}
    \item \textbf{Step 1}: Choose a question from YouGov (\url{https://yougov.co.uk/topics/overview/trackers})
    \item \textbf{Step 2}: Ask LLM the same question to see if they return similar responses (you can try to ask if they have any reference for their answers)
    \item \textbf{Step 3}: Create different "respondents" using YouGov's demographic breakdowns
    \item \textbf{Step 4}: Ask LLM the same question to see if they return similar responses based on the demograpic profile; repeat \textbf{Step 4} using different demograpic profiles
  \end{itemize}
  \vspace{0.15cm}
  \item Questions: How reliable are AI-generated responses, using survey data as the benchmark? How do you compare the performance of different LLMs? Does the topic you choose matter?
\end{itemize}

# Alternative surveys for exercise

\begin{itemize}
\small
  \item British Social Attitudes (\url{https://natcen.ac.uk/british-social-attitudes})
  \vspace{0.5cm}
  \item Pew Research Center (\url{https://www.pewresearch.org/datasets/})
  \vspace{0.5cm}
  \item World Value Survey (\url{https://www.worldvaluessurvey.org/wvs.jsp})
  \vspace{0.5cm}
  \item ... and more
\end{itemize}

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/pa_llm_critical.pdf")
```

# Accessing LLMs using API

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/vscode.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://code.visualstudio.com/api/extension-guides/language-model}
\end{center}

# Concluding remarks

\begin{itemize}
  \item The past three weeks only allow us to scratch the surface, and yet we have touched upon how GenAI can be used for three major research tasks: Analysis, coding and similation
  \vspace{0.5cm}
  \item Learn how to use AI tools creatively to increase your research productivity; critical thinking remains crucial and should not undermine the agency of human researchers
  \vspace{0.5cm}
  \item AI hype may as well be a bubble if we look at it again in a decade -- use it with caution (and make sure you talk to your)
  \vspace{0.5cm}
  \item Transparency is key: Explain how and why you use AI carefully in your research so your peers can evaluate the quality/rigour of your work
\end{itemize}

# "How People Are Really Using Gen AI in 2025" (Havard Business Review)

```{r echo=FALSE, out.width="30%", fig.align="center"}
knitr::include_graphics("Figs/hbr.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://hbr.org/2025/04/how-people-are-really-using-gen-ai-in-2025}
\end{center}

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/zurich.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/mit_response.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/mit_case.pdf")
```