---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 9: Communicating}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/week9.pdf")
```

# Today's plan

\begin{itemize}
\small
  \item Quick go-around: Citations for final essay
  \vspace{0.2cm}
  \item Recap from Week 8: Bureaucratic politics and political corruption
  \begin{itemize}
    \item Sources of corruption
    \item Corruption and inequality (Uslaner)
    \item Reading exercise: "Corruption and Inefficiency in Local Public Infrastructure"
  \end{itemize}
  \vspace{0.2cm}
  \item Refresher lecture: Political communication in the digital age
  \vspace{0.1cm}
  \begin{itemize}
    \item Media freedom and democracy
    \item How social media reshapes political communication
    \item Extra: Media censorship and propaganda in authoritarian regimes
    \item Reading exercise: "AI and Democracy"
  \end{itemize}
\end{itemize}

# Recap: Political corruption

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/Corruption/ti_2025.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
\url{https://www.transparency.org/en/cpi/2025}
\end{center}

# Measuring corruption

\begin{itemize}
  \item Using surveys
  \vspace{0.1cm}
  \begin{itemize}
    \item Experts versus citizens?
    \item Perceptions versus experiences?
  \end{itemize}
  \vspace{0.4cm}
  \item Using legal investigations and prosecutions (or other official records) -- e.g., Yuhua Wang's (Harvard) China’s Corruption Investigations Dataset
\end{itemize}

# Types of corruption: Decentralized versus centralized (William Easterly 2001)

\begin{columns}
\column{0.38\linewidth}
  \centering
  \includegraphics[scale=0.25]{Figs/Corruption/easterly}
\column{0.5\linewidth}
  \begin{itemize}
  \small
    \item Under \textbf{decentralized} corruption, there are many bribe takers, and their imposition of bribes is not coordinated among them
    \vspace{0.3cm}
    \item Under \textbf{centralized} corruption, a government leader organizes all corruption activity in the economy and
determines the shares of each official in the ill-gotten proceeds
  \end{itemize}
\end{columns}

# Types of corruption (Yuen Yuen Ang 2020)

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/Corruption/ang.pdf")
```
\vspace{0.1cm}
\begin{itemize}
\small
  \item Speed money: Petty bribes paid to low-level bureaucrats to circumvent hurdles or delays
  \item Access money: High-stakes rewards offered by elite capitalists to powerful officials in exchange for exclusive, lucrative privileges
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/Corruption/ang_compare.pdf")
```

# "Causes" of corruption (Treisman 2000)

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/Corruption/corruption_cover.pdf")
```

---

\begin{itemize}
  \item Corruption is \textbf{lower} in
  \vspace{0.1cm}
  \begin{itemize}
    \item countries with common law systems
    \item Britain and its former colonies
    \item countries with a Protestant tradition
    \item democratic countries and those with a freer press and more vigorous civic associations
    \item more economically developed countries, where populations are more educated and literate, and where the normative separation between "public" and "private" is clearer
    \item countries with higher relative salaries in public office
    \item countries where political stability os greater
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/Corruption/uslaner_1.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/Corruption/corruption_results.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/Corruption/uslaner.pdf")
```

# Refresher lecture: Political communication in the digital age

\begin{itemize}
\small
  \item Media freedom and democracy
    \vspace{0.2cm}
    \item Democracy and social media
    \begin{itemize}
      \item Echo chamber and polarization
      \item Loss of "local" politics
      \item Misinformation and disinformation
    \end{itemize}
    \vspace{0.2cm}
    \item Extra: Media censorship and propaganda in authoritarian regimes
    \begin{itemize}
      \item Media censorship (King, Pan, and Roberts 2013; Lorentzen 2014)
      \item Propaganda (Wedeen 1999; Huang 2015)
    \end{itemize}
    \vspace{0.2cm}
    \item Reading exercise: "AI and Democracy"
\end{itemize}

# World Press Freedom Index 2025 by Reporters Without Borders

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/rsf.pdf")
```
\vspace{0.2cm}
\begin{center}
\url{https://rsf.org/en/index}
\end{center}

# How social media drives political polarization through "echo" chambers

\begin{itemize}
  \item \textbf{Ideological sorting}: Platforms provide the infrastructure for politically like-minded individuals to identify and congregate with one another easily
  \vspace{0.2cm}
  \item \textbf{Reinforcement loops}: Users are primarily exposed to content that validates their existing worldviews, strengthening their current beliefs rather than challenging them
  \vspace{0.2cm}
  \item \textbf{Algorithmic filtering}: Ranking algorithms prioritize engagement, often hiding opposing viewpoints to keep users within a "filter bubble" of familiar information
  \vspace{0.2cm}
  \item \textbf{Political isolation}: This environment creates a digital silo that effectively shields citizens from diverse perspectives and healthy democratic dissent
\end{itemize}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/truth.pdf")
```

# "Is Online Political Communication More Than an Echo Chamber?" (Barberá et al 2015)

\begin{itemize}
  \item \textbf{Context-Dependent Echo Chambers}: Communication resembles an "echo chamber" primarily during political events (e.g., elections, government shutdowns), whereas nonpolitical events (e.g., the Super Bowl) foster a more unified "national conversation"
  \vspace{0.2cm}
  \item \textbf{Dynamic Polarization}: Some topics, such as the Newtown shootings, can shift over time, beginning as a broad national discussion before transitioning into a polarized exchange
  \vspace{0.2cm}
  \item \textbf{Ideological Asymmetry}: Liberals are more likely than conservatives to engage in cross-ideological information sharing, a trend the authors link to underlying psychological and motivational differences
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/left_authoritarianism.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
\url{https://www.bps.org.uk/research-digest/left-wing-authoritarianism-real-and-needs-be-taken-seriously-political-psychology}
\end{center}

# Decline of local news (and why is it problematic)


```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/local_uk.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/local_usa.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/local_usa2.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/local_usa2_results.pdf")
```

# Misinformation versus disinformation

\begin{itemize}
  \item Misinformation is fake news that is created and spread \textbf{by mistake}, often by someone who does not realise that it it false
  \vspace{0.4cm}
  \item Disinformation is fake news that is created and spread \textbf{deliberately} by someone who knows full well that it is false.
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/meta.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
\url{https://www.bbc.co.uk/news/articles/cly74mpy8klo}
\end{center}

# Censorshop in authoritarian regimes

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/censor_cover.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
\url{doi:10.1017/S0003055413000014}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/censor_results.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/peter_1.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/wedeen.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/huang.pdf")
```

# Reading exercise: "AI and democracy"

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/guardian.pdf")
```
\vspace{0.2cm}
\begin{center}
\tiny
\url{https://www.theguardian.com/technology/2026/jan/22/experts-warn-of-threat-to-democracy-by-ai-bot-swarms-infesting-social-media}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/science.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
\url{https://www.science.org/doi/10.1126/science.adz1697}
\end{center}