---
title: |
  |
  |
  | Investigating the Social World
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Lecture 3: Finding Rigour}
\end{center}
\vspace{0.7cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/cover.pdf")
```

# Wrapping up loose ends

\begin{itemize}
\small
	\item Varieties of "theory" in social research: They contrast each other but also (logically and intellectually) related to each other
	\vspace{0.1cm}
	\begin{itemize}
    \item Theory as paradigm and approach (Geddes 2003)
	  \item Theory for interpretation, explanation and emancipation (Cornelissen et al 2021)
	  \item Theory as propositions, generalization, vocabulary and critique (Martin 2015)
    \item \textbf{Last week}: Theory of ways to seeing, explaining and envisioning
	\end{itemize}
	\vspace{0.2cm}
	\item Key takeaway messages and tasks: "Cultural competence" for social researchers
	\vspace{0.1cm}
	\begin{itemize}
		\item Explain your analytical decisions as a researcher (and yet no need to be apologetic)
		\item Recognize and appreciate the (theoretical) contribution of others
		\item Search for the common ground for fruitful collaborations; avoid dogmatic feud and hunger game
	\end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/crt.pdf")
```
\begin{center}
\tiny
\url{https://www.nytimes.com/article/what-is-critical-race-theory.html}
\end{center}

# CRT as a way of seeing

\vspace{0.2cm}
```{r echo=FALSE, out.width="35%", fig.align="center"}
knitr::include_graphics("Figs/kc_new.pdf")
```
\begin{center}
\scriptsize
Professor Kimberlé Crenshaw (Columbia and UCLA) 
\end{center}
\vspace{0.4cm}
\begin{small}
CRT is is \textbf{a way of seeing}, attending to, accounting for, tracing and analyzing \textbf{the ways} that race is produced, \textbf{the ways} that racial inequality is facilitated, and \textbf{the ways} that our history has created these inequalities that now can be almost effortlessly reproduced unless we attend to the existence of these inequalities.
\end{small}

# CRT as a way of envisioning

\vspace{0.1cm}
```{r echo=FALSE, out.width="35%", fig.align="center"}
knitr::include_graphics("Figs/op_new.pdf")
```
\begin{center}
\scriptsize
Dr OiYan Poon (CAF Co-Lab)
\end{center}
\vspace{0.5cm}
\begin{small}
"'The founders of CRT critiqued liberal ideologies, and that they called on research scholars to seek out and understand the roots of why racial disparities are so persistent, and \textbf{to systematically dismantle racism}.'"
\end{small}

# "The statistical theory of racism and sexism"

```{r echo=FALSE, out.width="27%", fig.align="center"}
knitr::include_graphics("Figs/becker.pdf")
```
\vspace{0.3cm}
\begin{itemize}
\small
\item \textbf{Statistical discrimination}: Based on stereotypes that occur when decision makers use aggregate group characteristics to evaluate performance
\item \textbf{Taste-based discrimination}: Based on prior established prejudice or dislike (and thus triggering avoidance)
\end{itemize}

# Going back to business: Where we are now

\begin{itemize}
  \item Recap: To \textbf{produce valid knowledge claims} (cool, but how?)
  \pause
  \vspace{0.6cm}
  \item Last week: "\textbf{Theory comes to rescue}" (!)
  \vspace{0.6cm}
  \item Coming weeks: "\textbf{Empirics comes to rescue}" (!?)
  \vspace{0.2cm}
  \begin{itemize}
    \item How do social researchers use theory and data in their work? 
    \vspace{0.1cm}
    \item Why should we use numbers to understand the social world?
    \vspace{0.1cm}
    \item How can we put our research in a broader theoretical and social contexts?
  \end{itemize}
\end{itemize}

# What makes your research "rigorous?"

\pause
\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/claude.pdf")
```
\begin{center}
\small
Model in use: Claude 3.5 Sonnet 
\end{center}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/kkv_hd.pdf")
```

---

\begin{itemize}
  \item Why we should pay attention to empirics
  \vspace{1cm}
  \item How theory and empirics can be aligned
  \vspace{1cm}
  \item Epistemology, methodology and theory-empirics alignment (next week)
\end{itemize}

# Why empirics: Trade-off of theory-building

\begin{itemize}
  \item Theory as a way of seeing
  \vspace{0.1cm}
  \begin{itemize}
    \item Pros: Generate ideas and concepts to reveal nuances of the social world
    \item Cons: "Everyone is right" -- a collection of opinions and (analytical) perspectives? 
  \end{itemize}
  \vspace{0.3cm}
  \item Theory as a way of explaining
  \vspace{0.1cm}
  \begin{itemize}
    \item Pros: Lay out a specific and parsimonious process and mechanisms through which the social world "works" 
    \item Cons: "Everyone loses" -- domination and loss of complexities
  \end{itemize}
\end{itemize}
\vspace{0.6cm}
\begin{center}
What do you think? Do you have a preference here?
\end{center}

# Why empirics: The "sandcastle" pitfall

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/castles}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  "'My name is Ozymandias, king of kings:\
  Look on my works, ye Mighty, and despair!'\
  Nothing beside remains. Round the decay'\
  Of that colossal wreck, boundless and bare\
  The lone and level sands stretch far away."\
  $\qquad$ from Percy Bysshe Shelley, "Ozymandias"
  \endcol
\endcols

# How theory aligns with empirics

\begin{itemize}
  \item Induction: Seeking possibilities
  \vspace{0.6cm}
  \item Deduction: Seeking confirmation
  \vspace{0.6cm}
  \item Abduction: Seeking improvement (enrichment)
\end{itemize}

---

\begin{itemize}
  \item \textbf{Induction}: Seeking possibilities
  \vspace{0.1cm}
  \begin{itemize}
    \item Producing broad claims based on our observations showing systematic patterns
    \item Using (non-exhaustive) observations to inform theory-building possibilities
    \item Prone to any unfair or unwarranted rejections with only an (or a few) "exception" or "outlier"
  \end{itemize}
  \vspace{0.6cm}
  \item Deduction: Seeking confirmation
  \vspace{0.6cm}
  \item Abduction: Seeking improvement
\end{itemize}

---

\begin{itemize}
  \item Induction: Seeking possibilities
  \vspace{0.6cm}
  \item \textbf{Deduction}: Seeking confirmation
  \vspace{0.1cm}
  \begin{itemize}   
    \item Evaluating theoretical claims by using data to "test" the hypotheses based from these claims
    \item Ideal if attempting to adjudicate competing theoretical claims; often in line with positivism
    \item Rejecting hypotheses does not necessarily rule out the theory (but it does reduce our faith in the theory, however)
  \end{itemize}
  \vspace{0.6cm}
  \item Abduction: Seeking improvement
\end{itemize}

---

\begin{itemize}
  \item Induction: Seeking possibilities
  \vspace{0.6cm}
  \item Deduction: Seeking confirmation
  \vspace{0.6cm}
  \item \textbf{Abduction}: Seeking improvement
   \vspace{0.1cm}
  \begin{itemize}
    \item Starting by (surprisingly) finding that a hypothesis does NOT work well and may trigger the need for new data and interpretations/explanations
    \item Can be seen as a sequential integration of deductive-inductive processes
    \item Common for the task of theoretical refinement (or improvement)
  \end{itemize}
\end{itemize}

# Examples: Three tales of "culture" and development

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/examples.pdf")
```

---

\vspace{0.5cm}
\begin{itemize}
  \item Research question?
  \vspace{0.6cm}
  \item Theoretical claims or arguments?
  \vspace{0.6cm}
  \item Empirical strategy (and epistemology)?
  \vspace{0.6cm}
  \item Contribution and significance?
\end{itemize}

# Induction: "World Economic Development" (Khan 1979)

\begin{itemize}
  \item \textbf{Research question}: Why did NICs in East Asia achieve high economic growth rates between the 1950s and 1990s?
  \vspace{0.3cm}
  \item \textbf{Theoretical claims or arguments}: These countries were economically successful because of their "Asian values"
  \vspace{0.3cm}
  \item \textbf{Empirical strategy}: (Descriptive) case studies mostly
  \vspace{0.3cm}
  \item \textbf{Significance and contribution}
  \vspace{0.1cm}
  \begin{itemize}
    \item No longer popular (or accepted) due to the lack of empirical support and conceptual setbacks (circular and essentialist)
    \item Critically reflecting on the relationship between culture and development for explanatory purpose
  \end{itemize}
\end{itemize}

# Deduction: "The Oil Curse" (Ross 2012)

\begin{itemize}
  \item \textbf{Research question}: Why are women in the Middle East politically marginalized?
  \vspace{0.2cm}
  \item \textbf{Theoretical claims or arguments}: Petroleum perpetuates patriarchy because oil creates a specific industrial structure that reduces women's participation in the labor market
  \vspace{0.2cm}
  \item \textbf{Empirical strategy}: (Inferential) multiple regression to test the following hypotheses:
  \begin{itemize}
    \item A rise in the value of oil production will reduce female participation in the labor force
    \item A rise in the value of oil production will reduce female political influence
  \end{itemize}
  \vspace{0.2cm}
  \item \textbf{Significance and contribution}: Pushing us to think away from only using Islam as the explanation
\end{itemize}

# Abduction: "Translating the Devil" (Meyer 1999)

\begin{small}
\begin{itemize}
  \item \textbf{Research question}: How have Christian beliefs been accepted by local people in former European colonies in Africa?
  \vspace{0.4cm}
  \item \textbf{Theoretical claims or arguments}: Locals "appropriated" Christian beliefs and transformed Peki Ewe through the (re)imagination of the notion of devil
  \vspace{0.4cm}
  \item \textbf{Empirical strategy}: Using archival research (in Germany and the UK) and ethnography (in Ghana) were employed to document various lived experiences as an outsider from Europe
  \vspace{0.4cm}
  \item \textbf{Significant and contribution}: Enrich Weber's theoretical framework by considering a more complicated, nuanced process of religious conversion with the focus on the agency of local communities
\end{itemize}
\end{small}

# Concluding remarks

\begin{itemize}
  \item Alignment between theory-empirics provide the source of rigor in social research
  \vspace{0.3cm}
  \item Bolster (the credibility of) theory through the establishment theory-empirics alignment and \textit{high-quality} evidence
  \vspace{0.1cm}
  \begin{itemize}
    \item Why is it a good idea to use "numbers" to study the social world (even if you do not know quants)?
    \item Why is causality important? Is causality always a must?
    \item How can we generalize our insights beyond a case?
  \end{itemize}
  \vspace{0.3cm}
  \item Bring epistemology into the discussion: \textbf{How rigid are these epistemology-methodology links?}
  \begin{itemize}
    \item \textbf{Positivism}: Theory as explanation; quantitative
    \item \textbf{Constructivism} Theory as interpretation; qualitative
  \end{itemize}
\end{itemize}

# Discussion: Two worlds of development studies (really?)

\begin{itemize}
  \item Do you agree with the binary distinction between positivism and constructivism?
  \vspace{0.3cm}
  \item Does your research fall into one of these epistemological stances? How do you define your epistemological stance?
  \vspace{0.3cm}
  \item Do you agree with the alignment between epistemological positions and the choice of methods?
  \vspace{0.3cm}
  \item Is such alignment always the case? Can we be somewhere in between? 
  \vspace{0.3cm}
  How do we know when a research work is rigorous?
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/wd.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/epis_cont.pdf")
```

---

\vspace{0.5cm}
\begin{center}
Thank you!
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bbk_tv.pdf")
```
\vspace{0.3cm}
\begin{center}
\texttt{c.cheng@bbk.ac.uk}
\end{center}

