---
title: |
  |
  |
  | Investigating the Social World
  |
  |
author: |
  | Week 3: In Search for Rigour
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\vspace{0.7cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/lego_2.pdf")
```
\vspace{0.5cm}
\begin{center}
Welcome back!
\end{center}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/deepmind.pdf")
```

# Wrapping up the loose ends

\begin{itemize}
	\item Varieties of "theory" in social research: They contrast each other but (logically and intellectually) related to each other
	\vspace{0.1cm}
	\begin{itemize}
	  \item Theory as paradigm and approach (Geddes 2003) 
	  \item Theory for interpretation, explanation and emancipation (Cornelissen et al 2021)
	\end{itemize}
	\vspace{0.3cm}
	\item Key takeaway messages and tasks
	\vspace{0.1cm}
	\begin{itemize}
		\item Explain your analyticial decisions as a researcher (but no need to be apologetic)
		\item Recognize and appreciate the contribution of others
		\item Search for the common ground for fruitful collaborations; avoid dogmatic feud and squid game
	\end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/crt.pdf")
```

# CRT as an interpretative theory

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.5]{Figs/crenshaw}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  CRT is "'is \textbf{a way of seeing}, attending to, accounting for, tracing and analyzing \textbf{the ways} that race is produced," she said, '\textbf{the ways} that racial inequality is facilitated, and \textbf{the ways} that our history has created these inequalities that now can be almost effortlessly reproduced unless we attend to the existence of these inequalities.'"
  \endcol
\endcols

# CRT as an emancipatory theory

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/oiyang}
    \end{figure}
    \vspace{0.2cm}
  \endcol
  \begincol{.5\textwidth}
  "'The founders of CRT critiqued liberal ideologies, and that they called on research scholars to seek out and \textbf{understand the roots of why racial disparities are so persistent, and to systemically dismantle racism}.'"
  \endcol
\endcols

# Explanatory theory of inter-group inequality

\pause
\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/becker}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
    \item \textbf{Statistical discrimination} is based on stereotypes that occur when rational, information-seeking decision makers use aggregate group characteristics to evaluate individuals
  \vspace{0.45cm}
    \item \textbf{Taste-based discrimination} is based on prejudice or racial and gender bias
  \end{itemize}
  \endcol
\endcols

# Breaking the ice: Where we are now

\begin{itemize}
  \item Recap: To \textbf{produce valid knowledge claims} (cool, but how?)
  \pause
  \vspace{0.6cm}
  \item Coming weeks: "\textbf{Empirics comes to rescue}" (?)
  \vspace{0.2cm}
  \begin{itemize}
    \item How do social researchers use theory and data in their work? 
    \vspace{0.1cm}
    \item Why should we use numbers to understand the social world?
    \vspace{0.1cm}
    \item How can we put our research in a broader theoretical and social contexts?
  \end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/kkv_hd.pdf")
```

---

\vspace{0.2cm}
\begin{center}
\textbf{Part II}
\end{center}
\vspace{0.1cm}
```{r echo=FALSE, out.width="55%", fig.align="center"}
knitr::include_graphics("Figs/theme_2.pdf")
```
\vspace{0.1cm}
\begin{center}
\textbf{Building rigor in social research by aligning theory and empirics}
\end{center}

# What is rigor according to AI

\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/theme_2}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{small}
  \begin{itemize}
    \item The image showcases a large, fortress-like structure made of books and \textbf{scientific instruments}, symbolizing the strength and solid foundation of rigorous research.
    \vspace{0.3cm}
    \item This structure is encircled by a moat of flowing \textbf{data and statistics}, highlighting the \textbf{depth and thoroughness} required in research.
  \end{itemize}
  \end{small}
  \endcol
\endcols

# What is rigor according to AI

\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/theme_2}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{small}
  \begin{itemize}
    \item Atop the fortress, researchers are depicted using telescopes, microscopes, and computers, representing their \textbf{meticulous and detailed} work.
    \vspace{0.3cm}
    \item The clear and bright sky above the structure suggests the \textbf{clarity and enlightenment} achieved through rigorous study.
  \end{itemize}
  \end{small}
  \endcol
\endcols

---

\begin{itemize}
  \item Why we should pay attention to empirics
  \vspace{1cm}
  \item How theory and empirics can be aligned
  \vspace{1cm}
  \item Epistemology and theory-empirics alignment
\end{itemize}

# Why empirics: Tradeoff of theory-building

\begin{itemize}
  \item Theory as interpretation
  \vspace{0.1cm}
  \begin{itemize}
    \item Pros: Generate ideas and concepts to reveal nuances of the social world
    \item Cons: "Everyone is right" -- a collection of opinions and perspectives 
  \end{itemize}
  \vspace{0.3cm}
  \item Theory as explanation
  \vspace{0.1cm}
  \begin{itemize}
    \item Pros: Lay out a specific and parsimonious process and mechanisms through which the social world "works" 
    \item Cons: "Everyone loses" -- domination and loss of complexities
  \end{itemize}
\end{itemize}

# Why empirics: The "sandcastle" pitfall

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/castles}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  "'My name is Ozymandias, king of kings: Look on my works, ye Mighty, and despair!' Nothing beside remains. Round the decay of that colossal wreck, boundless and bare The lone and level sands stretch far away."
  \endcol
\endcols

# How theory aligns with empirics

\begin{itemize}
  \item \textbf{Induction}
  \vspace{0.1cm}
  \begin{itemize}
    \item Producing broad claims based on our observations showing systematic patterns
    \item Prone to any unfair or unwarranted rejections with only an (or a few) "exception" or "outlier"
    \item Better for showing the existence a "theoretical" possibility awaiting further research
  \end{itemize}
  \vspace{0.6cm}
  \item Deduction
  \vspace{0.6cm}
  \item Abduction
\end{itemize}

# How theory aligns with empirics

\begin{itemize}
  \item Induction
  \vspace{0.6cm}
  \item \textbf{Deduction}
  \vspace{0.1cm}
  \begin{itemize}   
    \item Evaluating theoretical claims by using data to "test" the hypotheses based from these claims
    \item Ideal if attempting to adjudicate competing theoretical claims; often in line with positivism
    \item Rejecting hypotheses does not necessarily rule out the theory (but it does reduce our faith in the theory, however)
  \end{itemize}
  \vspace{0.6cm}
  \item Abduction
\end{itemize}

# How theory aligns with empirics

\begin{itemize}
  \item Induction
  \vspace{0.6cm}
  \item Deduction
  \vspace{0.6cm}
  \item \textbf{Abduction}
   \vspace{0.1cm}
  \begin{itemize}
    \item Starting by (surprisingly) finding that a hypothesis does NOT work well and may trigger the need for new data and interpretations/explanations
    \item Can be seen as a sequential integration of deductive-inductive processes
    \item Common for the task of theoretical refinement (or improvement)
  \end{itemize}
\end{itemize}

# Examples from "culture" and development

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/examples.pdf")
```

---

\vspace{0.5cm}
\begin{itemize}
  \item Research question?
  \vspace{0.6cm}
  \item Theoretical claims or arguments?
  \vspace{0.6cm}
  \item Empirical strategy (and epistemology)?
  \vspace{0.6cm}
  \item Contribution and significance?
\end{itemize}

# Induction: "World Economic Development: 1979 and Beyond" (Khan 1979)

\begin{itemize}
  \item \textbf{Research question}: Why did NICs in East Asia achieve high economic growth rates between the 1950s and 1990s?
  \vspace{0.3cm}
  \item \textbf{Theoretical claims or arguments}: These countries were economically successful because of their "Asian values"
  \vspace{0.3cm}
  \item \textbf{Empirical strategy}: (Descriptive) case studies mostly
  \vspace{0.3cm}
  \item \textbf{Significance and contribution}
  \vspace{0.3cm}
  \begin{itemize}
    \item No longer popular (or accepted) due to the lack of empirical support and conceptual setbacks (circular and essentialist)
    \item Critically reflecting on the relationship between culture and development for explanatory purpose
  \end{itemize}
\end{itemize}

# Deduction: "The Oil Curse" (Ross 2012)

\begin{itemize}
  \item \textbf{Research question}: Why are women in the Middle East politically marginalized?
  \vspace{0.2cm}
  \item \textbf{Theoretical claims or arguments}: Petroleum perpetuates patriarchy because oil creates a specific industrial structure that reduces women's participation in the labor market
  \vspace{0.2cm}
  \item \textbf{Empirical strategy}: (Inferential) multiple regression to test the following hypotheses:
  \begin{itemize}
    \item A rise in the value of oil production will reduce female participation in the labor force
    \item A rise in the value of oil production will reduce female political influence
  \end{itemize}
  \vspace{0.2cm}
  \item \textbf{Significance and contribution}: Pushing us to think away from only using Islam as the explanation
\end{itemize}

# Abduction: "Translating the Devil" (Meyer 1999)

\begin{small}
\begin{itemize}
  \item \textbf{Research question}: How have Christian beliefs been accepted by local people in former European colonies in Africa?
  \vspace{0.4cm}
  \item \textbf{Theoretical claims or arguments}: Locals "appropriated" Christian beliefs and transformed Peki Ewe through the (re)imagination of the notion of devil
  \vspace{0.4cm}
  \item \textbf{Empirical strategy}: Using archival research (in Germany and the UK) and ethnography (in Ghana) were employed to document various lived experiences as an outsider from Europe
  \vspace{0.4cm}
  \item \textbf{Significant and contribution}: Enrich Weber's theoretical framework by considering a more complicated, nuanced process of religious conversion with the focus on the agency of local communities
\end{itemize}
\end{small}

# Concluding remarks

\begin{itemize}
  \item Alignment between theory-empirics provide the source of rigor in social research
  \vspace{0.3cm}
  \item Bolster the (credibility) of theory through the establishment theory-empirics alignment and \textit{high-quality} evidence
  \vspace{0.1cm}
  \begin{itemize}
    \item Why is it a good idea to use "numbers" to study the social world (even if you do not know quants)?
    \item Why is causality important? Is causality always a must?
    \item How can we generalize our insights beyond a case?
  \end{itemize}
  \vspace{0.3cm}
  \item Bring epistemology into the discussion: \textbf{How rigid are these epistemology-methodology links?}
  \begin{itemize}
    \item \textbf{Positivism}: Theory as explanation; quantitative
    \item \textbf{Constructivism} Theory as interpretation; qualitative
  \end{itemize}
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/epis_bin.pdf")
```

# Discussion: Two worlds of development studies (really?)

\begin{itemize}
  \item Do you agree with the binary distinction between positivism and constructivism?
  \vspace{0.3cm}
  \item Does your research fall into one of these epistemological stances? How do you define your epistemological stance?
  \vspace{0.3cm}
  \item Do you agree with the alignment between epistemological positions and the choice of methods?
  \vspace{0.3cm}
  \item Is such alignment always the case? Can we be somewhere in between? 
  \vspace{0.3cm}
  How do we know when a research work is rigorous?
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/epis_cont.pdf")
```

---

\vspace{0.5cm}
\begin{center}
Thank you!
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bbk_tv.pdf")
```
\vspace{0.3cm}
\begin{center}
\texttt{c.cheng@bbk.ac.uk}
\end{center}
