---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 4: Ruling}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week4.pdf")
```

# Today's plan

\begin{itemize}
  \item Logistics: Attendance and assessment 
  \vspace{0.3cm}
  \item Recap from previous weeks
  \begin{itemize}
    \item Nationalism and democracy (and varieties of democracy)
    \item Reading exercise: "Constitutional Design and Quality of Democracy"
  \end{itemize}
  \vspace{0.3cm}
  \item Refresher lecture: Democracy versus autocracy
  \vspace{0.1cm}
  \begin{itemize}
    \item Political development and regime change
    \item Comparative authoritarianism
  \end{itemize}
\end{itemize}

# Reading exercise summary: Nationalism and democracy

\begin{itemize}
  \item Nationalism is one of the most important political ideologies in contemporary era as it is tied to the creation of nation-states
  \vspace{0.2cm}
  \item Nationalism may or may not go along with democracy, as with a strong sense of shared identities
  \begin{itemize}
    \item is crucial for a "healthy" democracy with strong civic culture; see Robert Putnam's \textit{Making Democracy Work} (1993)
    \item can pave the foundation for far-right movements against minorities (e.g., India under BJP) and political repression in dicatatorships (e.g., North Korea)
  \end{itemize}
  \vspace{0.2cm}
  \item Nationalism could be categorized by (1) the narrative content (inclusive versus exclusive) and (2) the degree to which the narrative is open to contestation (open versus close)
  \vspace{0.2cm}
  \item Further questions to reflect on "British nationalism" 
\end{itemize}

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/reading_con.pdf")
```

# Recap: Varieties of democratic constitutions

\begin{itemize}
\small
  \item Varieties of constitutional democracy
  \vspace{0.1cm}
  \begin{itemize}
    \item Executive system: presidential versus parliamentary
    \item Electoral system: majoritarian/plurality, PR and mixed-member (and the resulting party system)
    \item Legislative system: uni- or bi-cameral
    \item Central-local relations: federal, unitary and decentralized
  \end{itemize}
  \vspace{0.2cm}
  \item Two visions of democracy: Majoritarian Model versus Consensus Model
  \vspace{0.1cm}
  \begin{itemize}
    \item Power concentration and dispersion (Lijphart 1999; Powell 2000), and the resulting design of judiciary
    \item Rigid or adaptive constitutions (Levitsky and Murillo 2009)
  \end{itemize}
  \item Democratic constitutional design is a response to "Madison's Dilemma" -- creating a government strong enough to govern while limited enough to protect civil liberties
\end{itemize}

# Democracy and economic development

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/dev_book.pdf")
```
\vspace{0.2cm}
\begin{itemize}
  \item Rule of law, constraints on executive power and private property rights
  \item Electoral accountability and government responsiveness
\end{itemize}

# Refresher lecture

\begin{itemize}
  \item Political development and regime change
  \begin{itemize}
    \item By driving force
    \item By starting point
  \end{itemize}
  \vspace{0.3cm}
  \item Comparative authoritarianism
  \begin{itemize}
    \item Definition, terminology and typology
    \item Tactics of authoritarian governance: Repression, co-optation and information control
  \end{itemize}
  \vspace{0.3cm}
  \item Regime type and development
\end{itemize}

# Regime change and political development

\begin{itemize}
\small
  \item "Modernization" as the key driving force
  \begin{itemize}
    \item Economic modernization (?)
    \item Cultural modernization (?) 
  \end{itemize}
  \vspace{0.3cm}
  \item Regime change can start from anywhere
  \begin{itemize}
    \item Internal: Top-down (elites) and/or bottom-up (civil society)
    \item External: Foreign powers and/or diaspora
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/adamp.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/error_cover.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/error.pdf")
```

--- 

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/berlin.pdf")
```
\vspace{0.1cm}
\begin{center}
Monday demonstrations in East Germany (1989-1991)
\end{center}

--- 

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/egypt.pdf")
```
\vspace{0.1cm}
\begin{center}
Arab Spring and the fall of Mubarak in Egypt (2011)
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/backsliding.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/weimar.pdf")
```

# Further questions to think about

\begin{itemize}
  \item By the underlying driving force and starting point, how many combinations are possible here? 
  \vspace{0.3cm}
  \item Do regime changes always lead to democratization?  Democratic transition versus consolidation?
  \vspace{0.3cm}
  \item Recent trend of "autocratization" and "democratic backsliding" -- why?
\end{itemize}

# Comparative authoritarianism: Definition and terminology

\begin{itemize}
\small
  \item Cheibub et al (2010): For a regime to be democratic, both the \textbf{chief executive office} and the \textbf{legislative body} must be filled by regular and competitive elections (ex ante uncertainty, ex post irreversibility and repeatability)
  \vspace{0.2cm}
  \item Geddes et al (2015)
  \vspace{0.1cm}
  \begin{itemize}
    \item An executive achieved power through undemocratic means; \textbf{"undemocratic" refers to any means besides direct, reasonably fair, competitive elections} 
    \item The government achieved power through democratic means, but subsequently changed the formal or informal rules
    \item The military prevented parties that substantial numbers of citizens who would be expected to vote for from competing and/or dictated important policy choices
  \end{itemize}
  \vspace{0.2cm}
  \item Terminology: Dictatorships and autocracies; authoritarianism and totalitarianism
\end{itemize}

# Varieties of authoritarianism (typology versus dichotomy)

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/GWF/type.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/GWF/survival.pdf")
```
\vspace{0.2cm}
\begin{itemize}
  \item Military regimes are most prone to collapse
  \item Monarchies and single-party regimes are the most resilient
\end{itemize}

# Governing authoritarian regimes

\vspace{0.3cm}
\begin{columns}
  \begin{column}{0.3\textwidth}
  \begin{figure}
  \includegraphics[width=1\linewidth]{Figs/handbook}\\
  \end{figure}
  \end{column}
  \begin{column}{0.5\textwidth}
    \begin{itemize}
    \item Physical repression and coercion
    \vspace{0.3cm}
    \item Information control and censorship
    \vspace{0.3cm}
    \item Co-optation and mobilization through institutions and patronage
    \end{itemize}
  \end{column}
\end{columns}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/toolkit.pdf")
```

---

\begin{itemize}
  \item Against individuals versus general population
  \vspace{0.3cm}
  \item Logistics of political repression
  \begin{itemize}
    \item Information dilemma
    \item Principal-agent delegation
    \item Loyalty versus coercion
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/lee2.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
Lee Hsien Loong: "No one should be denied medical care because they cannot afford it and that is my commitment to you. I introduced the \textbf{Pioneer Generation Package} (PGP) in 2014. Now, the Government will be rolling out the `Merdeka Package' to those in their 60s, who had just missed out on qualifying for the PGP."
\end{center}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/voting.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/dprk.pdf")
```
\vspace{0.1cm}
\begin{center}
General elections in North Korea (2014)
\end{center}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/elections.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/electoral.pdf")
```
