---
title: |
  |
  |
  | Social Network Analysis: A Primer
  |
  |
author: |
  | Masterclass in Social Research
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Plan for the day

\begin{itemize}
  \item Why social network analysis
  \vspace{0.3cm}
  \item Investigating the social world with networks
  \vspace{0.3cm}
  \item How do researchers study social networks
  \vspace{0.3cm}
  \item Looking ahead and additional resources
  \vspace{0.3cm}
  \item Exercise: Find a network for your research interest/agenda
\end{itemize}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/boy.pdf")
```
\vspace{0.4cm}
\textit{"No man is an island entire of itself; every man is a piece of the continent, a part of the main."}

# Why social networks?

\begin{itemize}
  \item The social world is complex and involves a variety of interactive dynamics among individuals and organizations that together form the social "structure"
  \vspace{0.3cm}
  \item Social network analysis (SNA) is both an approach and a method, and has received enormous scholarly attention
  \vspace{0.1cm}
  \begin{itemize}
    \item As an approach, SNA encourages us to adopt a "relational" approach to study (and perhaps theorize) the world 
    \item As a method, SNA requires us to move beyond the traditional disciplinary barriers so as to contextualize, describe/measure, analyze and explain/interpret how different relations shape the outcome of interest systematically
  \end{itemize}
  \vspace{0.3cm}
  \item Ultimately, SNA is part of the larger field of \textbf{network science} and aims to uncover the "principles" of "complex" social system (e.g., chaos theory)
\end{itemize}

# Investigating the world with networks

\begin{itemize}
  \item Network for "more-than-human" worlds
  \vspace{0.5cm}
  \item Social networks in politics
  \vspace{0.1cm}
  \begin{itemize}
    \item Social capital
    \item Collective action
    \item Party ID and vote choices
  \end{itemize}
  \vspace{0.5cm}
  \item Social networks and organizational behaviors
  \vspace{0.1cm}
  \begin{itemize}
    \item Structural holes
    \item Weak ties
    \item Community structure
  \end{itemize}
\end{itemize}

# Energy supply Network in the US

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/power.pdf")
```

# Social capital and democratic performance

\vspace{0.3cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/putnam.pdf")
```

# Repression and social movements

\vspace{0.3cm}
```{r echo=FALSE, out.width="110%", fig.align="center"}
knitr::include_graphics("Figs/siegel11.pdf")
```
\vspace{0.2cm}
\begin{scriptsize}
Siegel, David A. 2011. "When Does Repression Work? Collective Action in Social Networks." \textit{Journal of Politics 73}(4): 993-1010.
\end{scriptsize}

# Military coup and regime transition

\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/haiti.pdf")
```
\vspace{0.2cm}
\begin{scriptsize}
Naidu, Suresh, James A. Robinson, and Lauren E. Young. 2021. "Social Origins of Dictatorships: Elite Networks and Political Transitions in Haiti." \textit{American Political Science Review 115}(3): 900-916.
\end{scriptsize}

# Partisan preferences and vote choices in democracies

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/elec_usa.pdf")
```
\vspace{0.2cm}
\begin{scriptsize}
Leiter, Debra, Andreas Murr, Ericka Rascón Ramírez, and Mary Stegmaier. 2018. "Social Networks and Citizen Election Forecasting: The More Friends the Better." \textit{International Journal of Forecasting 34}(2): 235-248.
\end{scriptsize}

# How do researchers study social networks

\begin{itemize}
    \item A \textbf{network} (\textbf{graph}) consists of a set of \textbf{nodes} (or \textbf{vertices}) with certain pairs of these nodes connected by \textbf{edges} (i.e., \textbf{link}).
    \vspace{0.3cm}
    \begin{itemize}
      \item A node can be an individual or an organization or a country.
      \item An edge can be weighted (e.g., trade) and/or directed (e.g., friendship) to represent information exchanges, emotional connections, and financial transactions
    \end{itemize}
    \vspace{1cm}
    \item Two nodes are \textbf{neighbors} if they are connected by an edge.
\end{itemize}

# Example: Inter-marriages between big families in Renaissance Florentine (Padgett and Ansell 1993)

\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/marriage.pdf")
```
\vspace{0.5cm}
\begin{scriptsize}
Padgett, John F., and Christopher K. Ansell. 1993. "Robust Action and the Rise of the Medici, 1400-1434." \textit{American Journal of Sociology 98}(6): 1259-1319.
\end{scriptsize}

# How do researchers study social networks

\begin{itemize}
  \item \textbf{Descriptive} SNA: Various qualitative and quantitative summary measures of a network's connectivity
  \vspace{0.1cm}
  \begin{itemize}
    \item Nodes and edges in a network
    \item Structural features of a network (e.g., community detection)
  \end{itemize}
  \vspace{1cm}
  \item \textbf{Inferential} (generative) SNA, such as quantitative analysis on the underlying dynamic process of network formation and computer-assisted data analysis and simulation (e.g., ERGM)
\end{itemize}

# Nodes and edges in a network

\begin{itemize}
  \item Degree: The number of edges for a node (i.e., how connected a node is in a network)
  \vspace{0.3cm}
  \item Density: The number of edges out of the max possible of edges of a graph
  \vspace{0.3cm}
  \item Centrality: The importance (?) of a node in a network
  \vspace{0.1cm}
  \begin{itemize}
    \item Degree centrality: A node is important if it connects to many others in the network
    \item Eigenvector centrality: A node is important if it connects to other nodes that connect to many other nodes in the network
    \item Closeness centrality: A node is important if it is on average close to many other nodes in the network
    \item Betweenness centrality: A node is important if it is on the paths between many other nodes in the graph
  \end{itemize}
\end{itemize}

# Taking the structure seriously: Community fragmentation and local governance in Global South

\vspace{0.3cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/cruzapsr.pdf")
```
\vspace{0.2cm}
\begin{scriptsize}
Cruz, Cesi, Julien Labbone, and Pablo Querubin. 2020. "Social Network Structures and the Politics of Public Goods Provision: Evidence from the Philippines." \textit{American Political Science Review 114}(2): 486-501.
\end{scriptsize}

# Taking the structure seriously: Weak ties between two densely connected networks

\vspace{0.3cm}
```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/weak.pdf")
```
\vspace{0.2cm}
\begin{scriptsize}
Granovetter, Mark S. 1973. "The Strength of Weak Ties." \textit{American Journal of Sociology 78}(6): 1360-1380.
\end{scriptsize}

# Taking the structure seriously: Structural holes and brokers between two densely connected networks

\vspace{0.3cm}
```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/burt.pdf")
```
\vspace{0.2cm}
\begin{scriptsize}
Burt, Ronald S. 2004. "Structural Holes and Good Ideas." \textit{American Journal of Sociology 110}(2): 349–399.
\end{scriptsize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pic_gran.pdf")
```
\vspace{0.5cm}
\begin{center}
\begin{scriptsize}
From the \textbf{Reliants Project}: \url{https://www.reliantsproject.com/}.
\end{scriptsize}
\end{center}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pic_burt.pdf")
```
\vspace{0.5cm}
\begin{center}
\begin{scriptsize}
From the \textbf{Reliants Project}: \url{https://www.reliantsproject.com/}.
\end{scriptsize}
\end{center}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="93%", fig.align="center"}
knitr::include_graphics("Figs/books_pop.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="93%", fig.align="center"}
knitr::include_graphics("Figs/books_comp.pdf")
```

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/sfi.pdf")
```
\vspace{0.5cm}
\begin{small}
"\textbf{The Santa Fe Institute} is an independent, nonprofit theoretical research institute located in \textbf{Santa Fe, New Mexico, United States} and dedicated to the \textbf{multidisciplinary study of the fundamental principles of complex adaptive systems}, including \textbf{physical}, \textbf{computational}, \textbf{biological} (or ecological), and \textbf{social} systems."
\end{small}

# Looking ahead

\begin{itemize}
  \item Social network analysis encourages us to study the social world and human behaviors through the lens of "relations"
  \vspace{0.5cm}
  \begin{itemize}
    \item The nodes or vertices of interest can be human (e.g., voters) and/or non-human (e.g., firms and countries)
     \vspace{0.3cm}
    \item The ties or edges of interest need to be defined and thought through carefully
  \end{itemize}
\end{itemize}

# Looking ahead

\begin{itemize}
  \item Social network analysis (or network science) demands expertise from a variety theoretical knowledge and empirical approaches
  \vspace{0.3cm}
  \begin{itemize}
    \item Theoretical knowledge: Mathematics, physics, engineering, anthropology, sociology, economics, political science and more (the list can go on)
    \vspace{0.2cm}
    \item Empirical approaches: Qualitative (e.g., archives, ethnography and interviews), quantitative (e.g., surveys and descriptive/inferential SNA), and computational (e.g., agent-based modeling and machine learning)
  \end{itemize}
  \vspace{0.5cm}
  \item Choose any combinations that suit your research agenda and analytical goals; find the right collaborators.
\end{itemize}

# Looking ahead

\begin{itemize}
  \item Social network analysis remains challenging
  \vspace{0.5cm}
  \begin{itemize}
    \item Training across and collaboration among different subject areas can be difficult and rare
    \vspace{0.3cm}
    \item Theory-building and data analysis (measurement/operationalization`, collection, and analysis) can be daunting
    \vspace{0.3cm}
    \item Understanding (mechanisms) and interpreting (meanings) findings produced by (inductive and descriptive) data-driven research may contested (e.g., what it means to say a node is important)
  \end{itemize}
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/island.pdf")
```
\vspace{0.5cm}
\begin{center}
Thank you!
\end{center}

# Exercise: Find a network you can study

\begin{itemize}
  \item Reflect on your area(s) of substantive interest
  \vspace{0.3cm}
  \item Pick a particular structure, phenomenon and/or pattern of relations and describe how it might be represented using networks
  \vspace{0.1cm}
  \begin{itemize}
    \item What would the nodes represent? Are the nodes of interest human or non-human?
    \item What would the edges represent? Are the edges weighted?  Are the edges directed?
  \end{itemize}
  \vspace{0.3cm}
  \item Reflect on the advantages and disadvantages of your chosen representation -- what important features do you capture by using networks?
  \vspace{0.3cm}
  \item How do you collect the data? Any concerns with respect to logistics and/or ethics?
\end{itemize}
