---
title: |
  |
  |
  | Investigating the Social World
  |
  |
author: |
  | Week 5: Comparative and International Analysis
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\vspace{0.7cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/lego.pdf")
```
\vspace{0.5cm}
\begin{center}
Welcome back!
\end{center}

---

\vspace{0.2cm}
\begin{center}
\textbf{Part IV}
\end{center}
\vspace{0.4cm}
```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/theme.pdf")
```
\vspace{0.3cm}
\begin{center}
\textbf{Comparative and international analysis}
\end{center}

---

```{r echo=FALSE, out.width="40%", fig.align="center"}
knitr::include_graphics("Figs/theme.pdf")
```
\vspace{0.3cm}
\begin{center}
\small
"The image visually divides into two halves, each showcasing the unique characteristics of different cultures or countries, connected by a bridge of books, reports, and digital data streams. This bridge symbolizes the exchange of knowledge and research methods across nations. The central area, where researchers are seen collaborating, highlights the integrative nature of this field, emphasizing teamwork and the sharing of insights across borders."
\end{center}

---

\vspace{0.4cm}
\begin{itemize}
  \item Why comparative and intl analysis
  \vspace{0.7cm}
  \item Move from case studies to social science (Bates 2007; Pepinsky 2019)
  \vspace{0.7cm}
  \item Practical guides for generalization
  \vspace{0.7cm}
  \item Conclusion: Investigating the social world as a vocation
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/rosenbluth.pdf")
```

# Why comparative and intl analysis matters

\begin{itemize}
  \item Prevent selection bias (Geddes 1990; Collier and Mahoney 1996)
  \vspace{1cm}
  \item Avoid the pitfalls of "essentialism" (Fuchs 2001; Phillips 2010)
\end{itemize}

# The cases you choose affect the answers you get

\begin{itemize}
  \item The cases you choose can be partial and prevents useful generalization
  \vspace{0.1cm}
  \begin{itemize}
    \item The cases you choose may only represent one or a small number of similar cases in the domain/universe
    \item Without identifying the domain/universe/population, it is very difficult to ascertain the nature of your cases (typical/representation OR deviant/outlier/niche) and their analytical value (Gerring 2006)
  \end{itemize}
  \vspace{0.2cm}
  \item Selecting cases on dependent variable is very likely to produce faulty inference as we jump into the conclusion that
  \vspace{0.1cm}
  \begin{itemize}
    \item any characteristic shared by the selected cases is the cause
    \item the observed pattern can be extended to cases not included in the study
  \end{itemize}
\end{itemize}

# Labor repression and economic growth, 1970-81 (Geddes 2003)

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/sample.pdf")
```
\begin{center}
\small
When we only consider East Asian countries, labor repression seems to induce economic growth
\end{center}

# Labor repression and economic growth, 1970-81 (Geddes 2003)

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/all.pdf")
```
\begin{center}
\small
When we consider all countries, however, there exists no clear statistical association
\end{center}

# Be aware of the "pitfalls" of essentialism (Phillips 2010)

\begin{itemize}
  \item Essentialism can take at least the following four different forms in social research (Phillips 2010)
  \vspace{0.3cm}
  \begin{itemize}
    \item \textbf{Ignoring individual agency and micro-level variation}: Attribution of certain characteristics to all observations subsumed within a particular category
    \vspace{0.1cm}
    \item \textbf{Naturalizing things that are socially and historically constructed}: Attribution to the category by itself
    \vspace{0.1cm}
    \item \textbf{Sustaining stereotype and prejudice}: Invocation of a collectivity, presuming a homogenized and unified group
    \vspace{0.1cm}
    \item \textbf{Impeding innovation by suffocating contestation}: Sanctioning deviations from the presumed collective category uncritically
  \end{itemize}
  \pause
  \vspace{0.4cm}
  \item \textbf{Discussion}: Can you think of any examples in daily life or research you have read? Is attribution to categories always a bad idea? How can we get around these potential pitfalls?
\end{itemize}

# Scope condition is key for comparative and intl analysis

```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/tsai_mit.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
"The specification and empirical assessment of \textbf{scope conditions} are essential to building comparative political theory—theory that not only posits a causal relationship but precisely identifies -- and empirically validates -- \textbf{the conditions or contexts under which the theory holds} (\textit{Comparative Political Studies}, 2016)
\end{center}

# From case studies to social science: Three forms/stages of "comprehension"

\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/bates}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
    \item Apprehension
    \vspace{1cm}
    \item Explanation
    \vspace{1cm}
    \item Conviction
  \end{itemize}
  \endcol
\endcols

# From case studies to social science: Three forms/stages of "comprehension"

\begin{itemize}
  \item Apprehension
  \vspace{0.1cm}
  \begin{itemize}
    \item "Immersion" (often via ethnography and participant observation) as the key foundation of learning and understanding to inform judgment and sharpen intuition
    \item The experiences in the field often provides a "bullshit meter" (or more commonly as we put, "plausibility check")
    \item The first step "comes in the field itself" before "the researcher begins to separate from the field" and moves towards "explanation"
  \end{itemize}
  \vspace{0.5cm}
  \item Explanation
  \vspace{0.5cm}
  \item Conviction
\end{itemize}

# From case studies to social science: Three forms/stages of "comprehension"

\begin{itemize}
  \item Apprehension
  \vspace{0.5cm}
  \item Explanation
  \vspace{0.1cm}
  \begin{itemize}
    \item The movement towards "explanation" involves the "recoding" of the observations and encounters in the field
    \item Explanation establishes \textbf{the} lines of logic to link causes and consequences; the \textbf{thick} description provides the underpinnings for the exercise of abstraction and theory-building
    \item While \textbf{interpretation} aims to provide "a" meaning of the descriptions, \textbf{explanation} seeks to provide a logical link that render the outcome of interest necessary or inevitable
  \end{itemize}
  \vspace{0.5cm}
  \item Conviction
\end{itemize}

# From case studies to social science: Three forms/stages of "comprehension"

\begin{itemize}
  \item Apprehension
  \vspace{0.5cm}
  \item Explanation
  \vspace{0.5cm}
  \item Conviction
  \vspace{0.1cm}
  \begin{itemize}
    \item A stage to accord credence or find \textbf{confirmation} to our account
    \item Seeking "conviction" so as to yield outcomes consistent with the data and, in particular, \textbf{data other than that from which it was first derived}
    \item Moving out the sample to large-N analysis as the final road to conviction
  \end{itemize}
\end{itemize}

# Economic behaviors of coffee growers in East Africa and Colombia (Bates 2007)

\begin{itemize}
  \item Use the field to search for potentially testable and generalizable insight (through ethnography)
  \vspace{0.1cm}
  \begin{itemize}
    \item The behavior of the coffee growers represented a choice by rational actor under varying institutional environment
    \item Be cautious about the temptation of using culture to explain the consequences
  \end{itemize}
  \vspace{0.2cm}
  \item Scope through the initial cases for plausibility check and theory building (through archival research and interviews)
  \vspace{0.2cm}
  \item Pave the road to conviction with small-N and large-N evidence
  \vspace{0.1cm}
  \begin{itemize}
    \item Test the proposed logic in comparable new cases in Latin America, especially Colombia
    \item Test the proposed logic with large-N (muiltivariate) regression (all African countries between 1970-95)
  \end{itemize}
\end{itemize}

# Solidary groups and accountable local public goods provision in rural China (Tsai 2007)

\begin{itemize}
  \item Two months of \textbf{preliminary research} in seven provinces, followed by
  \vspace{0.1cm}
  \begin{itemize}
    \item eight months of \textbf{in-depth fieldwork} in Fujian
    \item detailed \textbf{case studies} on selected provinces, which reflect different levels of economic development, as well as regional differences in terrain, institutional history and social organization between north and South China
  \end{itemize}
  \vspace{0.4cm}
  \item Designed \textbf{surveys} to cover 300$+$ villages, carefully sampled from the selected provinces with assistance from local official partners and/or research assistants (mostly students)
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/singh_new.pdf")
```
\vspace{0.3cm}
\pause
\begin{center}
Can Singh's theoretical insights to other cases? What cases? How do we make sure the new cases are comparable to Singh's choice in the article?
\end{center}

# Exploring and extending Singh's scope condition (by a former ISW student)

\begin{itemize}
\small
  \item "It could be argued the \textbf{destabilisation of racialised identities within an ethnically diverse locality such as the gym may bear relevance in similar convivial sporting localities, in super-diverse, Global North cities}, with similar political conditions to London."
  \pause
  \vspace{0.2cm}
  \item "An alternative case to increase the scope condition of similar research to other contexts might be supplementary use of \textbf{quantitative methods, such as questionnaires measuring gym members' racial or gender identity salience in the gym compared to other locations, and statistical analysis of differences to serve as a replicable tool in different contexts}."
  \pause
  \vspace{0.2cm}
  \item "However, \textbf{this would signify an entirely different study}. Generalisations about this article's findings would require de-contextualisation, which is antithetical to the context-specific, complex qualitative methodological design and constructivist approach of Singh's research."
\end{itemize}

# Some practical suggestions for generalization

\begin{itemize}
\small
  \item A researcher can be driven by the general question rather than the case
  \vspace{0.2cm}
  \item If possible, think outside the box by formulating the research questions within and outside the given case (i.e., de-contextualization)
  \vspace{0.2cm}
  \item Define the \textbf{scope condition} carefully and explicitly -- that is, what are the comparable cases out there and why? Where does the case of interest come from?
  \vspace{0.2cm}
  \item Be careful with any cultural explanations as they may create some issues (e.g., reverse causality; circularity and tautology; loss of agency)
  \vspace{0.2cm}
  \item Do not force generalization when it is not the main objective of your research; be transparent and do not mislead yourself and your audiences
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/swedberg.pdf")
```
\vspace{0.3cm}
\begin{center}
\small
"Abstractions are used in all the sciences, and the capacity to make abstractions is a very useful tool when you theorize. It is also often helpful to go between different levels of abstraction, such as high, medium, and low. An abstraction is created in three steps: \textbf{You single out some phenomenon of interest}; \textbf{you cut it out and isolate it from its surroundings}; and \textbf{you remove various parts from it, so it will come out more clearly}."
\end{center}

# Conclusion: Investigating the social world as a vocation

\begin{itemize}
\small
  \item ISW is a special module, reflecting on the production of "valid" knowledge about the social world
  \begin{itemize}
    \item Know what you do and how to explain the objectives of your scholarly inquiry;
    \item Know how to tell the differences between different objectives and engage other researchers' work constructively and effectively
  \end{itemize}
  \vspace{0.1cm}
  \item Social research involves a series of careful thinking about epistemology, theory-building, and methodology (hence theory-empirics alignment)
  \begin{itemize}
    \item \textbf{Reveal nuances of the real world}: constructivism; theory as approach; interpretation; qualitative; inductive/abductive
    \item \textbf{Search for the "ultimate" truth}: positivism; theory as paradigm; explanation; quantitative; deductive
  \end{itemize}
  \vspace{0.1cm}
  \item Exceptions always exist and boundaries are constantly being contested and redefined
\end{itemize}

# Discussion: "Solidary Groups, Informal Accountability, and Local Public Goods Provision" (Tsai 2007)

\begin{itemize}
  \item What is the main research question? What is the scope condition?
  \vspace{0.7cm}
  \item Can you summarize her main theoretical argument? Can/should culture serve as the alternative explanation?
  \vspace{0.7cm}
  \item Are the findings significant and/or valid? Can you come up with another research project to extend this line of inquiry?
\end{itemize}

---

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \vspace{0.4cm}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/laitin}
    \end{figure}
    \vspace{0.2cm}
    \begin{center}
    \textbf{David D. Laitin} (Stanford)
    \end{center}
  \endcol
  \begincol{.5\textwidth}
  \begin{small}
  \begin{itemize}
    \item \textit{Why Muslim Integration Fails in \textbf{Christian-Heritage Societies}} (Cambridge: Harvard University Press, 2016) with Claire Adida and Marie-Anne Valfort
    \vspace{0.2cm}
    \item \textit{Identity in Formation: the \textbf{Russian-speaking} Populations in the Near Abroad} (Oxford: Oxford University Press, 2007)
    \vspace{0.2cm}
    \item \textit{Nations, States and Violence} (Ithaca: Cornell University Press, 1998)
    \vspace{0.2cm}
    \item \textit{Language Repertoires and State Construction in \textbf{Africa}} (Cambridge: Cambridge University Press, 1992)
  \end{itemize}
  \end{small}
  \endcol
\endcols

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/lachlan.pdf")
```

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/lachlan2.pdf")
```

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/scope_condition.pdf")
```
\vspace{0.4cm}
\begin{center}
\url{https://www.scopeconditionspodcast.com/}
\end{center}