---
title: |
  |
  |
  | Social Research Dissertation Workshop
  |
  | Part II: Navigating the Process  
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

## Today's workshop is for

\begin{itemize}
  \item MRes Global Politics/Politics/Public Policy and Management
  \vspace{1cm}
  \item MRes/MSc Social Research
\end{itemize}

## Plan for today

\begin{itemize}
  \item Overview of the dissertation process (again)
  \vspace{0.3cm}
  \item Action points before the deadline (\textbf{15 September 2023})
  \vspace{0.3cm}
  \item Available support and tips
  \vspace{0.3cm}
  \item Discussion and Q\&A
\end{itemize}

## Dissertation procress: An overview

\begin{itemize}
	\item Term 1 (2022): Prepare your \textbf{dissertation proposal form}; talk to the program director and/or your potential supervisor (if there is one)
	\vspace{0.2cm}
	\item Terms 2 and 3 (2023)
		\begin{itemize}
			\item January: Submit your proposal form by \textbf{20 January 2023}
			\item February to early March: Supervisor allocation; the Department sends a confirmation email about your dissertation supervisor
			\item March to June/early July: Supervision in progress; submit an ethics proposal (if necessary)
			\item July-August: Supervision ends (typically); complete your dissertation
			\item September: Submit your dissertation by \textbf{15 September 2023}
		\end{itemize}
	\vspace{0.2cm}
	\item After September 2023: Extension and deferral
\end{itemize}

## Action points once you have a supervisor

\begin{itemize}
  \item Meet and work with your supervisors
  \vspace{0.6cm}
  \item Prepare and submit the ethics proposal (covered by \textbf{Dr Barry Maydom})
  \vspace{0.6cm}
  \item Complete your research and dissertation
  \vspace{0.6cm}
  \item Deferring your dissertation (if necessary)
\end{itemize}

## Action point 1: Meet and work with your supervisor

\begin{itemize}
    \item Min \textbf{three} meetings before the end of Term 3 (Summer)
    \vspace{0.1cm}
    \begin{itemize}
        \item It is your responsibility to arrange meetings and keep in good contact with your supervisors
        \item \textbf{Supervisors can}: Offer guidance and support (but not necessarily direction); advise you on approach, scope, questions to be asked, and the outline structure and research design and methods
        \item \textbf{Supervisors may not}: Comment on drafts and meet you during the summer break
    \end{itemize}
    \vspace{0.3cm}
    \pause
    \item Read through the following documents for more information
    \vspace{0.1cm}
    \begin{itemize}
        \item "Guidelines for MRes and MSc Social Research Dissertations"
        \item "Policy on Supervised Dissertations for Taught Students"
    \end{itemize}
\end{itemize}

## Action point 2: Prepare and submit the ethics proposal

Covered by Dr Barry Maydom.

## Action point 3: Complete your research and dissertation

\begin{itemize}
  \item Working with yourself
  \vspace{1cm}
  \item Working with your peers
  \vspace{1cm}
  \item Additional support activities
\end{itemize}

## Action point 3: Complete your research and dissertation

\begin{itemize}
  \item \textbf{Working with yourself}
  \vspace{0.1cm}
  \begin{itemize}
    \item Pay close attention to personal wellbeing
    \item Manage your time and make weekly plans for your writing
  \end{itemize}
  \vspace{0.7cm}
  \item Working with your peers
  \vspace{0.7cm}
  \item Taking advantage of additional support activities
\end{itemize}

---

\vspace{0.3cm}
\begin{itemize}
    \item Self-care checklist
    \vspace{0.1cm}
    \begin{itemize}
    \item Get good sleep
    \item Nourish and move your body
    \item Do something fun
    \item Give yourself credit
    \item Connect with others (see the next part)
    \end{itemize}
    \vspace{0.8cm}
    \item Visit \texttt{\url{https://www.bbk.ac.uk/student-services/counselling-service}} for more information.
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="84%", fig.align="center"}
knitr::include_graphics("Figs/chicago.pdf")
```
\vspace{0.3cm}
\begin{center}
\begin{scriptsize}
\url{https://press.uchicago.edu/ucp/books/book/chicago/W/bo26985005.html}
\end{scriptsize}
\end{center}

## Action point 3: Complete your research and dissertation

\begin{itemize}
  \item Working with yourself
  \vspace{0.7cm}
  \item \textbf{Working with your peers}
  \vspace{0.1cm}
  \begin{itemize}
    \item Why you need writing groups
    \item Different forms of writing groups
    \item Planning your writing group(s)
  \end{itemize}
  \vspace{0.7cm}
  \item Taking advantage of additional support activities (Term 3)
\end{itemize}

---

\vspace{0.3cm}
\begin{itemize}
    \item Why you need writing groups
    \vspace{0.1cm}
    \begin{itemize}
        \item Accountability: Regular progress checks
        \item Social: Collegial support and fun
    \end{itemize}
    \vspace{0.2cm}
    \item Writing groups can take different forms
    \vspace{0.1cm}
    \begin{itemize}
        \item Accountability groups
        \item Peer review groups
        \item "Write-along" groups
    \end{itemize}
    \vspace{0.2cm}
    \item Writing groups take some careful (but not difficult) planning/moderation: \textbf{Size}, \textbf{frequency}, \textbf{subject areas and/or methods}, and \textbf{format} (online or in-person)
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="84%", fig.align="center"}
knitr::include_graphics("Figs/ucla_group.pdf")
```
\vspace{0.15cm}
\begin{center}
\begin{scriptsize}
\url{https://www.youtube.com/watch?v=tuy33I4jgDg}
\end{scriptsize}
\end{center}

## Action point 3: Complete your research and dissertation

\begin{itemize}
  \item Working with yourself
  \vspace{0.7cm}
  \item Working with your peers
  \vspace{0.7cm}
  \item \textbf{Taking advantage of additional support activities (Term 3)}
  \vspace{0.1cm}
  \begin{itemize}
    \item Masterclass in Social Research: "Pieces-of-Craft" workshops and Dissertation Fairs
    \item Dissertation bootcamps and other workshops organized by the College's Study Skills team
  \end{itemize}
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/msr2023.pdf")
```
\vspace{0.15cm}
\begin{center}
\begin{scriptsize}
A preview of the Masterclass.
\end{scriptsize}
\end{center}

## Action point 4: Deferring your dissertation

\begin{itemize}
  \item Students who wish to defer must complete the mitigating circumstances (MCs) form online via \textbf{My Birkbeck}
  \vspace{0.2cm}
  \begin{itemize}
    \item Obtain the approval from your program director
    \item Submit your request at least two weeks prior to the deadline
    \item Send any inquiries to the Department/School admin team
    \item (Ideally) you should also have the consent from your dissertation supervisor and have a clear plan to complete your dissertation before the next deadline
  \end{itemize}
  \vspace{0.5cm}
  \item Students can submit their deferred dissertation by either \textbf{15 May} or \textbf{15 September} of the following year (i.e., 2024)
\end{itemize}

## Concluding remarks (until I see some of you in the Masterclass)

\begin{itemize}
	\item There is no perfect dissertation; there is only finished dissertation
	\begin{itemize}
	  \item A master's dissertation resembles a peer-reviewed journal article, consisting of \textbf{Introduction}, \textbf{Theory/Literature Review}, \textbf{Research Design, Data and Methods}, \textbf{Results and Discussion}, and \textbf{Conclusion}
	  \item Use \textbf{Apppendix} to present additional (and yet not essential) information about your work
	  \item Make sure you are aware of what to avoid and the assessment criteria (available in the Guidelines)
	\end{itemize}
	\vspace{0.6cm}
	\item The key is to establish your identity as an independent researcher
\end{itemize}

## Concluding remarks (until I see some of you in the Masterclass)

\begin{itemize}
	\item There is no perfect dissertation; there is only finished dissertation
\	\vspace{0.6cm}
	\item The key is to establish your identity as an independent researcher
	\vspace{0.1cm}
	\begin{itemize}
		\item Use the uploaded articles and dissertation samples as your guide
		\item Use literature review to inform us of the significance and originality of your research (e.g., Moreno et al 2019)
		\item Use research design to establish \textbf{a sense of presence/existence} in the field -- avoid unnecessarily tedious and apologetic textbook-style justification and refrain from being too obsessed with jargon and terminology (e.g., "document" analysis)
	\end{itemize}
\end{itemize}

## Key contacts

\begin{itemize}
  \item Your dissertation supervisor
  \vspace{0.5cm}
  \item Your program director
  \vspace{0.1cm}
  \begin{itemize}
    \item For MRes students in Politics: Dr Barry Maydom (\href{mailto:b.maydom@bbk.ac.uk}{\texttt{b.maydom@bbk.ac.uk}})
    \item For MRes/MSc Social Research students: Dr Chao-Yo Cheng (\href{mailto:c.cheng@bbk.ac.uk}{\texttt{c.cheng@bbk.ac.uk}})
  \end{itemize}
  \vspace{0.5cm}
  \item Department admin team: \href{mailto:politics-admin@bbk.ac.uk}{\texttt{politics-admin@bbk.ac.uk}} (include your \textbf{student ID number} in the subject line of your email message)
\end{itemize}

---

\vspace{0.6cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bbk_tv.pdf")
```
\vspace{0.2cm}
\begin{center}
Thank you!
\end{center}

## Assessment criteria

\begin{itemize}
	\item The research question is well-defined, and contextualized
	\vspace{0.1cm} 
	\item An argument is specified, coherently presented and supported by evidence
	\vspace{0.1cm} 
	\item Alternative/rival arguments are analysed
	\vspace{0.1cm}
	\item The approach is critical, not descriptive
	\vspace{0.1cm}
	\item A relevant, suitable methodology is employed 
	\vspace{0.1cm}
	\item Relevant sources (data and literatures) have been consulted; knowledge of relevant literature, issues and debates is demonstrated 
	\vspace{0.1cm}
	\item The style and presentation is clear and careful, and appropriate academic conventions have been observed
\end{itemize}

---

\vspace{0.2cm}
\begin{itemize}
	\item WK1: Designing your plan for writing
	\item WK2: Advancing your argument
	\item WK3: Abstracting your article
	\item WK4: Selecting a journal (X)
	\item WK5: Refining your works cited
	\item WK6: Crafting your claims for significance
	\item WK7: Analyzing your evidence
	\item WK8: Presenting your evidence
	\item WK9: Strengthening your structure
	\item WK10: Opening and concluding your article
\end{itemize}