---
title: |
  |
  |
  | Introduction to IR
  |
  |
author: |
  | WK2 Seminar
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Welcome back

\begin{itemize}
  \item Questions?
  \vspace{1cm}
  \item Study/Writing Skills workshops
  \vspace{1cm}
  \item Why you should attend seminars
\end{itemize}

# WWI as a critical juncture in IR

\begin{itemize}
  \item Why WWI is so important?
  \vspace{1.5cm}
  \item Can you name at least one ancient/recent event of similar importance but perhaps for a different reason?
\end{itemize}

# History and IR

\begin{itemize}
  \item What issues might come up when we use history to support our claims?
  \vspace{1.5cm}
  \item How can we avoid the potential pitfalls?
  \vspace{1.5cm}
  \item Is Kennedy's article a good example of using history to study IR? Why and/or why not?
\end{itemize}

---

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/iraq}
    \end{figure}
  \endcol
  \pause
  \begincol{.5\textwidth}
  \textbf{Daily political life in dictatorships largely remains a black box.}
  \pause
  \vspace{0.4cm}
  \begin{small}
    \begin{itemize}
      \item What did the Ba'thist regime actually know about its citizens?
      \vspace{0.15cm}
      \item Why did it use blunt, seemingly suboptimal, forms of punishment against its population?
      \vspace{0.15cm}
      \item What explains variation in the types of compliance and resistance behaviors undertaken by Iraqis during Hussein’s dictatorship?
    \end{itemize}
  \end{small}  
  \endcol
\endcols

---

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/iraq}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
    \begin{small}
    \vspace{0.2cm}
    \textbf{Archives.}
    \vspace{0.2cm}
    \begin{itemize}
      \item "The Iraqi Secret Police Files" (University of Colorado Boulder)
      \vspace{0.2cm}
      \item Miscellaneous tapes and associated transcripts (Hoover Institution)
    \end{itemize}
    \vspace{0.8cm}
    \textbf{Interviews} via the Iraq Memory Foundation.
    \end{small}
  \endcol
\endcols

# General principles of historical research in IR

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.5]{Figs/history}
    \end{figure}
  \endcol
  \pause
  \begincol{.5\textwidth}
  \begin{small}
  \begin{itemize}
    \item Be clear about the questions. Think through the strategies (induction v deduction).
    \vspace{0.1cm}
    \item Be mindful of the underlying politics of information production, conservation, and declassification.
    \vspace{0.1cm}
    \item Be transparent as you are trying to form your interpretations (as well as the selection of details to be reported).
    \vspace{0.1cm}
    \item Be patient and open-minded -- read (and document) as much information as possible. Do not jump into conclusion.
  \end{itemize}
  \end{small}
  \endcol
\endcols

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/hoover.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.hoover.org/library-archives}
\end{scriptsize}
\end{center}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/loc.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.loc.gov/}
\end{scriptsize}
\end{center}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/bl.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.bl.uk/}
\end{scriptsize}
\end{center}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/nas.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.nationalarchives.gov.uk/}
\end{scriptsize}
\end{center}
