---
title: |
  |
  |
  | Investigating the Social World
  |
  |
author: |
  | Week 2
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\vspace{0.7cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/lego.pdf")
```
\vspace{0.5cm}
\begin{center}
Welcome back!
\end{center}

# Welcome back

\begin{itemize}
  \item Questions?
  
  \vspace{0.2cm}
  \item Student-staff communication
  \vspace{0.1cm}
  \begin{itemize}
    \item Emails
    \item Office hours
    \item Student representative
    \item Substack program newsletter
  \end{itemize}
  
  \vspace{0.2cm}
  \item Assessment I: The methodological review
  \vspace{0.1cm}
  \begin{itemize}
    \item Examine a published \textbf{peer-reviewed} article
    \item Answer \textbf{three} questions
    \item Write up to \textbf{500 words} (max) for each
    \item Due on midday \textbf{31 March 2023}
  \end{itemize}
\end{itemize}

# More on the methodological review

\begin{small}
\begin{itemize}
  \item You will read closely one of the articles provided by us -- the questions and articles will be released on Moodle \textbf{at the beginning of Reading Week}.
  \vspace{0.2cm}
  \item The questions will be based on the ones for group exercise in class.
  \vspace{0.2cm}
  \item No outside/additional research is needed. The presumption is you do not need any citations. No need to quote anything.
  \vspace{0.2cm}
  \item It is not a test -- there is no rubric, and feel free to surprise us. You can come to my office hours to discuss any questions and/or your preliminary responses with me.
  \vspace{0.2cm}
  \item \textbf{Do not plagiarize}.
\end{itemize}
\end{small}

# Breaking the ice

\begin{itemize}
  \item Recap: To \textbf{produce valid knowledge claims}  (cool, but how?)
  
  \vspace{0.6cm}
  \item Today: "\textbf{Theory comes to rescue}" (?)
  \vspace{0.1cm}
  \begin{itemize}
    \item What is your subject area? Have you taken any module with "theory" in the title?
    \vspace{0.1cm}
    
    \item What is the most important/famous theory in your subject area? And what makes it important/famous?
    \vspace{0.1cm}
    
    \item What is your main takeaway from the TED Talk by Simon Sinek. What is your main takeaway? Do you like what you hear? Is there really a "theory" in the talk?
  \end{itemize}
\end{itemize}

# Plan for today

\begin{itemize}
  \item Part I
    \begin{itemize}
      \item Lecture: What is theory and what theory does for you
      \item Discussion: Using theory in social research
    \end{itemize}
  \vspace{1cm}
  \item Part II 
    \begin{itemize}
      \item Lecture: Use empirical to build/evaluate a theory
      \item Discussion: (Re)thinking rational choice theory
    \end{itemize}
\end{itemize}

---

\vspace{0.2cm}
\begin{center}
\textbf{Part I}
\end{center}
\vspace{0.4cm}
```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/theory_small.pdf")
```
\vspace{0.4cm}
\begin{center}
\textbf{What is theory, and what does it do (does not do)?}
\end{center}

---

\vspace{0.5cm}
\begin{itemize}
  \item What do we use theory for in social research?
  \vspace{1cm}
  \item Do we mean the same thing when we say "theory?"
  \vspace{1cm}
  \item How do you present and formulate your theory?
\end{itemize}

# Using theory to "make sense of" the social world

\begin{itemize}
  \item Description and interpretation
  \vspace{0.8cm}
  \item Explanation and prediction
\end{itemize}

# Using theory to "make sense of" the social world

\begin{itemize}
  \item Description and interpretation
  \vspace{0.1cm}
  \begin{itemize}
    \item Focus on holding a certain level of descriptive accuracy
    \item Focus on providing what we see a "meaning"
    \item Focus on conceptualizing and defining
  \end{itemize}
  \vspace{0.8cm}
  \item Explanation and prediction
\end{itemize}

# Using theory to "make sense of" the social world

\begin{itemize}
  \item Description and interpretation
  \vspace{0.8cm}
  \item Explanation and prediction
  \vspace{0.1cm}
  \begin{itemize}
    \item Focus on unpacking the structural and process mechanisms between different factors
    \item Focus on reaching a certain level of de-contextualization, for generalization (to a set of similar phenomena)
    \item Focus on forecasting and predictions (?)
  \end{itemize}
\end{itemize}

# Do we mean the same thing when we say "theory?"

\begin{itemize}
  \item Theory as ideology
  \vspace{0.8cm}
  \item Theory as paradigm
  \vspace{0.8cm}
  \item Theory as approach
\end{itemize}

# Do we mean the same thing when we say "theory?"

\begin{itemize}
  \item Theory as ideology
  \vspace{0.15cm}
  \begin{itemize}
      \item Ways of "envisioning"
      \item A set of systematic arguments about how the social world should be
      \item Call for norm/value making/changing
  \end{itemize}
  \vspace{0.8cm}
  \item Theory as paradigm
  \vspace{0.8cm}
  \item Theory as approach
\end{itemize}

# Do we mean the same thing when we say "theory?"

\begin{itemize}
  \item Theory as ideology
  \vspace{0.8cm}
  \item Theory as paradigm
  \vspace{0.15cm}
  \begin{itemize}
    \item Ways of "explaining" (or predicting)
    \item A set of somewhat consistent explanations and hypotheses
    \item Taken together, a "coherent" worldview
  \end{itemize}
  \vspace{0.8cm}
  \item Theory as approach
\end{itemize}

# Do we mean the same thing when we say "theory?"

\begin{itemize}
  \item Theory as ideology
  \vspace{0.8cm}
  \item Theory as paradigm
  \vspace{0.8cm}
  \item Theory as approach
  \vspace{0.15cm}
  \begin{itemize}
    \item Ways of "viewing"
    \item A claim that certain factors deserve our attention
    \item A belief that certain methods are the most and appropriate means of understanding
  \end{itemize}
\end{itemize}

# Different forms of "theory"

\begin{itemize}
  \item Verbal
  \vspace{0.8cm}
  \item Formal (mathematical)
  \vspace{0.8cm}
  \item Combined
\end{itemize}

# Different forms of "theory"

\begin{itemize}
  \item Verbal
  \vspace{0.1cm}
  \begin{itemize}
    \item Perhaps the most common form of theoretical expressions
    \item Use of language is important -- the analysis of "meaning" (or the discourse)
    \item Abstraction through metaphor and analogy
  \end{itemize}
  \vspace{0.8cm}
  \item Formal (mathematical/computational)
  \vspace{0.8cm}
  \item Combined
\end{itemize}

# Different forms of "theory"

\begin{itemize}
  \item Verbal
  \vspace{0.8cm}
  \item Formal (mathematical/computational)
  \vspace{0.1cm}
  \begin{itemize}
    \item "A model or set of rules and assumptions used to understand various behaviors in mathematical terms" (APA)
    \item A lot of existing/emerging fields, such as agent-based modeling, positive/game theory, social "physics," and complexity theory (or complex system)
    \item An active area calling for vibrant, constructive interdisciplinary research
  \end{itemize}
  \vspace{0.8cm}
  \item Combined
\end{itemize}

---

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/reed}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  "\textbf{We have disagreements}, that is, not only about \textbf{how we establish the sheer existence} of this or that social phenomenon, but also about \textbf{how we can claim to correctly and effectively explain, criticize, or interpret} it. In my view it is these latter disagreements [...] at the core of controversies about social knowledge."
  \endcol
\endcols

# Different forms of "theory"

\begin{itemize}
  \item Verbal
  \vspace{0.8cm}
  \item Formal (mathematical)
  \vspace{0.8cm}
  \item Combined: The "Analytic Narratives" project
  \vspace{0.1cm}
  \begin{itemize}
    \item An initiative started by several political economists in the 1990s
    \item Intersection of economics, economic history, and political science
    \item Integration of "elegant" rational choice-like theory and "thick" descriptions
  \end{itemize}
\end{itemize}

---

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/sfi}
    \end{figure}
  \endcol
  \begincol{.4\textwidth}
  "The rise of \textbf{complexity theory}, an \textbf{interdisciplinary} field studying the emergent \textbf{behavior and patterns of the interactions} of simple (and not so simple) \textbf{components}, has been one of the most important responses to the ballooning of \textbf{knowledge}."
  \endcol
\endcols

# The "sandcatle" syndrome

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/castles}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  "'My name is Ozymandias, king of kings: Look on my works, ye Mighty, and despair!' Nothing beside remains. Round the decay of that colossal wreck, boundless and bare The lone and level sands stretch far away."
  \endcol
\endcols

# Discussion I: "Theory" in social research?

\begin{itemize}
  \item Group exercise
  \begin{itemize}
    \item Share the article you have with each other.
    \item What does the theory say and do, if there is one?
    \item What are the main strategies of theoretical expression and construction?
    \item Is the article a victim of the "sandcastle" syndrome?
  \end{itemize}
  \vspace{0.6cm}
  \item Plenary discussion: Come up with an article and share with the whole class.
\end{itemize}

---

\vspace{0.2cm}
\begin{center}
\textbf{Part II}
\end{center}
\vspace{0.4cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/lego_2.pdf")
```
\vspace{0.4cm}
\begin{center}
\textbf{"Building" your theory}
\end{center}

---

\vspace{0.5cm}
\begin{itemize}
  \item Empirics serve as the bedrock for the theory-building exercise.
  \vspace{1cm}
  \item Can you test a theory? Quick answer: Not quite.
  \vspace{1cm}
  \item How can we establish the "alignment" between your theory and empirics?
\end{itemize}

# "Testing" a theory?

\begin{itemize}
  \item "A theory needs to be falsifiable"
  
  \vspace{0.1cm}
  \begin{itemize}
    \item Falsification is important: We can reject something if it is inconsistent with our findings
    \item Intuitively, this statement may seen fine -- many theory-claiming people cherry pick the evidence to serve their need
    \item The myth (or partial understanding) stems from the confusion with respect to the distinction between theory and hypothesis
  \end{itemize}
  \vspace{0.6cm}
  \item Testing theory v observable implications
  \vspace{0.6cm}
  \item Critical tests
\end{itemize}

# "Testing" a theory?

\begin{itemize}
  \item "A theory needs to be falsifiable"
  \vspace{0.6cm}
  \item Testing theory v observable implications
  \begin{itemize}
    \item Hypothesis is an "observable" implication we derive from our theoretical claims
    \item But our observation may be explained or interpreted by more than one theory
    \item To "formalize" the idea: Theory is the \textbf{sufficient} condition for your hypothesis, but your hypothesis is NOT necessarily a necessary condition for your theory
  \end{itemize}
  \vspace{0.6cm}
  \item Critical tests
\end{itemize}

# "Testing" a theory?

\begin{itemize}
  \item "A theory needs to be falsifiable"
  \vspace{0.6cm}
  \item Testing theory v observable implications
  \vspace{0.6cm}
  \item Critical tests
  \vspace{0.1cm}
  \begin{itemize}
    \item Ruling out the rival explanations
    \item Additional tests are needed to making a compelling theoretical claim based on your findings
    \item A good example is medical checks -- having a cold can mean so many different things
  \end{itemize}
\end{itemize}

# Varieties of theory-empirics alignment

\begin{itemize}
  \item Deduction
  \vspace{0.1cm}
  \begin{itemize}   
    \item Evaluating theoretical claims by trying to "falsify" the hypotheses derived from these claims
    \item We use data to "test" our hypothesis, but rejecting hypotheses may not mean we have to reject the theory (but it does reduce our faith in the theory, however)
    \item Ideal if you attempt to adjudicate competing theoretical claims
  \end{itemize}
  \vspace{0.8cm}
  \item Induction
  \vspace{0.8cm}
  \item Abduction
\end{itemize}

# Varieties of theory-empirics alignment

\begin{itemize}
  \item Deduction
  \vspace{0.8cm}
  \item Induction
  \vspace{0.1cm}
  \begin{itemize}
    \item Producing potentially generalizable claims based on our (systematic) observations
    \item Prone to any unfair or unwarranted rejections with only an "exception" or "outlier"
    \item Better for showing the existence a "theoretical" possibility
  \end{itemize}
  \vspace{0.8cm}
  \item Abduction
\end{itemize}

# Varieties of theory-empirics alignment

\begin{itemize}
  \item Deduction
  \vspace{0.8cm}
  \item Induction
  \vspace{0.8cm}
  \item Abduction
  \vspace{0.1cm}
  \begin{itemize}
    \item Starting by (surprisingly) finding that a hypothesis does NOT work well and may trigger the need for new data and interpretations/explanations
    \item Can be seen as a sequential integration of deductive-inductive processes
    \item Common for the task of theoretical refinement (or improvement)
  \end{itemize}
\end{itemize}

# Discussion II: (Re)thinking rational choice theory

\begin{itemize}
  \item Come up with a succinct definition of "rational choice theory."
  \vspace{0.3cm}
  \item Why does Barbara Geddes say "rational choice theory is not a theory?" Do you agree?
  \vspace{0.3cm}
  \item Why does Jon Elster say rational choice theory is "too ambitious?" Do you agree?
  \vspace{0.3cm}
  \item Should we ditch rational choice theory? Or we can try to make it better, using induction, deduction, and/or abduction?
\end{itemize}

# Barbara's verdict (sort of) on rational choice

\begin{itemize}
  \item There are many \textbf{misperceptions} about rational choice.
  
  \begin{itemize}
    \item Rational choice does not assume people are "selfish."
    \item Rational choice does not assume people are driven by material interests.
    \item Rational choice does not assume preferences are stable or fixed.
    \item Rational choice does not ignore/discard ideology and/or history.
  \end{itemize}
  
  \vspace{0.2cm}
  \item Rational choice is an \textbf{approach} (i.e., ways of seeing), not a theory.
  
  \begin{itemize}
    \item Rational choice highlights the importance of "rationality." 
    \item Rational choice does not dictate a single or universal explanation/prediction.
    \item Rational choice provides a micro-foundation for us to research the social world.
  \end{itemize}
\end{itemize}

# Concluding remarks

\begin{itemize}
  \item Foster "cultural competency" among social researchers
  \begin{itemize}
    \item Be open-minded. "Theory" can refer to very different things, and it is important for us to respect (and understand) the differences.
    \item Be transparent. Your preferences and the choices you make (and do your best to explain them) in your research.
    \item Be cautious. The way you treat and build theory (paradigm v approach) will affect the answers you offer.
  \end{itemize}
  \vspace{0.5cm}
  
  \item Bolster the theory-empirics alignment with \textit{high-quality} evidence
  \begin{itemize}
    \item Why should we use "numbers" to study the social world?
    \item Why is causality important? Is causality always a must?
    \item How can we generalize our insights beyond a case?
  \end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/kkv_hd.pdf")
```

---

\vspace{0.5cm}
\begin{center}
Thank you!
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bbk_tv.pdf")
```
\vspace{0.3cm}
\begin{center}
\texttt{c.cheng@bbk.ac.uk}
\end{center}

---

\vspace{0.5cm}
\begin{center}
Thank you!
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/links.pdf")
```
\vspace{0.3cm}
\begin{center}
\texttt{c.cheng@bbk.ac.uk}
\end{center}