---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Weel 1: Comparing}
\end{center}
\vspace{0.4cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week1.pdf")
```

# Today's plan

\begin{itemize}
  \item Module organization and logistics (30 mins)
  \vspace{0.1cm}
  \begin{itemize}
    \item Module organization
    \item Weekly syllabus and assessment
    \item Student support and ground rules
  \end{itemize}
  \vspace{0.2cm}
  \item Refresher lecture and Q\&A (1-1.5 hr)
  \vspace{0.1cm}
  \begin{itemize}
    \item Why and what to compare
    \item Theory as explanation
    \item Case selection, empirical strategies and research design
  \end{itemize}
  \vspace{0.2cm}
  \item Reading exercise: "Democratic Development in Africa" (1 hr)
\end{itemize}

# Module organization

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/bbk_night.pdf")
```
\vspace{0.3cm}
\begin{itemize}
\small
  \item Students can opt to attend one of the two sections listed below
  \begin{itemize}
    \item Tuesdays, 2-5pm (BCB 208): Dr Chao-Yo Cheng (convenor)
    \item Wednesdays, 6-9pm (BCB 312): Dr Kevin Manton
  \end{itemize}
  \item Activities: Q\&A, "refresher" lecture, reading exercise/workshop etc
\end{itemize}

# Assessment

\begin{itemize}
  \item Seminar log (20\%): a structured write-up of the discussion addressing both Sections 1 and 2 in the \textbf{lecutre} outline (500 words max)
  \vspace{0.4cm}
  \item Mini critical review (20\%): a critical review of a selected \textbf{seminar} article between Weeks 1-4 in light of class discussion of the chosen week (500 words max)
  \vspace{0.4cm}
  \item Essay (60\%): a synthetic \textbf{analysis} of based on a recent event in the past 12 months (1,500 words max)
\end{itemize}

# Student support

\begincols
  \begincol{.4\textwidth}
    \vspace{0.4cm}
    \begin{figure}
    \centering
    \includegraphics[scale=0.4]{Figs/head}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \hspace{0.2cm}
  \small
  \noindent \textbf{Module convenor}
  \vspace{0.1cm}
  \begin{itemize}
    \item Office hours (Friday 2-4pm)
    \item \url{c.cheng@bbk.ac.uk}
  \end{itemize}
  \vspace{0.5cm}
  \noindent \textbf{Module administration}
  \vspace{0.1cm}
  \begin{itemize}
    \item Ask (\url{https://www.bbk.ac.uk/ask})
    \item \url{fhss-edu-sss@bbk.ac.uk}
  \end{itemize}
  \vspace{0.5cm}
  \noindent \textbf{Student Services}
  \endcol
\endcols

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/student.pdf")
```
\vspace{0.1cm}
\begin{center}
\url{https://www.bbk.ac.uk/student-services}
\end{center}

# Ground rules and tips

\begin{itemize}
  \item \textbf{Active learning}: We embrace \textit{interactive} learning; come to in-person seminars/tutorials, having watched the lecture videos and completed the readings for the week
  \vspace{0.3cm}
  \item \textbf{Stability with flexibility}: We aim for stability, but acknowledge the need for flexibility
  \vspace{0.3cm}
  \item \textbf{Good communication}: We want to make this \textit{fantastic}; please be your own best advocate and give us feedback and constructive suggestions
  \vspace{0.3cm}
  \item \textbf{Safety}: We all want to keep you safe during your year at Birkbeck but need to work collaboratively with you to make that happen
\end{itemize}

# Strategic/deep reading in social sciences (UC Berkeley)

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucb.pdf")
```
\vspace{0.5cm}
\begin{itemize}
  \item \textbf{Preview}: Get as much information about the reading before you actually read it
  \vspace{0.1cm}
  \item \textbf{Annotating}: Read with a pencil and making notes as you read
  \vspace{0.1cm}
  \item \textbf{Analyzing}: Break the reading apart to see how different parts relate to each other
  \vspace{0.1cm}
  \item \textbf{Responding}: Think again how the reading relates the topic of each week; come up with questions
\end{itemize}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/study.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
\url{https://www.bbk.ac.uk/student-services/learning-development}
\end{center}

# "Refresher" lecture

\begin{itemize}
  \item Why and what to compare
  \vspace{0.5cm}
  \item Theory as explanation
  \vspace{0.5cm}
  \item Case selection, empirical strategies and research design
\end{itemize}

# Why and what to compare

\begin{itemize}
  \item \textbf{Why compare}: From state-sponsored area studies to theory-driven "comparative politics"
  \vspace{0.5cm}
  \item What to compare? 
  \vspace{0.1cm}
  \begin{itemize}
    \item Cross-sectional: Cross-national (global/regional) versus single-country
    \item Over time: Historical versus contemporary
    \item Key: Scope condition and "variations"
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/jcscott.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://www.nytimes.com/2024/07/28/books/james-scott-dead.html}
\end{center}

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/soas.pdf")
```
\vspace{0.1cm}
\pause
\begin{center}
\tiny
\url{https://www.theguardian.com/higher-education-network/gallery/2016/feb/09/one-hundred-years-of-soas-in-pictures}
\end{center}

# Why and what to compare

\begin{itemize}
  \item Why compare: From state-sponsored area studies to theory-driven "comparative politics"
  \vspace{0.5cm}
  \item \textbf{What to compare?} 
  \vspace{0.1cm}
  \begin{itemize}
    \item Cross-sectional: Cross-national (global/regional) versus single-country
    \item Over time: Historical versus contemporary
    \item Key: Scope condition and "variations"
  \end{itemize}
\end{itemize}

# Cross-national: Natural resources and authoritarianism

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/ross.pdf")
```

# Regional: Pandemic and support for regional integration

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/european.pdf")
```

# National (w/o comparison): Pandemic and attitudes towards benefits in the United Kingdom

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/uk.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/uk2.pdf")
```

# National (w/ comparison): Community organizations and local public goods provision in rural China

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/tsai.pdf")
```

# Timing matters: "Endogenous" democratization

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/endogenous.pdf")
```

# History is useful: German elections in the 19th century

```{r echo=FALSE, out.width="40%", fig.align="center"}
knitr::include_graphics("Figs/mares.pdf")
```

# Theory and concepts in comparative politics

\begin{itemize}
  \item Key concepts
  \begin{itemize}
    \item Regime, government and governance (and where institutions fit in?)
    \item Power (capacity) versus authority (rights)
  \end{itemize}
  \vspace{0.3cm}
  \item Why theory matters or not
  \vspace{0.1cm}
  \begin{itemize}
    \item Theory as explanation
    \item Theory as ideology
  \end{itemize}  
  \vspace{0.3cm}
  \item Question: Can/should culture be used to explain politics? How about "personal" experiences and identity?
\end{itemize}

# Culture and politics

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/culture.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ross_islam.pdf")
```

# Methods: Case selection, empirical strategies and research design

\begin{itemize}
  \item Case selection
  \begin{itemize}
    \item Cross-national (global or regional) and single-country
    \item Representative, exemplary, protypical and deviant
  \end{itemize}
  \vspace{0.3cm}
  \item Research design: Methods of agreement versus methods of difference
  \vspace{0.3cm}
  \item Empirical strategies: Qualitative, quantitative and computational (or "mixed")
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/agree.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/diff.pdf")
```

# Methods: Case selection, empirical strategies and research design

\begin{itemize}
  \item Case selection
  \begin{itemize}
    \item Cross-national (global or regional) and single-country
    \item Representative, exemplary, protypical and deviant
  \end{itemize}
  \vspace{0.3cm}
  \item Research design: Methods of agreement versus methods of difference
  \vspace{0.3cm}
  \item Empirical strategies: Qualitative, quantitative and computational (or "mixed")
\end{itemize}

# Qualitative: Case study, ethnography and elite interviews

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/chandra.pdf")
```

# Quantitative: Statistical analysis of survey data

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/dan.pdf")
```

# Computational: Topic modeling of social media texts

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/kpm_china.pdf")
```

# Discuss before break

\begin{itemize}
  \vspace{0.4cm}
  \item Any particular "cases" or topic would like to explore more?
  \vspace{0.4cm} 
  \item How does "globalization" change the study of comparative politics?
  \vspace{0.4cm}
  \item Is there anything from the lecture that you would like to know more about? Is there anything from the lecture that you are not clear about?
\end{itemize}

# Elite politics (political succession) and foreign policymaking in China

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/shirk.pdf")
```

---

\vspace{0.3cm}
\begin{columns}
  \begin{column}{0.5\textwidth}
  \begin{figure}
  \includegraphics[width=0.8\linewidth]{Figs/marjan}\\
  \vspace{0.1cm}
  \footnotesize Marjane Satrapi
  \end{figure}
  \end{column}
  \hspace{0.1cm}
  \begin{column}{0.5\textwidth}
  \small
    "If I have one message to give to the secular American people, it's that \textbf{the world is not divided into countries}. The world is not divided between East and West. You are American, I am Iranian, we don't know each other, but we talk together and we understand each other perfectly. \textbf{The difference between you and your government is much bigger than the difference between you and me}. And the difference between me and my government is much bigger than the difference between me and you. And our governments are very much the same."
  \end{column}
\end{columns}