---
title: |
  | 
  |
  | CP$^2$ Week 11: Social Protection and Redistribution
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Motivation: Returning to the dictator's dilemma and handbook

\begin{itemize}
  \item Wintrobe 1989: Contrary to the popular belief, the use of repression can be counter-productive for dictators, usually dictatorship operates through political exchange (co-optation)
  \vspace{0.5cm}
  \item Bdm et al 2002: The key to political survival is to properly manage the "3-D Politics": \textbf{Nominal selectorate}, \textbf{real selectorate} and \textbf{winning coalition} through reshuffling and redistribution
\end{itemize}

# Recap: Relative size of real selectorate and winning coalition is important

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/bdm.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/lasswell.pdf")
```

# Why social protection matters in authoritarian regimes and beyond

\begin{itemize}
  \small
  \item Selected distributive allocations of goods and services can contribute to political survival
  \vspace{0.2cm}
  \item Depending on their preferences and the expected returns, politicians may have incentives to focus on
  \begin{itemize}
    \item The swing or the core voters in the elections (Dixit and Londregan, 1996)
    \item The poor, because the arranged transfers will create the largest marginal utility (Diaz-Cayeros, Estevez, and Magaloni, 2015)
    \item The urban population, because they are in a better position to mobilize against the regime and autocrat (i.e., urban bias)
    \item Ethnic minorities so as to appease separatism sentiment
  \end{itemize}
  \vspace{0.2cm}
  \item Dictators follow a simiar logic (Magaloni, 2006; Blaydes, 2011; Mahdavi, 2015)
\end{itemize}

# "Social Policy in Developing Countries" (Mares and Carnes 2009)

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/mares.pdf")
```

# Three models of social protection in China (Solinger 2015)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.4]{Figs/solinger}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
    \item To extend assistance to accord with social citizenship rights (or to fulfill the Confucian concept of the rite of benevolence)
    \vspace{0.2cm}
    \item To offer subsidies to attain support or to pacify anger and silence demands from the poor
    \vspace{0.2cm}
    \item To grant benefits (education, health care) to enhance the nation's productivity
  \end{itemize}
  \endcol
\endcols

---



# Case study

\begin{itemize}
  \item Health care
  \vspace{0.6cm}
  \item Poverty reduction
  \vspace{0.6cm}
  \item Responses to the Covid-19 crisis
\end{itemize}

# Poverty alleviation

\begin{itemize}
  \item Poverty alleviation started as a national campaign in the early 1980s and its changes since then can reflect the evolution of the Chinese administrative state
  \vspace{0.2cm}
  \begin{itemize}
    \item 1980s: Work-for-relief programs and allocation of tangible goods
    \item 1990s: Fiscal transfers to and between local governments
    \item 2000s: Campaigns moving to targeting villages 
    \item 2010s: Campaigns moving to targeting households
  \end{itemize}
  \vspace{0.3cm}
  \item How should we understand these changes?
\end{itemize}

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/zeng.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/zuo.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/xinjiang.pdf")
```

# Health care

\begin{itemize}
  \item China provides a variety of health insurance plans for Chinese citizens, to name a few:
  \begin{itemize}
  \vspace{0.3cm}
    \item Urban Employee Basic Medical Insurance
    \vspace{0.2cm}
    \item New Rural Cooperative Medical Insurance
    \vspace{0.2cm}
    \item Urban Resident Basic Medical Insurance, which did not depend on employment status
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/health_gen.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/health_cov.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/health_region.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/health_reg.pdf")
```

# China's responses to Covid-19: Xi as a leader (Yuen Yuen Ang 2020)

\begin{itemize}
  \item Yuen Yuen Ang 2020: China under Xi has displayed both strengths and lethal shortcomings in handling COVID-19
  \vspace{0.1cm}
  \begin{itemize}
    \item Successfully mobilized a strong national response once the paramount leader gave the green light to act
    \item Lack of transparency and Xi’s initial inaction delayed decisive measures to curb the intial outbreak
  \end{itemize}
  \vspace{0.3cm}
\end{itemize}

# China's responses to Covid-19: China as a Leninist state (Victor Shih 2021)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.34]{Figs/covid}
    \end{figure}
  \endcol
  \hspace{0.2cm}
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
    \item The initial information suppression, the party largely succeeded in preventing the spread of not only information but also panic about COVID-19
    \vspace{0.2cm}
    \item The Leninist mobilization with the help from the co-opted civil society contributed to effective containment
    \vspace{0.2cm}
    \item The responses vary significantly across the country and show its inability to address challenges faced by both the urban and rural vulnerable population and migrant workers.
\end{itemize}
  \endcol
\endcols

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/dai.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/chengtao.pdf")
```

# Concluding remarks: Inequality is a political weapon

\begin{itemize}
  \item Social protection and redistribution contribute to authoritarian survival, and yet the mechanisms can vary, depending on the targeting beneficiaries and the type of resources being allocated
  \vspace{0.1cm}
  \begin{itemize}
    \item Distributive allocations can involve the combination of particularistic and local public goods
    \item Distributive allocations can help with political control, both in terms of coercion and co-optation
    \item Distributive allocations can yield micro- and macro-level political implications
  \end{itemize}
  \vspace{0.3cm}
  \item China's gigantic welfare state may be overshadowed by foreseeable economic challenges
  \vspace{0.3cm}
  \item The politics of social protection and redistribution may further question the conventional democracy v autocracy divide
\end{itemize}
