---
title: |
  |
  |
  | Social Research Dissertation Autumn Workshop
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

## Today's workshop is for

\begin{itemize}
  \item MRes students in Politics
  \vspace{0.7cm}
  \item MRes/MSc Social Research students
\end{itemize}

## Plan for today

\begin{itemize}
	\item Why dissertation (and what is it)?
  \vspace{0.2cm}
	\item An overview of the dissertation process
  \vspace{0.2cm}
	\item Getting started: Dissertation proposal form
  \vspace{0.2cm}
	\begin{itemize}
		\item Personal information
		\item Title and subject area
		\item "Scholarly" literature
		\item Data and methods
		\item Choosing (naming) your potential supervisor
	\end{itemize}
  \vspace{0.2cm}
	\item Concluding remarks and Q\&A
\end{itemize}

## Why dissertation? And what is it?

\begin{itemize}
  \item An \textbf{independent}, \textbf{original} research project, guided by your supervisor(s)
  \vspace{0.3cm}
  \item Resembles a \textbf{peer-reviewed journal article} (12,000-15,000 words, excluding appendix)
  \vspace{0.3cm}
  \item A \textbf{rite of passage}
  \begin{itemize}
    \item from a student to a scholar
    \item from consuming to producing knowledge
    \item from an empirical "technician" to a well-informed social "researcher"
  \end{itemize}
\end{itemize}

## Dissertation process: An overview

\begin{itemize}
	\item Term 1: Prepare your \textbf{dissertation proposal form}; talk to the program director and/or your potential supervisor (if there is one)
	\vspace{0.2cm}
	\item Terms 2 and 3
		\begin{itemize}
			\item January: Submit your proposal form (check deadlines on Moodle)
			\item February to early March: Supervisor allocation
			\item March to June/early July: Supervision in progress; complete ethics application and risk assessment, if necessary
			\item July-August: Supervision ends (typically) and complete your dissertation
			\item September: Submit your dissertation by \textbf{15 September 2024}
		\end{itemize}
	\vspace{0.2cm}
	\item After September 2024: Extension and deferral (TBA)
\end{itemize}

## Upcoming dissertation activities

\begin{itemize}
	\item Term 2: Dissertation workshop (by Dr Barry Maydom)
	\begin{itemize}
		\item Navigating the writing of your dissertation
		\item Managing the relationship with your supervisor
		\item Preparing ethics application and risk assessment
		\item Locating additional help or support
	\end{itemize}
	\vspace{0.4cm}
	\item Term 3: Social Research Dissertation Fairs (as part of Masterclass)
	\vspace{0.4cm}
	\item Term 3/summer break: Dissertation bootcamps (check with the Study Skills Team)
\end{itemize}

## Proposal form

\pause
\begin{small}
\begin{itemize}
  \item The proposal form is \textbf{NOT marked}; no written comments will be provided (but you can/should come to my office hours to talk about it)
  \vspace{0.3cm}
  \item The proposal form \textbf{helps us find you a supervisor} that best suits your research project
  \vspace{0.3cm}
  \item The proposal form \textbf{shows your tentative ideas} for the dissertation
  \vspace{0.3cm}
  \item The proposal form should be \textbf{short, succinct} but \textbf{sufficiently coherent and informative} (2-2.5 pages max)
  \vspace{0.3cm}
  \item The proposal form is due in January on Moodle; submission link and \textbf{sample proposal forms} are available on Moodle
\end{itemize}
\end{small}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/form.pdf")
```

## Proposal form: Basic personal information

\begin{itemize}
  \item Name
  \item Email address
  \item MSc/MRes programme
  \item Full-time or part-time
\end{itemize}
\pause
\vspace{1.2cm}
\begin{center}
\textbf{The information you provide needs to be consistent with what we see on MyBirkbeck}
\end{center}

## Proposal form: Title and subject area

\begin{itemize}
  \item Title of proposed dissertation: Short v long? Creative v informative?
  \vspace{0.3cm}
  \pause
  \item Description of subject area: A brief statement to introduce your research (like a paper's abstract)
  \begin{itemize}
    \item What are the \textbf{research questions}?
    \item What is the \textbf{focus of previous research literature}?
    \item What is the \textbf{context and/or background}? What is the \textbf{scope condition}?
    \item What is \textbf{distinctive about your theory position or intellectual approach}?
    \item What are \textbf{the tentative methods and data sources}? 
    \item \textbf{[Not yet]} What are your benchmark findings? What "new facts" have you found? Or what key substantive conclusions do you draw? The value-added or originality of your work within this field?	
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/mason.pdf")
```
\begin{center}
\begin{scriptsize}
Mason, J. 2002. \textit{Qualitative researching} (2nd). London: Sage, p.18.
\end{scriptsize}
\end{center}

## Proposal form: "Scholarly literature"

\begin{itemize}
  \item List the relevant "scholarly" literature(s); arrange the literature(s) by topic/keyword if you'd like
  \pause
  \vspace{0.2cm}
  \item Identify the gap, such as
  \begin{itemize}
    \item Theoretical/empirical tensions or contradictions
    \item Different or alternative theoretical perspectives
    \item Different methodological approach (e.g., build on more inductive work to propose some hypothesis-testing deductive approach)
    \item Extension to new context (e.g., different region, time period)
    \item Some conventional wisdom left untested or understudied (e.g., critical element not engaged with, need to consider other conditions)
  \end{itemize}
  \vspace{0.2cm}
  \item Use literature to think through your RQs and supervisor preferences
\end{itemize}

---

```{r echo=FALSE, out.width="94%", fig.align="center"}
knitr::include_graphics("Figs/say.pdf")
```

## Proposal form: "Primary research material"

\begin{itemize}
  \item A suggestive research design where you
  \begin{itemize}
    \item describe the type(s) of evidence which will be collected and analyzed to address the research question
    \item show that the proposed evidence can feasibly be collected given the time constraint
    \item outline feasible method(s) that will be used to analyze the evidence
  \end{itemize}
  \vspace{0.2cm}
  \item Note it is okay to be tentative (we know things will change)
  \vspace{0.2cm}
  \item The key is to build a coherent alignment between question, theory, and empirics; you can use your observations to \textit{evaluate} or \textit{challenge} a theory
  \item The key is to provide original analysis, which can be based on existing data
\end{itemize}

---

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/ssrc.pdf")
```
\begin{center}
\begin{small}
\url{https://www.ssrc.org/publications/the-art-of-writing-proposals/}
\end{small}
\end{center}

## Proposal form: Potential supervisor

Supervisor allocation is a \textbf{collaborative} process between the Program Director and you; you can/need to
\pause
\vspace{0.3cm}
\begin{itemize}
  \item Nominate one or more (permanent) academic staff members from
  \begin{itemize}
    \item School of Social Sciences
    \item Other subject areas outside SSS, such as applied linguistics, criminology, and law
  \end{itemize}
  \vspace{0.3cm}
  \item Show some knowledge of the work by your potential supervisor and provide  explanations in the proposal form
  \vspace{0.3cm}
  \item Start informal conversations with me and/or your potential supervisors
\end{itemize}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/staff.pdf")
```
\begin{center}
\begin{scriptsize}
\url{https://www.bbk.ac.uk/school/social-sciences/our-staff}
\end{scriptsize}
\end{center}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/tom.pdf")
```

## Course of action (until we meet again in 2024)

\begin{itemize}
  \item Revise workshop slides and recommended readings on Moodle
  \vspace{1cm}
  \item Start thinking about your dissertation (if you haven't) and talk to your personal tutor and/or potential supervisor
  \vspace{1cm}
  \item Work on the proposal form and submit it by the deadline
\end{itemize}

## Key contacts

\begin{itemize}
  \item MRes in Politics: Dr Barry Maydom (\href{mailto:b.maydom@bbk.ac.uk}{\texttt{b.maydom@bbk.ac.uk}})
  \vspace{0.7cm}
  \item MRes/MSc Social Research: Dr Chao-Yo Cheng (\href{mailto:c.cheng@bbk.ac.uk}{\texttt{c.cheng@bbk.ac.uk}})
\end{itemize}

---

Thank you.