---
title: |
  | 
  |
  | Doing Political and Social Research 
  |
  |
author: |
  | Research Proposals
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Overview

\begin{itemize}
  \item A \textbf{2500-word} ($+/- 10\%$) research design to outline a planned research project
  \vspace{0.1cm}
  \item Submit your work by \textbf{Midday, 10 January November 2025}
  \vspace{0.1cm}
  \begin{itemize}
  	\item Attach the assessment coversheet
  	\item Enter your student ID in the field of Submission Title
  \end{itemize}
\end{itemize}
\vspace{0.6cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/turnitin.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/process.pdf")
```

---

\vspace{0.2cm}
\begin{itemize}
\small
	\item What \textbf{ontological and epistemological foundations} underpin your research philosophy?
	\item What is your research \textbf{question}?
	\item Where does your proposed research build on \textbf{existing studies}?
	\item What is your \textbf{theory}? Will you be testing an existing theory, developing a new theory or adapting an exisiting theory to a different context?
	\item What are the key \textbf{concepts} in your research project? How will you operationalise and measure them?
	\item Which \textbf{method(s)} will you use to answer your question? Why do you choose this method, or sets of method? How will you collect your data? How will you analyse the data you collect? What practical difficulties might you face when collecting and analysing your data? How will you overcome them?
	\item What \textbf{ethical issues} might arise during your research project? How will you ensure that your research meets relevant ethical standards?
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/moodle.pdf")
```

# Extra: More on a research proposal

\begin{itemize}
  \item State the \textbf{research question(s)} that you could answer within the constraints of a dissertation
  \item Provide the \textbf{background literature} -- what is the body of literature that your RQs emerges from and that you plan to speak to? 
  \item Present a \textbf{research design} to explain data collection and analysis
  \item Discuss the \textbf{limitations and critical considerations} specific to your research design
  \item Bibliography
\end{itemize}

---

\begin{itemize}
	\item Research question(s) that are sufficiently focused and tractable
	\item Literature review that sets the relevant context and motivates the research question -- what is the gap?
	\item A research design that
	\begin{itemize}
		\item describes the type(s) of evidence which will be collected and analyzed to address the research question;
		\item shows that the proposed evidence can feasibly be collected in the timeframe available;
		\item outlines feasible method(s) that will be used to analyze the evidence
	\end{itemize}
	\item A thoughtful discussion on the expected relative strengths and limitations of the combination of evidence and analytical method
\end{itemize}

# "On the Art of Writing Proposals" (SSRC)

\begin{figure}
  \centering
  \includegraphics[scale=0.25]{Figs/ssrc}
\end{figure}
\begin{center}
\begin{small}
\url{https://www.ssrc.org/publications/the-art-of-writing-proposals/}
\end{small}
\end{center}

# Extra: Explaining your design: Process tracing and case studies

\begin{itemize}
	\item Literature review: How do you use the literature to motivate the study? What is the puzzle? Why is this study interesting, in relation to what we already know or assume?
	\item How do you justify the selection of cases for comparison? What is the evidence that these cases are different on the independent and dependent variables while being mostly similar in other aspects?
	\item What is your proposed explanation? Where does the evidence come from? What is the most convincing piece of evidence?
	\item Are there any alternative explanations in addition to the central argument?
	\item Which part of your argument could be generalized to other contexts? What is the scope condition?
\end{itemize}

# Extra: Explaining your design: Quantitative research design

\begin{itemize}
  \item Basic components
  \begin{itemize}
    \item The key empirical \textbf{question(s)}?
	  \item The key \textbf{hypotheses} (empirical or testable implications) based on the theoretical literature?
	  \item The key \textbf{independent} (explanatory) and \textbf{dependent} (outcome) variables?
	  \item The key \textbf{statistical result(s)} we expect to observe in the regression table(s)?
  \end{itemize} 
  \item Advanced components
  \begin{itemize}
    \item Are all necessary \textbf{control} variables (covariates) included in the model?
	  \item Are there any additional tests to assess the \textbf{robustness} of the main results? Are there any tests to study the \textbf{heterogeneous} effect (i.e., do the main findings change if we divide up the observations)?
	  \item What else can be or should have been tested?
  \end{itemize}
\end{itemize}

# Extra: Explaining your design: Qualitative research design

\begin{itemize}
  \item What is the key concept from the literature that you seek to engage with? How do you motivate your study with reference to the literature?
  \item How do you justify the case selection? Is it a typical case? An exceptional case? What puzzle can it contribute to solving?  
  \item How will the qualitative methods used allow you to answer the research questions? Is there anything that might be missed out, with your design?  
  \item Is anything in your research paper generalizable to other settings? 
  \item How does your research provide a sense of "being there?" 
\end{itemize}

---

\begin{figure}
\centering
  \includegraphics[scale=0.35]{Figs/tom}
\end{figure}
\vspace{0.2cm}
\begin{scriptsize}
\begin{center}
\url{https://press.uchicago.edu/ucp/books/book/chicago/W/bo131341275.html}
\end{center}
\end{scriptsize}
