---
title: |
  | 
  |
  | Chinese Politics in Comparative Perspective
  |
  |
author: |
  | Learning Journal
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Overview 

\begin{itemize}
  \item \textbf{Main task}: Read a published peer-review article in leading political science journals to reflect on how China is researched and compared
  \vspace{0.3cm}
  \begin{itemize}
    \item Choose an article from the following journals (the article should be published after 2010 and explicitly focuses on China)
    \vspace{0.2cm}
    \item Bring the article to class to share with everyone in \textbf{Week 6}
    \vspace{0.2cm}
    \item Read the article and write an essay to answer two questions, using up to \textbf{1,000 words} (max)
    \vspace{0.2cm}
    \item Submit your work by \textbf{Midday, 7 April 2024}
  \end{itemize}
  \vspace{0.3cm}
  \item No external references are required or should be used
\end{itemize}

---

\begin{itemize}
  \item American Political Science Review
  \item British Journal of Political Science
  \item Comparative Politics
  \item Comparative Political Studies
  \item Journal of Politics
  \item Political Research Quarterly
  \item Political Studies
  \item Studies in Comparative International Development
  \item World Politics
\end{itemize}

---

\vspace{1cm}
\begin{itemize}
\small
  \item In your words, explain how the author(s) think China can be compared with the country (or countries). Can you identify a country that can also be included using the same comparative logic? {Do you agree with such a comparison?} Discuss.
  \vspace{0.5cm}
  \item Use \textbf{Google Scholar} to explore how other scholars have cited the article you select. What are some common themes among these articles? Is the article you choose predominantly cited by scholars who specialize in China, or has it also garnered attention from scholars who do not focus on China? Do you believe that the article, while centered on China, holds broader implications for scholars who study other countries or regions? Why? Discuss.
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/example.pdf")
```

# Keep in mind

\begin{itemize}
  \item Answer every part of both questions
  \item Attach the asseesment coversheet (Moodle)
  \item Enter your student ID in the field of \textbf{Submission Title} (see below)
\end{itemize}
\vspace{0.6cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/turnitin.pdf")
```