---
title: |
  | 
  |
  | Postgraduate Social Research Programs 
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Plan for today

\begin{itemize}
  \item Program structure and requirements
  \vspace{0.2cm}
  \item Learning resources and opportunities
  \vspace{0.2cm}
  \item Postgraduate academic life
  \vspace{0.2cm}
  \item Q\&A
\end{itemize}

# Program structure and requirements

\begin{itemize}
\small
  \item Four 15-credit \textbf{research methods} modules (60 credits in total)
  \begin{itemize}
  \begin{small}
    \item Qualitative Social Research Methods and Qualitative Research Design
    \item Introduction to Quantitative Social Research, Applied Social Data Analysis in \texttt{R} and Advanced Topics in Quantitative Social Research
    \item Experiments in Social Science
  \end{small}  
  \end{itemize}
  \vspace{0.25cm}
  \item Two 30-credit \textbf{theory/subject} options (60 credits in total)
  \begin{itemize}
  \begin{small}
    \item MSc and PG Diploma: Investigating the Social World and Masterclass in Social Research
    \item MRes: Two options from your subject area
  \end{small}
  \end{itemize}
  \vspace{0.25cm}
  \item \textbf{Social Research Dissertation} (60 credits), including four dissertation workshops
\end{itemize}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/timetable_2024.pdf")
```
\vspace{0.3cm}
\begin{center}
Compulsory and Core SR modules, 2024-2025
\end{center}

# Program structure and requirements

\begin{itemize}
  \item We offer flexibility such that you can 
  \vspace{0.1cm}
  \begin{itemize}
    \item Choose to finish your study full-time or part-time
    \item Audit modules and/or take modules outside your subject area
    \item Progress to MSc/MRes after you finish PG Certificate/Diploma
    \item Exit with PG Cert/Diploma even if you start with MSc/MRes
  \end{itemize}
  \vspace{0.2cm}
  \item We will revisit your modules and/or your plan at/after Birkbeck in 1:1 meetings
\end{itemize}
\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/com.pdf")
```

# Dissertation process: An overview

\begin{itemize}
\small
  \item Autumn (Term 1)
  \begin{itemize}
    \item November: Workshop on how to get started
    \item Preliminary meetings with me and/or potential supervisor
  \end{itemize}
  \vspace{0.7cm}
  \item Spring (Term 2)
  \begin{itemize}
    \item January-February: Submit your proposal form (20 January) and supervisor allocation
    \item March: Workshop on literature review and research ethics
  \end{itemize}
  \vspace{0.7cm}
  \item Summer (Term 3)
  \begin{itemize}
    \item June: Dissertation Fairs through Masterclass in Social Research
    \item July: Workshop on writing up your dissertation (optional)
  \end{itemize}
  \item September: Submit dissertation (September 15) or request deferral (TBA)
\end{itemize}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/tom.pdf")
```

# Studying at Birkbeck

\begin{itemize}
  \item Each module has a handbook, with detailed description of contact points, aims and objectives, learning outcomes, syllabus, and assessment requirements
  \vspace{0.4cm}
  \item Most modules will be taught by pre-recorded lectures combined with in-person seminars or lab tutorials (for quantitative modules)
  \vspace{0.4cm}
  \item You will need to finish pre-sessional activities (e.g., readings and lecture recordings) before the live sessions and actively engage in class activities
  \vspace{0.4cm}
  \item You will have the chance to provide feedback (student-staff exchanges and module evaluation)
\end{itemize}

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/masterclass.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/services.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/mybbk.pdf")
```

# Learning outside classrooms

\begin{itemize}
  \item Birkbeck's Study Skills workshops
  \vspace{0.2cm}
  \item Birkbeck's research centres and institutes
  \vspace{0.2cm}
  \item Birkbeck's career services; opportunities on LinkedIn, \url{jobs.ac.uk}, and \url{idealist.org}
  \vspace{0.2cm}
  \item Libraries, including those outside Birkbeck, such as the Senate House, SOAS, UCL, and the British Library -- check out \textbf{SCONUL}
  \vspace{0.2cm}
  \item Additional events and resources outside Birkbeck (e.g., UoL colleges, Turing and NCRM) 
\end{itemize}

# Learning outside classrooms: Study Skills workshops

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/study_skill.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.bbk.ac.uk/student-services/learning-development}
\end{scriptsize}
\end{center}

# Learning outside classrooms: Research centres and institutes

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/centre_inst.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.bbk.ac.uk/research/centres}
\end{scriptsize}
\end{center}

# Learning outside classrooms: Career planning

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/careers.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.bbk.ac.uk/student-services/careers-and-enterprise}
\end{scriptsize}
\end{center}

# Academic life as a PG student

\begin{itemize}
  \item \textbf{Active learning}: We embrace \textit{interactive} learning; come to in-person seminars/tutorials, having Watched the lecture videos and completed the readings or other preparation for the week
  \vspace{0.2cm}
  \item \textbf{Stability with flexibility}: We aim for stability, but acknowledge the need for flexibility
  \vspace{0.2cm}
  \item \textbf{Good communication}: We want to make this \textit{fantastic}; please be your own best advocate and give us feedback and constructive suggestions
  \vspace{0.2cm}
  \item \textbf{Safety}: We all want to keep you safe during your year at Birkbeck but need to work collaboratively with you to make that happen
\end{itemize}

# Key support and contact points

\begin{itemize}
  \item \textbf{Program director} and personal tutor: \textbf{Dr Chao-Yo Cheng} (\href{mailto:c.cheng@bbk.ac.uk}{\texttt{c.cheng@bbk.ac.uk}}) for academic and research questions; office hours \textbf{between 3-5pm on Fridays} via Microsoft Teams (or by appointment)
  \vspace{0.2cm}
  \item \textbf{Program administrator} for admin or procedural matters
  \vspace{0.2cm}
  \item \textbf{Module conveners} (see Moodle pages)
  \vspace{0.2cm}
  \item \textbf{Dissertation supervisor}
  \vspace{0.2cm}
  \item \textbf{Student Services} provided by the College
\end{itemize}

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/bbkask.pdf")
```
\begin{center}
\url{https://www.bbk.ac.uk/ask}
\end{center}
\vspace{0.5cm}
\begin{itemize}
\small
  \item All student inquiries should first be submitted to the \textbf{Ask} system
  \vspace{0.1cm}
  \item Unless instructed, students are suggested not email individual administrators
  \vspace{0.1cm}
  \item If unsure or encountering exceptional circumstances, get in touch with the program director for advice
\end{itemize}

---

\vspace{1cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bbk_tv.pdf")
```
\vspace{0.5cm}
\begin{center}
Stay curious, stay active and stay healthy
\end{center}
