---
title: |
  |
  |
  | Investigating the Social World
  |
  |
author: |
  | Week 11
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\vspace{0.7cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/lego.pdf")
```
\vspace{0.5cm}
\begin{center}
Welcome back!
\end{center}

---

\vspace{0.8cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/cgg.pdf")
```
\vspace{0.3cm}
\begin{center}
\textbf{Comparative and international analysis}
\end{center}

---

\vspace{0.4cm}
\begin{itemize}
  \item Why comparative and intl analysis
  \vspace{0.7cm}
  \item Move from case studies to social science
  \vspace{0.7cm}
  \item Practical guides for generalization
  \vspace{0.7cm}
  \item Conclusion: Investigating the social world as a vocation
\end{itemize}

# Why comparative and intl analysis

\begin{itemize}
  \item Prevent selection bias (Geddes 1990; Collier and Mahoney 1996)
  \vspace{1cm}
  \item Get around essentialism (Fuchs 2001; Phillips 2010)
  \vspace{1cm}
  \item Boost significance/importance
\end{itemize}

# The cases you choose affect the answers you get

\begin{itemize}
  \item The cases you choose can be problematic for the purpose of generalization.
  \vspace{0.1cm}
  \begin{itemize}
    \item The cases you choose may only represent one or a small number of similar cases in the domain/universe.
    \item Without identifying the domain/universe/population, it is very difficult to ascertain the nature of your cases (typical/representation OR deviant/outlier/niche) and their analytical value (Gerring 2006).
  \end{itemize}
  \vspace{0.2cm}
  \item Selecting cases on dependent variable is very likely to produce faulty inference. 
  \vspace{0.1cm}
  \begin{itemize}
    \item Jumping into the conclusion that any characteristic shared by the selected cases is the cause.
    \item Jumping into the conclusion that the observed pattern can be extended to cases not included in the study.
  \end{itemize}
\end{itemize}

# Labor repression and economic growth, 1970-81 (Geddes 2003)

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/sample.pdf")
```
\begin{center}
\small
When we only consider countries in Asia, labor repression seems to drive economic growth.
\end{center}

# Labor repression and economic growth, 1970-81 (Geddes 2003)

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/all.pdf")
```
\begin{center}
\small
When we consider all countries, there exists no clear statistical association.
\end{center}

# Be aware of the "pitfalls" of essentialism 

\begin{itemize}
  \item Essentialism can take at least the following four different forms in social research (Phillips 2010)
  \vspace{0.1cm}
  \begin{itemize}
    \item Attribution of certain characteristics to all observations subsumed within a particular category (ignoring individual agency and micro-level variation)
    \item Attribution to the category by itself (naturalizing things that are socially and historically constructed)
    \item Invocation of a collectivity, presuming a homogenized and unified group (sustaining stereotype)
    \item Policing of the presumed collective category uncritically (impeding innovation by suffocating contestation)
  \end{itemize}
  \pause
  \vspace{0.4cm}
  \item \textbf{Discussion}: Is attribution to categories always a bad idea? How can we get around these potential issues?
\end{itemize}

# From case studies to social science: Two forms of "comprehension"

\begin{itemize}
  \item Apprehension
  \vspace{1cm}
  \item Explanation
\end{itemize}

# From case studies to social science: Two forms of "comprehension"

\begin{itemize}
  \item Apprehension
  \vspace{0.1cm}
  \begin{itemize}
    \item "Immersion" (often via ethnography) as the key foundation of learning and understanding to inform judgment and sharpen intuition.
    \item The experiences in the field often provides a "bullshit meter" (or more commonly as we put, "plausibility check").
    \item The first step "comes in the field itself" before "the researcher begins to separate from the field" and moves towards "explanation." 
  \end{itemize}
  \vspace{0.7cm}
  \item Explanation
\end{itemize}

# From case studies to social science: Two forms of "comprehension"

\begin{itemize}
  \item Apprehension
  \vspace{0.7cm}
  \item Explanation
  \vspace{0.1cm}
  \begin{itemize}
    \item The movement towards "explanation" involves the "recoding" of the observations and encounters in the field.
    \item Explanation establishes \textbf{the} lines of logic to link causes and consequences; the \textbf{thick} description provides the underpinnings for the exercise of abstraction and theory-building.
    \item Explanation is different from \textbf{interpretation}, which focuses on providing "a" meaning of the descriptions.
  \end{itemize}
\end{itemize}

# Economic behaviors of coffee growers in East Africa and Colombia

\begin{itemize}
  \item Use the field to search for potentially testable and generalizable insight (through ethnography)
  \vspace{0.1cm}
  \begin{itemize}
    \item The behavior of the coffee growers represented a choice by rational actor under varying institutional environment
    \item Be cautious about the temptation of using culture to explain the consequences
  \end{itemize}
  \vspace{0.2cm}
  \item Scope through the initial cases for plausibility check and theory building (through archival research and interviews)
  \vspace{0.2cm}
  \item Pave the road to conviction with small-N and large-N evidence
  \vspace{0.1cm}
  \begin{itemize}
    \item Test the proposed logic in comparable new cases
    \item Test the proposed logic with large-N (muiltivariate) regression
  \end{itemize}
\end{itemize}

# Some practical suggestions for generalization

\begin{small}
\begin{itemize}
  \item A researcher can be driven by the general question rather than the case (e.g., Rob Bates and David Laitin).
  \vspace{0.12cm}
  \item If possible, think outside the box by formulating the research questions within and outside the given case (i.e., de-contextualization).
  \vspace{0.12cm}
  \item Define the \textbf{scope} condition carefully and explicitly -- that is, what are the comparable cases out there and why? Where does the case of interest come from?
  \vspace{0.12cm}
  \item Use the key stakeholders and factors in the case to search for the possibility of comparable cases.
  \vspace{0.12cm}
  \item Be careful with any cultural explanations as they may create some issues (e.g., reverse causality; circularity and tautology; loss of agency).
  \vspace{0.12cm}
  \item Do not force generalization when it is not the main objective of your research. Do not be delusional and mislead yourself and your audiences.
\end{itemize}
\end{small}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/mason.pdf")
```
\begin{center}
\begin{scriptsize}
Mason, Jennifer. 2002. \textit{Qualitative researching} (2nd). London: Sage, p.18.
\end{scriptsize}
\end{center}

---

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \vspace{0.4cm}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/laitin}
    \end{figure}
    \vspace{0.2cm}
    \begin{center}
    \textbf{David D. Laitin} (Stanford)
    \end{center}
  \endcol
  \begincol{.5\textwidth}
  \begin{small}
  \begin{itemize}
    \item \textit{Why Muslim Integration Fails in \textbf{Christian-Heritage Societies}} (Cambridge: Harvard University Press, 2016) with Claire Adida and Marie-Anne Valfort
    \vspace{0.2cm}
    \item \textit{Identity in Formation: the \textbf{Russian-speaking} Populations in the Near Abroad} (Oxford: Oxford University Press, 2007)
    \vspace{0.2cm}
    \item \textit{Nations, States and Violence} (Ithaca: Cornell University Press, 1998)
    \vspace{0.2cm}
    \item \textit{Language Repertoires and State Construction in \textbf{Africa}} (Cambridge: Cambridge University Press, 1992)
  \end{itemize}
  \end{small}
  \endcol
\endcols

# Conclusion: Investigating the social world as a vocation

\begin{itemize}
  \item ISW is a special module, reflecting on the production of "valid" knowledge about the social world.
  \begin{itemize}
    \item You need to know what you do and how to explain the objectives of your scholarly inquiry;
    \item You need to know how to tell the differences between different objectives and engage other researchers' work constructively and effectively.
  \end{itemize}
  \item Social research involves a series of careful thinking about ontology/epistemology, theory-building, and methodology.
  \begin{itemize}
    \item \textbf{Reveal nuances of the real world}: constructivism; theory as approach; interpretation; qualitative; inductive/abductive
    \item \textbf{Search for the "ultimate" truth}: positivism; theory as paradigm; explanation; quantitative; deductive
  \end{itemize}
  \item Exceptions always exist and boundaries are constantly being contested and redefined (e.g., Bates).
\end{itemize}

---

\vspace{0.7cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/epis_cont.pdf")
```

# Discussion: "Solidary Groups, Informal Accountability, and Local Public Goods Provision" (Tsai 2007)

\begin{itemize}
  \item What is the main research question? What is the scope condition?
  \vspace{0.5cm}
  \item Can you summarize her main theoretical argument? Can/should culture serve as the alternative explanation?
  \vspace{0.5cm}
  \item Are the findings significant and/or valid? Can you come up with another research project to extend this line of inquiry?
\end{itemize}

---

More questions to consider about the research design and methods:
\vspace{0.4cm}
\begin{itemize}
  \item Measurement of key analytical variables (solidary groups and local public goods provision)
  \vspace{0.2cm}
  \item Choice of estimator (seemingly unrelated regression)
  \vspace{0.2cm}
  \item Identification strategy (instrumental variable)
  \vspace{0.2cm}
  \item Robustness and additional/critical tests
  \vspace{0.2cm}
  \item Extension and further analysis?
\end{itemize}
