---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | Open Social Science: Transparency and Replicability
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Logistics: Dissertation Fairs

\begin{itemize}
	\item Six students have submitted their abstracts and bios
	\vspace{0.2cm}
	\item We should be able to accommodate your first week preferences -- each week will feature two (or three if we have any late submissions) students
	\vspace{0.2cm}
	\item You will also receive a separate email from me containing the comments for your reference
	\vspace{0.2cm}
	\item Preliminary program released on Moodle tomorrow; you are welcome to invite your supervisors
\end{itemize}

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/nk_crop.pdf")
```
\vspace{0.1cm}
\begin{center}
\url{https://www.columbiaspectator.com/news/2019/09/12/history-professor-charles-armstrong-found-guilty-of-plagiarism-to-retire-in-2020/}
\end{center}

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucla_crop.pdf")
```
\vspace{0.1cm}
\begin{center}
\url{https://www.science.org/doi/10.1126/science.1256151}
\end{center}

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="105%", fig.align="center"}
knitr::include_graphics("Figs/pnas.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\textit{Proceedings of the National Academy of Sciences} (2022)
\end{scriptsize}
\end{center}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/bitss.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.bitss.org/}
\end{scriptsize}
\end{center}

# What We Can/Should Do

\begin{itemize}
	\item Transparency
	\vspace{0.1cm}
	\begin{itemize}
		\item "Research Design" section
		\item Pre-analysis plan (PAP)
	\end{itemize}
	\vspace{0.2cm}
	\item Replicability: Data and code sharing, such as Harvard Dataverse and Birkbeck Data Repository
	\vspace{0.2cm}
	\item Knowledge accumulation: Meta-analysis and systematic review?
	\vspace{0.2cm}
	\item Triangulation: Mixed- and multi-method research?
\end{itemize}

# Discussion

\begin{itemize}
	\item What are the social research methods or empirical approaches you will be using in your dissertation?
	\vspace{0.4cm}
	\item What information in the main text should you provide in writing aligning with the broad ideas of transparent and replicable social sciences research?
	\vspace{0.4cm}
	\item What additional information should you provide in the appendix? Any practical and ethical concerns?
\end{itemize}