---
title: |
  | 
  |
  | CP$^2$ Week 3: Structures of the Party-State
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Recap: From empire to nation-state (1911-1949)

\begin{itemize}
  \item State building was challenging
  \vspace{0.1cm}
  \begin{itemize}
    \item Powerful subnational political forces (e.g., warlords under Beiyang govt and CPC under Nationalist govt)
    \item Competing revolutionary visions (i.e., CPC v KMT,)
    \item External threats and influence: Japan, the Soviet Union and USA 
  \end{itemize}
  \vspace{0.3cm}
  \item Nation building as unfinished business (to this day)
  \vspace{0.1cm}
  \begin{itemize}
    \item Taiwan as a Japanese colony
    \item Peripheral areas (Mongolia, Xinjiang and Tibet/Xizang)
  \end{itemize}
\end{itemize}

# CPC v KMT

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/kmt.pdf")
```
\vspace{0.2cm}
\begin{itemize}
\small
  \item KMT started as a loose aliiance of anti-Qing rebel groups in the late 19th century; recreated in 1919 by Sun Yat-sen with the help from Moscow
  \item Chiang Kai-shek with his control over the Nationalist Army managed to become the leader of China through the "412 Incident" in 1927 and \textbf{Northern Expedition} between 1926 and 1928
  \item KMT remained fragmented despite the constant effort of state-building during the Golden Decade
\end{itemize}

# CPC v KMT

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/cpc.pdf")
```
\vspace{0.2cm}
\begin{itemize}
\small
  \item CPC started as an urban Communist party in 1921; retreated to the countryside in 1927 to rebuild its support base
  \item Mao Zedong rose to lead the Party during the \textbf{Long March} (1934-1935), by defeating the faction in favor of following the "command" from Moscow 
  \item Mao managed to build a coherent party through a serious of political movements in the 1940s
\end{itemize}

# Warloards in China (1925)

```{r echo=FALSE, out.width="82%", fig.align="center"}
knitr::include_graphics("Figs/wp_warloards.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://www.westpoint.edu/research/centers-and-institutes/digital-history-center/digital-history-center-atlases}
\end{center}

# Long March (1934-1935)

```{r echo=FALSE, out.width="82%", fig.align="center"}
knitr::include_graphics("Figs/wp_longmarch.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://www.westpoint.edu/research/centers-and-institutes/digital-history-center/digital-history-center-atlases}
\end{center}

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/cps_ma_ji.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/revs.pdf")
```

# Local states in the early 20th century (Remick 2002)

\begin{itemize}
  \item Key idea(s): The behavior and capacity of local states vary significantly in China, as these two are shaped by central policy and local social contexts (history, limitations and ideology)
  \vspace{0.3cm}
  \item Some questions to consider:
  \vspace{0.1cm}
  \begin{itemize}
    \item What is a state and state "capacity?" How did Remick define it? Why is state-building an important question? Why did she and many scholars choose to focus on taxation and public finance?
    \item Why do we need to look central and local state-building as two distinctive and yet interrelated processes?
    \item She mentioned a couple of countries to put her argument in a comparative perspective. Why? Does it make sense to you? Any other cases you can think of?
  \end{itemize}
\end{itemize}

---

\begin{center}
\textbf{Week 3: Party-state Structures}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/cover.pdf")
```

# Overview

\begin{itemize}
  \item The Chinese Party-state: Key elements and characteristics
  \vspace{0.1cm}
  \begin{itemize}
    \item The central Party leadership
    \item The central state
    \item Presence of the Party in the state
  \end{itemize}
  \vspace{0.5cm}
  \item Institutionalization of the Chinese Party-state
  \vspace{0.5cm}
  \item Chinese Party-state in comparative perspective
\end{itemize}

# The Party's Central Leadership

\begin{itemize}
\small
  \item The highest decision-making body is the \textbf{Standing Committee} of the \textbf{Politburo}, headed by the \textbf{General Secretary}, under them there are
  \vspace{0.05cm}
  \begin{itemize}
    \item Central Party Offices: General, Organization (\textit{nomenklukla}), Propaganda (Publicity), United Front and Intl Liaison 
    \item Secretariat (i.e., secretary of individual Politburo SC members)
    \item Central Military Commission
    \item Central Commission for Discipline Inspection
    \item Leading Small Groups
  \end{itemize}
  \vspace{0.3cm}
  \item Politburo members, including the General Secretary, are chosen through "internal" elections
\end{itemize}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/chart.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pmeeting.pdf")
```
\vspace{0.1cm}
\begin{center}
Politburo is the apex of the system
\end{center}

---

\begin{itemize}
  \item The delegates of the \textbf{National Party Congress}, which meets every five years, elect the \textbf{Central Committee}
  \vspace{0.5cm}
  \item The \textbf{General Secretary} is "elected" during the First Plenum (or \textbf{Plenary} session) of the \textbf{Central Committee}
  \vspace{0.5cm}
  \item The \textbf{General Secretary} was only confirmed the highest leader of China after 1992 under \textbf{Jiang Zemin}
  \vspace{0.5cm}
  \item The current post holder is \textbf{Xi Jinping}
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/selection.pdf")
```
\vspace{0.2cm}
\begin{center}
The selection process illustrates the notion of "\textbf{democratic centralism}" ("more candidates than available seats" elections)
\end{center}

# The Central State

\begin{itemize}
  \item \textbf{Legislature}: National People's Congress (NPC)
  \vspace{0.1cm}
  \begin{itemize}
    \item Meets every year in Beijing
    \item Elects the \textbf{President} and ruling cadres of other state organs every five year
    \item Operates daily through the Standing Committee and Special Committees
    \item Provides descriptive representation of different social groups and parties
  \end{itemize}
  \vspace{0.3cm}
  \item \textbf{Executive}: The State Council
  \vspace{0.1cm}
  \begin{itemize}
    \item Headed by the \textbf{Premier} (or Prime Minister), appointed by the \textbf{President} and approved by NPC
    \item Operates through Ministries and different central commissions and Leading Small Groups
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/npc_out.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/npc_2023.pdf")
```
\vspace{0.1cm}
\begin{center}
The First Meeting of the 14th National People's Congress in March 2023
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/stamp.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/map.pdf")
```

# Party-state: Key characteristics and debates

\begin{itemize}
\small
  \item Party-state helps to characterize the Party's ruling position in the Chinese political system
  \begin{itemize}
    \item \textbf{Party committee} is present in every government department and office
    \item The central Party offices "manage" and coordinate different functional government departments (e.g., Propaganda/Publicity and Ministry of Education)
    \item Leading Working Groups reported to the Party center
    \item Key Politburo members leaders also have posts in the government
    \item Key important offices are separately listed but actually the same one (e.g., Central Military Commission)
    \item Key Party and government offices work jointly despite being two different offices (e.g., Central Commission for Discipline Inspection and the Ministry of Supervision)
  \end{itemize}
  \item Scholars in China (and the Party) advocate to use the term "\textbf{Party-government}" to demonstrate the Party's extensive presence in the Chinese bureaucracy
\end{itemize}

# Institutionalization of the Party-state over generations

\begin{itemize}
  \item Political succession, developed by Deng Xiaoping and confirmed in the early 1990s
  \vspace{0.1cm}
  \begin{itemize}
    \item Term limit (except the President and Vice-President)
    \item Retirement age
    \item Successor-in-training
    \item Grandpa-designated successor
    \item Comprehensive takeover
  \vspace{0.3cm}
  \end{itemize}
  \item Collective leadership, established by Deng after the Cultural Revolution
\end{itemize}

# Defining authoritarian regimes (Geddes et al 2015)

\begin{columns}
  \column{0.5\linewidth}
  \begin{itemize}
  \small
  \item An executive achieved power through \textbf{undemocratic means} (anything other than direct, reasonably fair, competitive elections)
  \vspace{0.3cm}
  \item The government achieved power through democratic means, but subsequently \textbf{changed the formal or informal rules}
  \vspace{0.3cm}
  \item The \textbf{military} prevented parties that substantial numbers of citizens would be expected to vote for from competing and/or dictated important policy choices
  \end{itemize}
  \column{0.5\linewidth}
  \centering
  \hspace{0.2cm}
  \includegraphics[scale=0.4]{Figs/bg}
\end{columns}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/dd_gwf.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://sites.psu.edu/dictators/}
\end{center}

# Varieties of authoritarian regimes 

\begin{itemize}
  \item \textbf{Personalist}: Access to office and fruits of office depends on much more the discretion of an individual leader; not identical with \textbf{monarchy}
  \vspace{0.4cm}
  \item \textbf{Military}: A group of officers (e.g., junta) decides who will rule and exercise some influence on policy; carrying with the seeds of disintegration on their own while elites in other two types of regime have more incentives to cooperate with each other
  \vspace{0.4cm}
  \item \textbf{Single-party} (or hegemonic party): Access to political office and control over policy are dominated by one party, though other parties may legally exist and compete in elections
  \vspace{0.4cm}
  \item \textbf{Mixed} or hybrid
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/GWF/type.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://sites.psu.edu/dictators/}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/GWF/survival.pdf")
```

# What makes single-party regimes special

\begin{itemize}
  \item Key to survival is to make sure elite cohesion through rents and economic success (especially in the case of China)
  \vspace{0.3cm}
  \item Single-party regimes favor internal and societal co-optation
  \vspace{0.1cm}
  \begin{itemize}
    \item Party holds organizational capacity to control the career paths of officials, organizes the distribution of benefits to supporters, and mobilizes citizens to show support for party leaders in other ways
    \item Party cadres simply want to hold their offices and tend to stand by to observe the winners of internal struggle
    \item Party has broad basis of popular support and has the institutional and organizational sources to stay responsive and co-opt for greater participation (Ghandi 2008)
  \end{itemize}
  \vspace{0.3cm}
  \item Caveat: Parties with weak external rival and easy rent access (e.g., the Philippines) tend to be divided and collapse (Smith 2005)
\end{itemize}

# "The Institutional Turn in Comparative Authoritarianism" (Penpinsky 2014)

\begin{itemize}
  \item Why institutions matter for authoritarian survival?
  \vspace{0.1cm}
  \begin{itemize}
    \item Institutions are constraints: Elections, parties and legislatures contain factional conflicts while regulating the mobilization of mass support
    \item Institutions are dictator's instruments: They do what their creators want them to do (e.g., co-optation and power distribution) and can be adjusted
  \end{itemize}
  \vspace{0.3cm}
  \item Why institutions are important for the study of comparative authoritarianism
  \vspace{0.1cm}
  \begin{itemize}
    \item Path dependence and increasing returns: Institutions reproduce the political and social order
    \item Institutions shed light on problems of redistribution and policy making that motivate regime behavior
  \end{itemize}
\end{itemize}

# Discussion: "Party All the Time" (Thorton 2021)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.6]{Figs/thornton_book}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
    \item What are some of the key differences between being a ruling party and a revolutionary party?
    \vspace{0.3cm}
    \item How does CCP manage to succeed both as a revolutionary before 1949 and as a ruling party after the PRC was established?
    \vspace{0.3cm}
    \item Why do many authoritarian leaders rule through the party? How do they build their parties?
    \vspace{0.3cm}
  \end{itemize}
  \endcol
\endcols

# Concluding remarks: More questions to think about

\begin{itemize}
  \item Institutional durability and (policy) changes in authoritarian regimes
  \vspace{0.5cm}
  \item One-party dominant regime (e.g., Japan and Italy) and hegemonic party regime (e.g., Mexico and China)
  \vspace{0.5cm}
  \item Domestic politics and foreign policy
\end{itemize}

---

\vspace{1cm}
See you next week!

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/lee90.pdf")
```
\vspace{0.2cm}
\begin{center}
\scriptsize
\url{http://ark.cdlib.org/ark:/13030/ft9n39p3pc/}
\end{center}