---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | AI for Social Research: Qualitative
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/pandora.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/logos.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/snake_new.pdf")
```
\vspace{0.2cm}
\begin{center}
\url{https://www.aisnakeoil.com/}
\end{center}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/snake_mit.pdf")
```
\vspace{0.2cm}
\begin{center}
\url{https://www.youtube.com/watch?v=C3TqcUEFR58}
\end{center}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/tasks.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/film02.pdf")
```

# Machine learning: Supervised, unsupervised and reinforcement

\begin{itemize}
  \item Classification (or regression with regularization) -- spam emails
  \vspace{0.4cm}
  \item Pattern recognition (or clustering) -- facial recognition
  \vspace{0.4cm}
  \item Prediction (or forecasting) -- stock market forcasting and GPT (generative predictive transformer)
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/type.pdf")
```

# Example: Bots on social media

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/bots.pdf")
```
\vspace{0.2cm}
\begin{center}
\tiny
\url{https://www.socialfollowers.uk/blogs/how-social-media-bots-work-and-how-to-detect-them/}
\end{center}

# Example: Vocal pitch of women politicians

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/congress.pdf")
```
\vspace{0.2cm}
\begin{center}
\scriptsize
\url{https://doi.org/10.1017/S0003055419000467}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/congress_fig.pdf")
```

# Example: Factional politics in post-1949 China

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/china.pdf")
```

# Generative AI, GPT and large language models (LLMs)

\begin{itemize}
  \item Fine tuning and/or prompt enigeering?
  \vspace{0.5cm}
  \item General and/or domain-specific LLMs?
  \vspace{0.5cm}
  \item Metrics for performance evaluation: Intelligent? accurate? efficient (computationally)? transparent? safe? ethical? democratic? open access?
\end{itemize}

---

\begincols
  \begincol{.4\textwidth}
    \vspace{0.4cm}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/flowchart}
    \end{figure}
  \endcol
  \hspace{0.2cm}
  \begincol{.5\textwidth}
  \pause
  \vspace{0.2cm}
  \begin{itemize}
    \item Is it intuitive to understand?
    \vspace{1cm}
    \item Can you try to come up with your prompt to create a (similar) flowchart?
    \vspace{1cm}
    \item What is the prompt you use? Share it with us.
  \end{itemize}
  \endcol
\endcols

---

```{r echo=FALSE, out.width="77%", fig.align="center"}
knitr::include_graphics("Figs/deepseek.pdf")
```

# Exercise: Running critical discourse analysis with AI

\begin{itemize}
\small
  \item \textbf{Step 1}: Ask AI to produce a brief definition of critical discourse analysis -- critically evaluate the response provided by AI
  \vspace{0.1cm}
  \begin{itemize}
    \item What response do you get? Do you agree with it or not? Is it the same as what Margarita discussed last week?
    \item Can you try to engage with AI to "improve" the definition?
  \end{itemize}
  \vspace{0.3cm}
  \item \textbf{Step 2}: Upload the "Return to Work" letter and ask AI to carry out a critical discourse analysis based on the definition -- what do you get? 
  \vspace{0.3cm}
  \item \textbf{Questions to think about}
  \begin{itemize}
    \item Are you comfortable with what we have just done? Why not? What might be the caveat (Morgan 2023)?
    \item Does it make sense to analyze the "discourse" produced by AI or Human-AI interactions? How?
  \end{itemize}
\end{itemize}

# Exercise: Analyzing interview transcripts

\begin{itemize}
\small
  \item \textbf{Source}: "Urban conceptions of nature during the ongoing COVID-19 pandemic" (Matthew Wood 2021) -- see \url{https://researchdata.bbk.ac.uk/id/eprint/161/}
  \vspace{0.3cm}
  \item \textbf{Step 1}: Download the data and skim through the information on BiRD
  \vspace{0.3cm}
  \item \textbf{Step 2}: Replicate the prompts used by Morgan (2023) to analyze the interview transcripts
  \vspace{0.3cm}
  \item \textbf{Questions to think about}
  \begin{itemize}
    \item Are you happy with Morgan's prompts? Can you get around the limitations or issues pointed out by Morgan?
    \item How do you evaluate the quality of the analysis produced by AI? What might be some useful criteria?
    \item How do your know AI is more or less nuanced when it comes to coding and \textbf{interpretation}?
  \end{itemize}
\end{itemize}

# Exercise: Conducting interviews using AI agent (Geiecke and Jaravel 2024)

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/lse.pdf")
```

# Concluding remarks

\begin{itemize}
  \item Combining the use of AI with other software packages for qualitative research (e.g., ATLAS.ti and MAXQDA)?
  \vspace{0.4cm}
  \item "Strategic ignorance" is key -- the priority is to use the tool and treat LLM as a "black" box?
  \vspace{0.4cm}
  \item End of "lone wolf" social researchers? Check out the \textbf{Turing Way}
  \vspace{0.4cm}
  \item "Only because we can does not mean we should" (from Jurassic Park) -- do you agree?
  \vspace{0.4cm}
  \item Substitute versus empowerment?
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/turing.pdf")
```
\vspace{0.2cm}
\begin{center}
\scriptsize
\url{https://book.the-turing-way.org/collaboration/collaboration}
\end{center}
