---
title: |
  |
  |
  | AI for Social Research: A Primer
  |
  |
author: |
  | 2026 Masterclass in Social Research
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# From last week: "AI-4-Science"

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/ai4sci}
    \end{figure}
  \endcol
  \hspace{0.2cm}
  \begincol{.5\textwidth}
  "AI for Science (AI4S) represents the \textbf{convergence} of artificial intelligence (AI) innovation in scientific research and AI-driven scientific discovery, demonstrating their deep integration, and the establishment of a transformative research paradigm."
  \endcol
\endcols

---

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/ai4sci}
    \end{figure}
  \endcol
  \hspace{0.2cm}
  \begincol{.5\textwidth}
  \small
  "Traditional scientific discovery involves \textbf{generating and validating candidate hypotheses} from a large solution space, often characterized by low efficiency and challenges in identifying high-quality solutions. AI harnesses its \textbf{powerful data processing and analytical capabilities} to navigate solution spaces more efficiently, enabling the generation of high-quality candidate hypotheses."
  \endcol
\endcols

# Plan for today 

\begin{itemize}
  \item What makes LLMs so powerful (and intimidating)
  \vspace{0.5cm}
  \item LLMs and the social sciences
  \vspace{0.5cm}
  \item What researchers can do with AI
  \vspace{0.5cm}
  \item Discussion and looking ahead (for the next two weeks)
\end{itemize}

# Motivation and objectives

\begin{itemize}
  \item A moment of existential crisis? AI is reshaping how "knowledge" is produced
  \vspace{0.3cm}
  \item Competing narratives: hype versus skepticism -- where is the middle ground?
  \vspace{0.3cm}
  \item Social sciences face a specific challenge: AI both \textit{is} a social
  phenomenon and \textit{transforms} how we study social phenomena
  \vspace{0.3cm}
  \item Goal: develop an informed, critical, and productive relationship with (gen-)AI tools
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/openai.pdf")
```
\vspace{0.2cm}
\begin{center}
\url{https://openai.com/research}
\end{center}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/last_gen.pdf")
```
\vspace{0.3cm}
\begin{center}
\scriptsize
\url{https://datasociety.net/points/could-this-be-the-last-generation-of-graduate-students/}
\end{center}

----

\vspace{1cm}
"\textit{The comparison between graduate students and AI systems within university settings is a debate over the purpose of graduate education, which has always contained a tension between disciplinary formation and labor. Students are apprenticed into a field, but they also teach and keep research projects moving. AI systems can take over many of the visible tasks through which graduate work has long been justified}."

# What makes LLMs groundbreaking: "Autocomplete" at scale

\begin{itemize}
  \item \textbf{Statistical learning} -- LLMs learn \textbf{probability distributions} over language, predicting the next \textbf{token} given context, with no hand-crafted rules or symbolic representations
  \vspace{0.2cm}
  \item \textbf{Scaling laws} -- model capability improves \textit{predictably} with compute, parameters and data; intelligence as an emergent property of scale rather than explicit programming
  \vspace{0.2cm}
  \item A huge amount of \textbf{training data} -- models, (pre-)trained on WWW-scale corpora, absorb linguistic patterns, factual associations, and implicit cultural knowledge across domains and languages
  \vspace{0.2cm}
  \item \textbf{Hardware} -- GPU parallelism makes \textbf{billion-parameter training feasible}; compute constraints now shape the geopolitics of AI (chip export controls, sovereign AI infrastructure)
\end{itemize}

---

\vspace{0.7cm}
\begin{center}
"My \texttt{(?????)} jumped into my suitcase when I am packing and refused to leave me alone."
\end{center}
\pause
\vspace{0.1cm}
```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/cats.pdf")
```

# How do LLMs actually work? Pattern recognition and forcasting

\begin{itemize}
  \item At its core, an LLM is a \textbf{next-token predictor} -- given a sequence of words, it estimates the most probable continuation
  \vspace{0.3cm}
  \item This is neither a \textbf{retrieval} exercise (looking something up) nor an attempt of {reasoning} (following logical rules) -- it is \textbf{pattern completion learned from statistical regularities} in text
  \vspace{0.3cm}
  \item The surprising result: predicting text at sufficient scale produces behaviour that \textit{looks like} understanding, translation, coding, and argumentation -- without any of these being explicitly programmed
  \vspace{0.3cm}
  \item Implication for researchers: the model has no beliefs, intentions, or knowledge -- only \textbf{learned associations between tokens}
  \vspace{0.3cm}
  \item The \textbf{hallucination} problem: Fluent does not mean true
\end{itemize}

# How the use of AI has changed: From GPT to AGI

\begin{itemize}
  \item \textbf{Chatbot and API interface} -- "conversational" AI; direct prompt--response interaction
  \vspace{0.2cm}
  \item \textbf{Prompt engineering and context} -- structured prompting, chain-of-thought, and retrieval-augmented generation (RAG)
  \vspace{0.2cm}
  \item \textbf{Fine-tuning domain-specific models} -- adapting pretrained LLMs to specialized corpora and tasks
  \vspace{0.2cm}
  \item \textbf{Agentic AI and automated pipelines} -- multi-step reasoning, tool use, and autonomous task execution
  \vspace{0.2cm}
  \item \textbf{Agentic frameworks and harness engineering} -- orchestrating agents at scale (e.g., OpenClaw, Claude Skills)
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/llm_agent.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/logos_llm.pdf")
```

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/llm_coding.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/llm_ranking_coding.pdf")
```
\vspace{0.2cm}
\begin{center}
\url{https://openrouter.ai/}
\end{center}

# LLMs and social sciences: Three domains

\begin{itemize}
\small
  \item \textbf{Applied/professional uses} -- AI augmenting professional practice, such as \textbf{creative arts}, education (teachers and students), newsrooms, healthcare and counseling, and crime control
  \vspace{0.2cm}
  \item \textbf{Human-centered AI} and/or \textbf{human-AI interactions} -- AI as a force of structural change in human society and behaviors, such as
  \vspace{0.1cm}
  \begin{itemize}
    \item Labour market and economic development
    \item Misinformation and democratic backsliding
    \item Climate change and sustainable development
  \end{itemize}
  \vspace{0.2cm}
  \item \textbf{Geopolitics/governance of AI} -- development, regulation and governance, such as the Deepseek moment and US-China AI competition
\end{itemize}

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/anthropic.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://www.anthropic.com/news/anthropic-education-report-how-university-students-use-claude}
\end{center}

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/lse_jon.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://lse-dsi.github.io/genial/}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/lse_jon_table.pdf")
```

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/gain.pdf")
```
\begin{center}
\small
\url{https://generative-ai-newsroom.com/}
\end{center}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/deepseek_germany.pdf")
```

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/era_cover.pdf")
```

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/ada_delegation.pdf")
```
\vspace{0.1cm}
\begin{center}
\url{https://www.adalovelaceinstitute.org/report/dilemmas-of-delegation/}
\end{center}

# AI as a research assistant: Practical starting points (Hendriksen 2023)

\begin{itemize}
  \item Identifying and refining research ideas, questions, and relevant literatures (e.g., "what has been done in this area?") -- 
  \vspace{0.3cm}
  \item Locating potential co-authors and key/influential researchers in a field
  \vspace{0.3cm}
  \item Stress-testing ideas via simulated critical readers (e.g., "how will my reviewers respond?")
  \vspace{0.3cm}
  \item Drafting research applications, proposals, and code (with caution)
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ai_research.pdf")
```

# What social researchers can (should?) do with AI: Scale, scope, speed

\begin{itemize}
  \item \textbf{Simulation} -- explore human behaviour through AI "impersonation"; agent-based modeling (ABM) and human-machine comparisons
  \vspace{0.3cm}
  \item \textbf{Data analysis and generation} -- annotation, grounded-theory coding, NLP, discourse and thematic analysis (Geiecke \& Jaravel 2024; Korinek 2023;
  Leek et al.\ 2024)
  \vspace{0.3cm}
  \item \textbf{Survey and experimental research} -- design, implementation and replication of surveys and experiments (Bail 2024)
  \vspace{0.3cm}
  \item \textbf{Statistical programming} -- "vibe coding" and beyond (more in Advanced Topics in Quantitative Social Research)
\end{itemize}

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/llm_survey_cover.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/llm_survey.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/llm_prediction_cover.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/llm_prediction.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/pa_llm.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/pa_llm_critical.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ai_game.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/lse.pdf")
```

# LLMs and "interviews" (Geiecke and Jaravel 2024)

\vspace{0.3cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/lse.pdf")
```

---

\begin{itemize}
\small
  \item Use "single" LLM agent (OpenAI's GPT-4o) to carry out interviews on different topics, such as
  \vspace{0.1cm}
  \begin{itemize}
    \item Meaning of life
    \item Political preferences (over the 2024 French snap election)
    \item Educational and occupational choices
  \end{itemize}
  \vspace{0.3cm}
  \item Evaluate LLM peformance 
  \vspace{0.1cm}
  \begin{itemize}
    \item Compare AI-generated transcripts to human experts (in this case, "sociologists")
    \item Solicit assessment from respondents, e.g. how well the content of the interview captures their views (or "empathetic") and whether they would prefer to participate in an interview with AI or human researchers
  \end{itemize}
  \vspace{0.3cm}
  \item Why AI-led/assisted interviews might be a good idea? What else can AI should be used? What else can we do to improve the design? Or should we just drop the idea?
  %\begin{itemize}
  %  \item Scale and cost
  %  \item Privacy and sensitivity bias
  %\end{itemize}
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/lse_llm.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/lse_pref.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/lse_quality.pdf")
```

---



# Caveats: What can go wrong

\begin{itemize}
  \item \textbf{Human bias encoded at scale} -- but also an opportunity to reverse-engineer how biases are (re)produced
  \vspace{0.2cm}
  \item \textbf{Junk science and quality control} -- "hallucination" and confabulation; domain-specific, context engineering, RAG and evaluation frameworks as partial solutions
  \vspace{0.2cm}
  \item \textbf{Ethics as both problem and solution} -- issues of deception, consent and simulation in AI-assisted research
  \vspace{0.2cm}
  \item \textbf{Transparency and replicability} -- performance evaluation, prompt versioning, and systematic model comparison
\end{itemize}

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/snake_new.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/snake_tasks.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/film02.pdf")
```

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/zurich.pdf")
```
\vspace{0.2cm}
\begin{center}
\tiny
\url{https://www.theatlantic.com/technology/archive/2025/05/reddit-ai-persuasion-experiment-ethics/682676/}
\end{center}

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/mit_case.pdf")
```

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/mit_response.pdf")
```

# How should researchers respond?

\begin{itemize}
  \item Principle of \textbf{strategic ignorance} (McGoey 2019) -- you do not need to understand AI fully to use it wisely; know enough to ask the right questions
  \vspace{0.5cm}
  \item Identify the \textbf{skills that matter to you}: prompt/context/harness engineering, critical
  evaluation of outputs, understanding model limitations (at least at an intuitive level)
  \vspace{0.5cm}
  \item Follow \textbf{institutional guidance}: Birkbeck and LSE policies on AI use and academic integrity
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/evals.pdf")
```

---

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/ignore.pdf")
```

# Questions for discussion

\begin{itemize}
  \item Now that AI is a real possibility -- how might \textit{you} incorporate it into your research? How comfortable are you with using AI for research purposes, and why?
  \vspace{0.5cm}
  \item Christopher Bail (PNAS 2024) identifies many promising research opportunities -- which would you most like to explore? What opportunities does he miss?
  \vspace{0.5cm}
  \item What skills do you need to acquire to take genuine advantage of AI? How much do you actually need to \textit{know} about how AI works?
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pandora.pdf")
```

# Training data: The world encoded as text

\begin{itemize}
  \item LLMs are trained on internet-scale corpora: web pages, books, code, academic papers, social media, legal documents
  \vspace{0.3cm}
  \item The model does not \textit{store} this text -- it \textbf{compresses} it into billions of numerical parameters (weights) that encode statistical relationships between concepts
  \vspace{0.3cm}
  \item What the model "knows" reflects what was \textit{written down, in languages represented online, by people with internet access} -- a deeply non-representative sample of human knowledge
  \vspace{0.3cm}
  \item \textbf{Knowledge cutoff}: the model has no awareness of events after its training data ends -- it cannot update in real time (without retrieval augmentation)
\end{itemize}

# Scaling laws and emergent behaviour

\begin{itemize}
  \item Kaplan et al.\ (2020) showed that LLM performance improves \textit{predictably} with three factors: model size (parameters), training data volume, and compute
  \vspace{0.3cm}
  \item More surprising: certain capabilities -- multi-step reasoning, language translation, analogical thinking -- \textbf{emerge abruptly} at scale thresholds, without being explicitly trained
  \vspace{0.3cm}
  \item This makes frontier LLMs qualitatively different from earlier ML tools: their capabilities are \textbf{difficult to anticipate in advance} and difficult to fully audit after the fact
  \vspace{0.3cm}
  \item For social researchers: you are working with a tool whose full capability envelope is not yet known -- even to its developers
\end{itemize}

# Online short course: "AI for Social Research" (in planning)

\begin{itemize}
\small
  \item \textbf{Session 1}: Overview -- From GPT to AGI
  \vspace{0.1cm}
  \item \textbf{Session 2}: LLMs and qualitative social research
  \vspace{0.1cm}
  \item \textbf{Session 3}: LLMs for survey research
  \begin{itemize}
    \item questionnaire design and pretesting
    \item synthetic respondents and experiments
    \item replication and validation
  \end{itemize}
  \vspace{0.1cm}
  \item \textbf{Sessions 4-6}: LLMs in social data science
  \begin{itemize}
    \item natural language processing
    \item agent-based modeling and simulations
    \item agentic data analytical pipelines
  \end{itemize}
  \vspace{0.1cm}
  \item \textbf{Session 7}: Project workshop and looking ahead
\end{itemize}

# Appendix: GenAI for plausibility checks

\begin{itemize}
  \item Without using any conversational AI tools (e.g., ChatGPT, Claude and Gemini), imagine how you can integrate AI into your research practice and workflows
  \vspace{0.5cm}
  \item Ask the tools you prefer to see what they can say/do for your research (both in terms of the topic and methods)
  \vspace{0.5cm}
  \item Do these suggestions make sense to you? Any notable differences in terms of the quality of generated responses?
\end{itemize}
