---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 3: Codifying}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week3.pdf")
```

# Today's plan

\begin{itemize}
  \item Logistics: Seminar log and mini critical review (20\% each)
  \vspace{0.25cm}
  \item Recap: "Nations, States and Nation-States" (40 mins)
  \vspace{0.25cm}
  \item Refresher lecture (1-1.5 hr)
  \vspace{0.1cm}
  \begin{itemize}
    \item Comparing constitutions
    \item Judiciary in comparative perspective
  \end{itemize}
  \vspace{0.25cm}
  \item Reading exercise: "Constitutions and the Quality of Democracy" (next week)
\end{itemize}

# Seminar log

\begin{itemize}
\small
  \item Due on \textbf{February 13}, 2026 (Friday)
  \vspace{0.2cm}
  \item Write a \textbf{500-word} piece of connected prose (max)
  \vspace{0.2cm}
  \item Choose one topic from Week 1-4 and answer the questions listed at the end of the lecture outline
  \vspace{0.2cm}
  \item Review the readings (both requred and optional ones), the recorded lecture and the refresher lecture in class
  \vspace{0.2cm}
  \item Your response should
  \vspace{0.1cm}
  \begin{itemize}
    \item provide an overview of what you have learned and some of the key points you will take with you
    \item provide thoughts from class discussion to demonstrate your reflection and understanding of the topics/issues, with specific examples
  \end{itemize}
\end{itemize}

# Mini critical review

\begin{itemize}
\small
  \item Due on \textbf{February 20}, 2026 (Friday)
  \vspace{0.2cm}
  \item Write a \textbf{500-word} piece of connected prose (max)
  \vspace{0.2cm}
  \item Choose one topic from Week 1-5
  \vspace{0.2cm}
  \item In your review:
  \vspace{0.1cm}
  \begin{itemize}
    \item discuss the reading exercise to review relevant topics of the week, based on the questions we discuss in class
    \item review the answers you heard in class discussion and provide your thoughts
  \end{itemize} 
\end{itemize}

# Reminders and tips

\begin{itemize}
  \item No need to be thorough -- perhaps better to be focused and in-depth
  \vspace{0.3cm}
  \item No need to worry too much about citations for these two assignments -- the key is to share your thoughts as informed by readings, lectures and class discussions
  \vspace{0.3cm}
  \item Balanace between descriptive versus analytical 
  \vspace{0.3cm}
  \item Come to office hours for verbal support
\end{itemize}

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/ga.pdf")
```
\begin{center}
\scriptsize
\url{https://goodauthority.org/}
\end{center}

# Exercise: "Democracy and Nationalism" (Eun A. Jo, College of William and Mary)

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/reading.pdf")
```
\begin{center}
\scriptsize
\url{https://goodauthority.org/news/good-to-know-nationalism-and-democracy/}
\end{center}

---

\vspace{0.3cm}
\begin{itemize}
  \item Is nationalism good or bad for democracy? Why?
  \vspace{0.3cm}
  \item Does the "typology" make sense to you? Can you place countries you know in the table? Any exceptions?
  \vspace{0.3cm}
  \item What activities can contribute to the construction of shared identity, namely "nation-building?"
  \vspace{0.3cm}
  \item The article concludes by saying nationalism needs a better "narrative." What does that mean?
\end{itemize}

# Recap: Nations, states and nation-states

\begin{itemize}
\small
  \item Nations
  \begin{itemize}
    \item "Imagined Communities" (Benedict Anderson 1983)
    \item Top-down and/or bottom-up construction of shared identity
  \end{itemize}
  \vspace{0.1cm}
  \item States
  \begin{itemize}
    \item Political autorities/entities with the legitimacy to use coercive power (Charles Tilly 1990)
    \item Centralizred armed forces (military and police), bureaucracy, and "infrastructure" (e.g., transportation)
    \item Taking various forms in history and contemporary era
  \end{itemize}
  \vspace{0.1cm}
  \item Nations and states do not always align, and when they do not
  \begin{itemize}
    \item Nations without states: armed conflicts and secession/independent movement; "Wars in the Balkans" in the 1990s
    \item Multi-ethnic/nation states: plural societies as put by Arend Lijphart 1977; Northern Ireland after the 1997 Good Friday Agreement
  \end{itemize}
\end{itemize}

# Refresher lecture

\begin{itemize}
  \item Comparing constitutions
  \vspace{0.1cm}
  \begin{itemize}
    \item For legal scholars
    \item For political scientists
  \end{itemize}
  \vspace{0.3cm}
  \item Judiciary in comparative perspective
  \vspace{0.1cm}
  \begin{itemize}
    \item Constitutional review
    \item Emergency power
  \end{itemize}
  \vspace{0.3cm}
  \item Looking forward: Two visions of democracy
\end{itemize}

# Comparing constitutions like a legal scholar

\begin{itemize}
  \item Ideology (e.g., the Federalist Papers)
  \vspace{0.2cm}
  \item Formats: Cofied versus uncodified? Written versus unwritten (norms)?
  \vspace{0.2cm}
  \item Texts
  \vspace{0.1cm}
  \begin{itemize}
    \item Length: Lengthy or succinnt?
    \item Substance: Principles or details?
    \item Wording: Simple or sophisticated? 
  \end{itemize}
  \vspace{0.2cm}
  \item Rules for changes: Hard/strong versus soft/weak? How difficult it is to change some parts of or the whole consttitution? (see Abramson and Barber 2019)
  \vspace{0.2cm}
  \item Key inquries: Coming up with a "perfect" constitutional law
\end{itemize}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/federalist.pdf")
```
\begin{center}
\scriptsize
\url{https://guides.loc.gov/federalist-papers/full-text}
\end{center}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/philly.pdf")
```

# "Comparative Constitutions Project"

\vspace{0.1cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/ccp.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
\url{https://comparativeconstitutionsproject.org}
\end{center}

# "Comparative Constitutions Project"

\vspace{0.1cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/ccp_profs.pdf")
```

---

\vspace{0.1cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/ccp_ranking.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://comparativeconstitutionsproject.org/ccp-rankings/}
\end{center}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/george2.pdf")
```
\vspace{0.2cm}
\begin{center}
\tiny
\url{https://constitution-unit.com/2016/06/06/countries-with-longer-constitutions-are-poorer-and-more-corrupt-evidence-from-the-oecd/}
\end{center}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/ucl_con.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
\url{https://youtu.be/dcdBnn2NXmM}
\end{center}

# Comparing constitutions like a political scientist

\begin{itemize}
\small
  \item Varieties of constitutional democracy
  \vspace{0.1cm}
  \begin{itemize}
    \item Executive system: presidential versus parliamentary
    \item Electoral system: majoritarian/plurality, PR and mixed-member (and the resulting party system)
    \item Legislative system: uni- or bi-cameral
    \item Central-local relations: federal, unitary and decentralized
  \end{itemize}
  \vspace{0.2cm}
  \item Key inquries: Underlying power dynamics driving the "creation" as well as the "practice"
  \vspace{0.1cm}
  \begin{itemize}
    \item Power concentration and dispersion (Lijphart 1999; Powell 2000), and the resulting design of judiciary
    \item Rigid or adapatative constitutions (Levitsky and Murillo 2009)
  \end{itemize}
  \vspace{0.1cm}
  \item Two visions of democracy: Westminster Model versus Concensus Model as a response to "Madison's dilemma"
\end{itemize}

# Constitutional "engineering" in plural socieities (Lijphart 1977)

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/lijphart.pdf")
```

# Consociational democracy: Power-sharing among groups

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ireland.pdf")
```

# Supreme courts and constitutional review

\begin{itemize}
  \item Judiciary in the constitution organizes relations between the executive and legislative branches of government:
  \vspace{0.2cm}
  \begin{itemize}
    \item \textbf{Parliamentary supremacy}: a principle according to which judges' decisions remain subordinate to decisions of the legislative majority (e.g., UK before 2009)
    \item \textbf{Judicial review}: the ability of a country's high court to invalidate laws the legislature has enacted by declaring them unconstitutional (e.g., USA, Germany and UK after 2009)
  \end{itemize}
  \vspace{0.3cm}
  \item What explans the adoption of either principle? What is the "politics" behind the operation of these principles?
\end{itemize}

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/usa.pdf")
```

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/germany.pdf")
```

---

\vspace{1.5cm}
"Domestic intelligence cannot push for a ban on the party -- that can only go through the \textbf{two houses of parliament}, \textbf{government} or \textbf{the constitutional court} -- but its latest decision could encourage others to start the process."

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/uk.pdf")
```

# "Why Govts Call A State of Emergency" (Bjørnskov and Voigt 2018)

\begin{itemize}
  \item Countries w/o constitutionalized emergency provisions declare states of emergency (SoE) significantly more often than those that have them
  \vspace{0.2cm}
  \item When the cost to declare SoE is low, governments will do so more frequently for natural disasters (but not for political "turmoils")
  \vspace{0.2cm}
  \item Large natural disasters are more likely to lead to an official SOE when the legislature holds more power
  \vspace{0.2cm}
  \item Takeaway lessons
  \vspace{0.1cm}
  \begin{itemize}
    \item Constitutioal provisions (as institutional checks) matter
    \item Not all SoEs are idential
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/soe_india.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://www.thehindu.com/books/a-list-of-books-documenting-emergency-in-india-its-excesses-and-consequences/article69734711.ece}
\end{center}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/soe_usa.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://www.whitehouse.gov/presidential-actions/2025/08/declaring-a-crime-emergency-in-the-district-of-columbia/}
\end{center}

# Case study: (Absence) of SoE declaration for the pandemic in the UK

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/soe_uk.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://constitution-unit.com/2016/06/06/countries-with-longer-constitutions-are-poorer-and-more-corrupt-evidence-from-the-oecd/}
\end{center}