---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 3: Codifying}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week3.pdf")
```

# Today's plan

\begin{itemize}
  \item Logistics: Seminar log and mini critical review (20\% each)
  \vspace{0.25cm}
  \item Recap: "Nations, States and Nation-States" (40 mins)
  \vspace{0.25cm}
  \item Refresher lecture (1-1.5 hr)
  \vspace{0.1cm}
  \begin{itemize}
    \item Comparing constitutions
    \item Judiciary in comparative perspective
  \end{itemize}
  \vspace{0.25cm}
  \item Reading exercise: "Constitutions and the Quality of Democracy" (next week)
\end{itemize}

# Seminar log

\begin{itemize}
\small
  \item Due on \textbf{February 13}, 2026 (Friday)
  \vspace{0.2cm}
  \item Write a \textbf{500-word} piece of connected prose (max)
  \vspace{0.2cm}
  \item Choose one topic from Week 1-4 and answer the questions listed at the end of the lecture outline
  \vspace{0.2cm}
  \item Review the readings (both requred and optional ones), the recorded lecture and the refresher lecture in class
  \vspace{0.2cm}
  \item Your response should
  \vspace{0.1cm}
  \begin{itemize}
    \item provide an overview of what you have learned and some of the key points you will take with you
    \item provide thoughts from class discussion to demonstrate your reflection and understanding of the topics/issues, with specific examples
  \end{itemize}
\end{itemize}

# Mini critical review

\begin{itemize}
\small
  \item Due on \textbf{February 20}, 2026 (Friday)
  \vspace{0.2cm}
  \item Write a \textbf{500-word} piece of connected prose (max)
  \vspace{0.2cm}
  \item Choose one topic from Week 1-5
  \vspace{0.2cm}
  \item In your review:
  \vspace{0.1cm}
  \begin{itemize}
    \item discuss the reading exercise to review relevant topics of the week, based on the questions we discuss in class
    \item review the answers you heard in class discussion and provide your thoughts
  \end{itemize} 
\end{itemize}

---

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/ga.pdf")
```
\begin{center}
\scriptsize
\url{https://goodauthority.org/}
\end{center}

# Exercise: "Democracy and Nationalism" (Eun A. Jo, College of William and Mary)

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/reading.pdf")
```
\begin{center}
\scriptsize
\url{https://goodauthority.org/news/good-to-know-nationalism-and-democracy/}
\end{center}

---

\vspace{0.3cm}
\begin{itemize}
  \item Is nationalism good or bad for democracy? Why?
  \vspace{0.3cm}
  \item Does the "typology" make sense to you? Can you place countries you know in the table? Any exceptions?
  \vspace{0.3cm}
  \item What activities can contribute to the construction of shared identity, namely "nation-building?"
  \vspace{0.3cm}
  \item The article concludes by saying nationalism needs a better "narrative." What does that mean?
\end{itemize}

# Recap: Nations, states and nation-states

\begin{itemize}
\small
  \item Nations
  \begin{itemize}
    \item "Imagined Communities" (see Benedict Anderson 1983)
    \item Top-down and/or bottom-up construction of shared identity
  \end{itemize}
  \vspace{0.1cm}
  \item States
  \begin{itemize}
    \item Political autorities/entities with the legitimacy to use coercive power (see Charles Tilly 1990)
    \item Centralizred armed forces (military and police), bureaucracy and infrastructure (e.g., transportation)
    \item Taking various forms in history and contemporary era
  \end{itemize}
  \vspace{0.1cm}
  \item Nations and states do not always align, and when they do not
  \begin{itemize}
    \item Intra-state or civil wars and secession/independent movement; Wars in the Balkans in the 1990s
    \item Multi-ethnic/nation states (or plural societies as put by Arend Lijphart 1977); Northern Ireland after the 1997 Good Friday Agreement
  \end{itemize}
\end{itemize}

# Refresher lecture

\begin{itemize}
  \item Comparing constitutions
  \vspace{0.1cm}
  \begin{itemize}
    \item For legal scholars
    \item For political scientists
  \end{itemize}
  \vspace{0.3cm}
  \item Judiciary in comparative perspective
  \vspace{0.1cm}
  \begin{itemize}
    \item Constitutional review
    \item Emergency power
  \end{itemize}
  \vspace{0.3cm}
  \item Looking forward: Two visions of democracy
\end{itemize}

# Comparing constitutions like a legal scholar

\begin{itemize}
  \item Formats: Cofied versus uncodified? Written versus unwritten?
  \vspace{0.2cm}
  \item Texts: Lengthy versus succinnt? Principles or details? Wording of articles? 
  \vspace{0.2cm}
  \item Rules to amend and change: How difficult it is to change some parts of or the whole consttitution?
\end{itemize}

# Debate on UK "constitution"

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucl_con.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
\url{https://youtu.be/dcdBnn2NXmM}
\end{center}

# "Constitute Project"

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/con_project.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
\url{https://www.constituteproject.org/}
\end{center}

# Comparing constitutions like a political scientist

\begin{itemize}
  \item Varieties of constitutions
  \vspace{0.1cm}
  \item Key inquries: Underlying power dynamics driving the "creation" as well as the "practice"
  \begin{itemize}
    \item Power concentration and dispersion (Lijphart 1999; Powell 2000)
    \item Rigid or adapatative constitutions (Levitsky and Murillo 2009; Abramson and Barber 2019)
  \end{itemize}
\end{itemize}

# Constitutional "engineering" in plural socieities (Lijphart 1977)

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/lijphart.pdf")
```

# Supreme courts and constitutional review

Basic logic.

---

USA

---

Germany

---

UK.

# Emergency power.

Collapse of Weimar democracy (check Sherri Berman).
