---
title: |
  | 
  |
  | Doing Political and Social Research 
  |
  |
author: |
  | Week 4
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucl.pdf")
```
\vspace{0.2cm}
\begin{center}
\scriptsize
\url{https://www.eventbrite.co.uk/cc/policy-and-practice-seminar-series-1627679}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/barbara.pdf")
```

---

```{r echo=FALSE, out.width="110%", fig.align="center"}
knitr::include_graphics("Figs/barbara2.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ibp.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/annual.pdf")
```
\vspace{0.2cm}
\begin{center}
\scriptsize
\url{https://www.annualreviews.org/content/journals/polisci}
\end{center}

# Find your research questions

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/mason.pdf")
```
\begin{center}
\begin{scriptsize}
Mason, J. 2002. \textit{Qualitative researching} (2nd). London: Sage, p.18.
\end{scriptsize}
\end{center}

# Identify the gap

\begin{itemize}
  \item Theoretical/empirical tensions or contradictions
  \vspace{0.2cm}
  \item Different or alternative theoretical perspectives
  \vspace{0.2cm}
  \item Different methodological approach (e.g., build on more inductive work to propose some hypothesis-testing deductive approach)
  \vspace{0.2cm}
  \item Extension to new context (e.g., different region, time period)
  \vspace{0.2cm}
  \item Some conventional wisdom left untested or understudied (e.g., critical element not engaged with, need to consider other conditions)
  \vspace{0.2cm}
  \item ...
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/george.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/auto.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/systematic.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/king.pdf")
```