---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | Featuring Research on Social/New Media
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/plats.pdf")
```

# Podcast for different subject areas

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/blogs_crop.pdf")
```

# Recommended podcast: New Books Network

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.5]{Figs/nbn_crop}
    \end{figure}
    \vspace{0.2cm}
	\begin{center}
	\textbf{\url{newbooksnetwork.com}}
	\end{center}
  \endcol
  \begincol{.5\textwidth}
  	\begin{itemize}
	\item A consortium of author-interview podcast channels:
	\vspace{0.3cm}
	\begin{itemize}
		\item Aiming to "raise the level of public discourse" by introducing scholars and other serious writers to a wide public via new media
		\vspace{0.3cm}
		\item Convering $100+$ subjects, disciplines, and genres
		\vspace{0.3cm}
		\item Publishing 50 to 75 episodes every week
		\vspace{0.3cm}
		\item Opening to suggestions or pitches
	\end{itemize}
	\end{itemize}
  \endcol
\endcols

# Short videos

\vspace{0.3cm}
```{r echo=FALSE, out.width="110%", fig.align="center"}
knitr::include_graphics("Figs/bbk3mt1_crop.pdf")
```
\vspace{0.5cm}
\begin{scriptsize}
\begin{center}
\textbf{\url{https://youtube.com/playlist?list=PL2Fy-5oxIlb7VLoWUC6j_eq9GJXzSjXfP}}
\end{center}
\end{scriptsize}

# Twitter threads

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/paul.pdf")
```
\vspace{0.1cm}
\begin{scriptsize}
\begin{center}
\textbf{\url{https://twitter.com/ProfPaulPoast/status/1630305966674640896?s=20}}
\end{center}
\end{scriptsize}

# Twitter threads

\begin{itemize}
	\item \textbf{[Optional]}: Follow Chao-yo (\texttt{@londonmeow0627}) on Twitter
	\vspace{0.3cm}
	\item Start with an introduction for spoilers and/or highlights, put concise statements about the research idea (RQ, literature and context/background) and a quick discussion on methods (if the methods are innovative) and findings
	\vspace{0.3cm}
	\item Some useful tips:
	\vspace{0.2cm}
	\begin{itemize}
		\item Acknowledge those who inspire your research; give credits to your collaborators
		\item Include GIFs/memes or short audios/videos to attract attention
		\item Engage with the replies constructively to boost engagement
		\item Use Twitter Analytics and other tools (e.g., TweetDeck) to manage tweets
	\end{itemize}
\end{itemize}

# Blog posts

\begin{small}
\begin{itemize}
	\item \textbf{REF} (Research Excellence Framework) encourages researchers to use blogs to communicate their research with the public
	\vspace{0.2cm}
	\item Some key principles:
	\begin{itemize}
		\item Use titles for spoilers
		\item Decide how you want to hook/motivate readers -- narrate in a reflective manner (it can be reasonably personal)
		\item Present the core of the topics/projects clearly and simply -- use subject-specific vocabulary wisely
		\item Link additional materials if necessary (not required for this assessment)
		\item Build your research profile -- organizational position (within and outside Birkbeck), links to social media profiles (e.g., Linkedin), and published works and/or any other research activities (e.g., NGOs or think tanks)
	\end{itemize}
	\vspace{0.2cm}
	\item What blogs are available at Birkbeck: \url{https://www.bbk.ac.uk/blogs}
\end{itemize}
\end{small}

# Blog posts

\begin{itemize}
	\item For Masterclass, you will have to write two 1,000-word (max) \textbf{blog posts} to reflect on two lectures.
	\vspace{0.3cm}
	\begin{itemize}
		\item One should be chosen from Weeks 2-5 (qualitative and historical social research)
		\vspace{0.1cm}
		\item One should be chosen from Weeks 7-11 (quantitative and social data science)
	\end{itemize}
\end{itemize}

```{=tex}
%\vspace{0.5cm}
%\item In the \textbf{critical reflection paper} (4,000 words max), you should also provide a more general discussion on how the lecture series can help with your research projects (e.g., dissertation) and/or ideas.
```

# Exercise: Composing a blog post on today's lecture

\begin{itemize}
	\item First, work on the questions by yourself (\textbf{15-20 mins})
	\vspace{1cm}
	\item Next, exchange your answers with your group partners (2-3 students) for feedback (\textbf{15 mins}).
	\vspace{1cm}
	\item Finally, share what you have with the class (\textbf{20-30 mins})
\end{itemize}

# Exercise: Composing a blog post on today's lecture

\begin{itemize}
	\item Write about 500 words (max) to answer the following questions.
	\begin{itemize}
		\item Summarize the key highlights of today's lecture to those who did not attend the lecture
		\item Discuss what have you learned from the lecture, in relation to any relevant research experiences in the past
			\begin{itemize}
				\item If you have done surveys before, did today's lecture provide any new insights?
				\item If you have not done any surveys before, what are some of the most useful things you learned from today's lecture?
		\end{itemize}
		\item How can today's lecture help with your research, even if you may not be thinking about using surveys at this moment?
		\item What else would you like to learn about surveys?
		\item Provide a short description about yourself as a researcher
	\end{itemize}
\end{itemize}