---
title: |
  | 
  |
  | CP$^2$ Week 1: State of the Field 
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Welcome to CP$^2$

\begin{itemize}
  \item Logistics and introduction
  \vspace{0.1cm}
    \begin{itemize}
    \item Module organization
    \item Weekly syllabus
    \item Assessment
    \item Communication
\end{itemize}
  \vspace{0.5cm}
  \item Lecture and discussion: "China and comparative politics"
\end{itemize}

# Module organization

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/bbk_night.pdf")
```
\begin{itemize}
  \item Every Thursday 6-8:30pm
  \item Rm 353, Malet Street
  \item In-person lecture and seminar discussion
  \item \textbf{Question}: Final week in Chinatown?
\end{itemize}

# Weekly syllbus

\begin{itemize}
\small
  \item \textbf{Week 1}: State of field
  \vspace{0.1cm}
  \item \textbf{Weeks 2-5}: The Party-state
  \begin{itemize}
    \item From empire to nation-state
    \item Structures of the party-state
    \item Informal politics and factionalism
    \item Central-local relations and local governance
  \end{itemize}
  \vspace{0.1cm}
  \item \textbf{Weeks 7-10}: Governance and development
  \begin{itemize}
    \item Coercion and control
    \item Co-optation and representation
    \item Social protection and redistribution
    \item State-led development (Guest Speaker: Dr Xin Han)
  \end{itemize}
  \vspace{0.1cm}
  \item \textbf{Week 11}: China as a method
\end{itemize}

# Assessment

\begin{itemize}
\small
  \item Learning journal (30\%)
  \vspace{0.1cm}
  \begin{itemize}
    \item Compete a 1,000-word essay (max) about a published peer-reviewed political science article
    \item Answer two questions to explore how China is being examined in comparative perspective
    \item More details coming in Week 2
  \end{itemize}
  \vspace{0.5cm}
  \item Final essay (70\%)
  \begin{itemize}
    \item Compete a 3,000 essay about a substantive question in relation to Chinese politics, using all provided materials through the module
    \item Key is to demonstrate a solid understanding of class materials
    \item More details coming in Week 6
  \end{itemize}
\end{itemize}

# Moodle and student support

\begin{itemize}
  \item Moodle
  \vspace{0.1cm}
  \begin{itemize}
    \item Background readings
    \item Additional resources
    \item Weekly videos and readings
    \item Lecture slide decks
  \end{itemize}
  \vspace{0.2cm}
  \item Office hours: Every Friday 3-5pm via Teams (or by appointment)
  \vspace{0.2cm}
  \item Send administrative inquries to the School via Ask
  \vspace{0.2cm}
  \item Study support services provided by the Faculty and College
\end{itemize}

# Ground rules

\begin{itemize}
  \item In-person attendance is compulsory
  \vspace{0.2cm}
  \item Students should finish all readings and be prepared to discuss them
  \vspace{0.2cm}
  \item Take advantage of small-class size to build a fruitful working relationship with the module convenor
  \vspace{0.2cm}
  \item Consult module convenor or the Study Skills team to learn about strategic academic reading
  \vspace{0.2cm}
  \item Self-care is essential and stay close communication with module convenor and the College
\end{itemize}

---

\vspace{1cm}
Any questions before we proceed?

# Introduction

\begin{itemize}
  \item About Chao-yo
  \vspace{0.8cm}
  \item About you: Please tell us your name, program/year, any relevant modules in Politics or other subject areas you have done, and what you would like to learn from this module
\end{itemize}

---

\vspace{1cm}
Again, any questions before we proceed?

---

\begin{center}
\textbf{Week 1: State of the Field}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/theme.pdf")
```

# Three generations of China scholars during the Cold War (Harding 1984)

\begin{itemize}
  \item 1st generation (1960s)
  \vspace{0.1cm}
  \begin{itemize}
    \item Focus on descriptive study and rely heavily on (limited) official data, such as Communist doctrine, policy, and organizational tables; many scholars completed their work in Taiwan 
    \item Research is mostly speculative and tentative (and yet suffering from over-generalization and under-conceptualization)
  \end{itemize}
  \vspace{0.5cm}
  \item 2nd generation (1970s)
  \vspace{0.5cm}
  \item 3rd generation (1980s)
\end{itemize}

# Three generations of China scholars during the Cold War (Harding 1984)

\begin{itemize}
  \item 1st generation (1960s)
  \vspace{0.5cm}
  \item 2nd generation (1970s)
  \vspace{0.1cm}
  \begin{itemize}
    \item Aim to be conceptual and rely heavily on Red Guard materials to get behind-the-scene insights
    \item Start to disaggregate China by moving to subnational comparison
  \end{itemize}
  \vspace{0.5cm}
  \item 3rd generation (1980s)
\end{itemize}

# Three generations of China scholars during the Cold War (Harding 1984)

\begin{itemize}
  \item 1st generation (1960s)
  \vspace{0.5cm}
  \item 2nd generation (1970s)
  \vspace{0.5cm}
  \item 3rd generation (1980s)
  \vspace{0.1cm}
  \begin{itemize}
    \item Engage more with comparative politics literature while utilizing the accumulation of empirical description of the Chinese regimes contributed by China scholars from previous generations
    \item China's Reform and Openness provides new materials and research opportunities
  \end{itemize}
\end{itemize}

# New generations of China scholars after the Cold War (Shambaugh 2023)

\begin{itemize}
  \item The first three generations of China scholars more or less "completed" the field transition from area studies to political "science"
  \vspace{0.1cm}
  \item China has also been used more generally to tackle broad theoretical puzzles in comparative politics and political economy
  \begin{itemize}
    \item The understanding of Chinese politics has in general moved from "totalitarianism" to "authoritarianism"
    \item Modern China studies have gradually moved away from the typical "cultural trap"
  \end{itemize}
  \vspace{0.1cm}
  \item Scholars have also been trained to use cutting-edge methods and rigorous research design
  \vspace{0.1cm}
  \item However, two recent events have changed the research landscape: One is the abolition of the term limits for the Chinese president; one is the escalation of the geopolitical tension between the US and China; the consequences are profound
\end{itemize}

# "Bringing in China" (Tsai 2017)

\begin{itemize}
  \item Many political scientists have under-utilized or ignored the theoretical relevance of the study of Chinese politics
  \vspace{0.2cm}
  \item And yet China can contribute to the general political science research in many different ways
  \vspace{.01cm}
  \begin{itemize}
    \item Using China to generate new research questions (by specifying the scope condition or conceptual equivalence)
    \item Using China to test the the scope of existing findings based on another country (e.g., the United States)
    \item Using China to demonstrate innovative methods of data collection combined with strong research design (e.g., King, Pan and Roberts 2013)
  \end{itemize}
  \vspace{0.2cm}
  \item There exist many opportunities of theory building, in particular for the literatures on authoritarian governance, political development, and political economy of development
\end{itemize}

# "Bringing in China" (Tsai 2017)

\begin{itemize}
  \item Many political scientists have under-utilized or ignored the theoretical relevance of the study of Chinese politics
  \vspace{0.2cm}
  \item And yet China can contribute to the general political science research in many different ways
  \vspace{0.2cm}
  \item There exist many opportunities of theory building, in particular for the literatures on authoritarian governance, political development, and political economy of development
  \vspace{0.1cm}
  \begin{itemize}
    \item Tsai seems to find China studies that can produce new research agenda more valuable, implicitly speaking. What do you think? 
    \item Which topic do you find most intruging? And are any of these questions relevant to the country you are most familiar with (e.g., United Kingdom)?
    \item Why does the question of "bringing in China into comparative politics" matter? What do you think? 
  \end{itemize}
\end{itemize}

---

\begin{itemize}
  \item Authoritarian political development
  \vspace{0.1cm}
  \begin{itemize}
    \item Quasi-democratic institutions in China (e.g., political parties, elections and legislatures)
    \item Role of media and the Internet; censorship and propaganda
    \item Legacies of revolution and colonialism for state-building
  \end{itemize}
  \vspace{0.4cm}
  \item Political behavior of (economic) development
  \vspace{0.1cm}
  \begin{itemize}
    \item "Participatory" development and civic engagement and non-electoral participation
    \item Strategic use of the law and formal institutions
    \item Intermediaries and informal institutions  
  \end{itemize}
\end{itemize}

---

\begin{itemize}
  \item "Governance on the Cheap" -- how do political leaders extend the reach of the state over local officials and citizens without formal bureaucratic institutions?
  \vspace{0.6cm}
  \item "Voluntary compliance"
  \vspace{0.6cm}
  \item "Institutional theories of political behavior"
\end{itemize}

# Discussion: "Why China's Reforms Have Delayed Democracy" (Gallagher 2002)

\begin{itemize}
  \item Which general theoretical puzzle did Gallagher attempt to speak to? (Hint: It is about Lipset)
  \vspace{0.3cm}
  \item Why is the case of "Reform and Openness" in China interesting or puzzling?
  \vspace{0.3cm}
  \item How did Gallagher dismiss conventional wisdom about the political implications of FDI liberalization and market reform? What is her alternative theory? How did her theory speak to the general theory of economic modernization and regime change?
  \vspace{0.3cm}
  \item Did Gallagher highlight the importance of culture in her argument? If not, why do you think that is the case? Do you agree with her?
\end{itemize}

---

\vspace{1cm}
See you next week!
