---
title: |
  | 
  |
  | CP$^2$ Week 7: Control and Coercion
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# 2025 "Two-Sessions"

```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/two2.pdf")
```
\vspace{0.1cm}
\begin{center}
Currently taking place in Beijing
\end{center}

---

\begin{itemize}
\small
  \item National People's Congress, March 5-14
  \vspace{0.1cm}
  \begin{itemize}
    \item Review the \textbf{Government Work Report} (政府工作报告)
    \item Examine the report on the implementation of the \textbf{2024 National Economic and Social Development Plan} and the 2025 report and draft
    \item Examine the report on the execution of the 2024 \textbf{Central and Local Budgets} and the 2025 report and draft
    \item Review the proposal "\textbf{Draft Amendment to the Law of the People's Republic of China on Deputies to the National and Local People's Congresses}" submitted by the NPC Standing Committee (中华人民共和国全国人民代表大会和地方各级人民代表大会代表法（修正草案）)  
    \item Review the work report of the \textbf{Standing Committee of the National People's Congress}, the \textbf{Supreme People's Court} and the \textbf{Supreme People's Procuratorate} 
  \end{itemize}
  \vspace{0.3cm}
  \item National Chinese People's Political Consultative Conference, March 4-10
\end{itemize}

# Plan for Weeks 7-8

\begin{itemize} 
  \item Assessment: Reminders and Q\&A
  \vspace{0.4cm}
  \item Lectures: Control and co-optation
  \vspace{0.1cm}
  \begin{itemize}
    \item Puzzle: "The dictator's dilemma" (or is there one?)
    \item Coercive control as the means to prevent elite split and societal collective action
    \item Co-optation and representation (and yet without electoral accountability) as the means to stay responsive and give the mass a stake in the regime's survival 
  \end{itemize}
\end{itemize}

# End-of-term bobba tea meet-up

\vspace{0.4cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bobba2.pdf")
```
\vspace{0.2cm}
\begin{center}
In Week 10, we can meet up at a bobba place nearby and walk to the classroom together -- thoughts?
\end{center}

# "The dictator's dilemma" (Wintrobe 1998)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/wintrobe}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \small
  \begin{itemize}
  \small
  \item Contrary to the popular belief, the use of repression can be counter-productive for dictators, usually dictatorship operates through political exchange (co-optation)
  \vspace{0.4cm}
  \item The dictators seek the loyal support of interest groups, especially among the military and other political and social forces having the strongest capacity to depose them
\end{itemize}
  \endcol
\endcols

# "The dictator's dilemma" (Wintrobe 1998)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/wintrobe}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \small
  \begin{itemize}
  \small
  \item Co-optation can be done by redistribution and incorporating groups into the regime apparatus
  \vspace{0.5cm}
  \item democracies do not necessarily redistribute more than dictatorships; autocrats may distribute more to buy support
\end{itemize}
  \endcol
\endcols

# "The Dictator's Handbook" (Bdm et al 2011)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/bdm}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item The key to political survival is to properly manage the "3-D Politics": \textbf{Nominal selectorate}, {real selectorate} and {winning coalition}
  \vspace{0.4cm}
  \item Any political leader, if conditions permitted, can reshuffle the members of these groups
  \vspace{0.4cm}
  \item The relative size of real selectorate and winning coalition will determine the preference of redistribution using public goods or particularistic private benefits
\end{itemize}
  \endcol
\endcols

# Relative size of real selectorate and winning coalition is important

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/selectorate.pdf")
```

---

```{r echo=FALSE, out.width="62%", fig.align="center"}
knitr::include_graphics("Figs/two_pla.pdf")
```
\begin{center}
\small
The military delegation at the annual plenary session of the National People’s Congress, which opens on Wednesday, will number 267, down from the 281 appointed in 2023. Fourteen have been removed because of corruption investigations."
\end{center}

# "Spin Dictators" (Guriev and Treisman 2022)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/spin_auto}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item \textbf{Fear dictators} maintains power by violently repressing and punishing opposition, controlling all communication tools, denouncing the pluralist democracy and blocking cross-border flows of information and people
  \vspace{0.4cm}
  \item \textbf{Spin dictators}, in contrast, focus on shaping public opinion than directly repressing people, which appears to be a more efficient and less costly alternative
  \end{itemize}
  \endcol
\endcols

# "Spin Dictators" (Guriev and Treisman 2022)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/spin_auto}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item Examples of "fear" dictators: Stalin in the former Soviet Union, Hitler in Nazi Germany and Mao in China
  \vspace{0.5cm}
  \item Examples of "spin" dictators: Victor Orbán in Hungary, Lee Hsien Loong in Singapore and Erdoğan in Turkey
  \vspace{0.5cm}
  \item Do you agree with the typology? Any other examples you can come up with?
  \end{itemize}
  \endcol
\endcols

# Placing the discussion in the Chinese context

\begin{itemize}
  \item Coercion and control: Preventing elite split and collective action
  \vspace{0.1cm}
  \begin{itemize}
    \item Elite purges and restructuring
    \item Surveillance and micro-management over civil society
    \item Media censorship and propaganda (or publicity)
  \end{itemize}
  \vspace{0.5cm}
  \item Co-optation and representation: Sharing the stake of the regime's survival
  \vspace{0.1cm}
  \begin{itemize}
    \item Natl and local legislatures
    \item Local elections (villages and legislative)
    \item Party membership
    \item Party-state employees
    \item Social policy
  \end{itemize}
\end{itemize}

# Elite purges and restructuring

\begin{itemize}
  \item Cultural revolution under Mao, such as Liu, Shih and Zhang (2018)
  \vspace{1cm}
  \item Anti-corruption campaign under Xi (allegedly), such as Lorentzen and Lu (2018)
\end{itemize}

# "The Fall of the Old Guards" (Liu, Shih and Zhang 2018)

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/res_liuetal.pdf")
```
\vspace{0.1cm}
\begin{small}
"The Cultural Revolution radically reoriented the composition of the elite selectorate from one favoring central agencies to one favoring local interests."
\end{small}

# Lorentzen and Lu (2018)

\begincols
  \begincol{.5\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.36]{Figs/res_ll}
    \end{figure}
  \endcol
  \hspace{0.6cm}
  \begincol{.3\textwidth}
  \small
  \small
  "Almost no individuals connected to the Premier were arrested, but connections to the broader Politburo Standing Committee were not protective."
  \endcol
\endcols

# Surveillance and micro-management

\begin{itemize}
  \item Propaganda and media censorship
  \vspace{0.4cm}
  \item Strategic oppression of "key" societal groups, such as
  \vspace{0.1cm}
  \begin{itemize}
    \item Women and LGBT groups
    \item Non-Han ethnicities
    \item Religious activities
  \end{itemize}
  \vspace{0.4cm}
  \item Innovative modles of control and coercion
  \vspace{0.1cm}
  \begin{itemize}
    \item Micro-level bargained authoritarianism (Lee and Zhang 2013)
    \item Outsourcing state coercion to "thugs" (Ong 2018)
    \item Societal "cellular" atomization and technology-assisted street-level management (Zhou 2022)
  \end{itemize}
\end{itemize}

# King, Pan and Roberts (2013)

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/media.pdf")
```

# Topics are not equally sensitive

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/media2.pdf")
```

# Controlled media freedom for "local officialdom"

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/peter.pdf")
```

# Unpersuasive propaganda as signal of control

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/huang.pdf")
```

# Micro-level management and surveillance

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/micro.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/zhou.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ong_both.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/social.pdf")
```

# Conclusion and looking ahead

\begin{itemize}
  \item Dictators rarely rule by force only, and they will try all means to induce compliance and loyalty
   \vspace{0.3cm}
  \item Authoritarian governance involves intense coalition building (within and outside the inner circle) and strikes the general equilibrium of coercion and co-optation
   \vspace{0.3cm}
  \item China is no exception, and yet a number of puzzles remain to be answered:
  \vspace{0.1cm}
  \begin{itemize}
    \item How can the Party sustain a system of responsiveness without complete electoral accountability?
    \item How can the Party strike the balance between the stick and carrot?
  \end{itemize}
\end{itemize}

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="45%", fig.align="center"}
knitr::include_graphics("Figs/doc.pdf")
```
\vspace{0.15cm}
\begin{center}
\small
"\textbf{I Want To Be A People's Representative}" by Zhitan Jia (2013)
\end{center}