---
title: |
  | 
  |
  | Doing Political and Social Research 
  |
  |
author: |
  | Week 6
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Logistics

\begin{itemize}
  \item Assessment
  \vspace{0.1cm}
  \begin{itemize}
    \item Quiz: Quiz 1 marking in progress; Quiz 2 open on \textbf{10 December 2024}
    \item Essay: Due on \textbf{25 November 2024}
    \item Proposal: Due on \textbf{10 January 2025}; examples on Moodle now
  \end{itemize}  
  \vspace{0.5cm}
  \item Polcies and guidelines
  \vspace{0.1cm}
  \begin{itemize}
    \item Moodle
    \item Office hours
    \item Mitigating circumstances
  \end{itemize}
\end{itemize}

# Strategic reading: The Berkeley style

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ucb.pdf")
```
\vspace{0.5cm}
\begin{itemize}
  \item \textbf{Preview}: Get as much information about the reading before you actually read it
  \vspace{0.35cm}
  \item \textbf{Annotating}: Read with a pencil and making notes as you read
  \vspace{0.35cm}
  \item \textbf{Analyzing}: Break the reading apart to see how different parts relate to each other
  \vspace{0.35cm}
  \item \textbf{Responding}: Think again how the reading relates the topic of each week; come up with summary, comments and questions
\end{itemize}

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/mckie.pdf")
```

# Recap: Concepts

\begin{itemize}
  \item How should we understand concepts? Do you take the classical view or the prototype/family resemblance view?
  \vspace{0.2cm}
  \item What makes for a good concept?
  \vspace{0.2cm}
  \item Consider a concept that you have come across in your studies.
  \vspace{0.1cm}
  \begin{itemize}
    \item Is it useful?
    \item Is it contested?
    \item Has it been stretched?
    \item Can it be sub-divided or aggregated with other concepts? Can you move up and down the ladder of abstraction?
  \end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/process.pdf")
```

# Datasets in political and social reseaech

\begin{itemize}
  \item Admin data, such as social security payment records, educational attainment records, health records, court records and tax records (UK Data Services)
  \vspace{0.3cm}
  \item Survey data, such as ANES and BSA
  \vspace{0.3cm}
  \item Socioeconomic data, such as QoG and VDem
  \vspace{0.3cm}
  \item Smart data: Text, image and sound
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ukdata.pdf")
```
\vspace{0.2cm}
\url{https://ukdataservice.ac.uk/}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/anes.pdf")
```
\vspace{0.2cm}
\url{https://electionstudies.org/}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/bsa.pdf")
```
\vspace{0.2cm}
\url{https://natcen.ac.uk/british-social-attitudes}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/afr.pdf")
```
\vspace{0.2cm}
\url{https://www.afrobarometer.org/}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/wvs.pdf")
```
\vspace{0.2cm}
\url{https://www.worldvaluessurvey.org/}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ihds.pdf")
```
\vspace{0.2cm}
\url{https://ihds.umd.edu/}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/qog.pdf")
```
\vspace{0.2cm}
\url{https://www.gu.se/en/quality-government}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/vdem.pdf")
```
\vspace{0.2cm}
\url{https://v-dem.net/}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/prio.pdf")
```
\vspace{0.2cm}
\url{https://www.prio.org/}

# Smart data: Social media posts

```{r echo=FALSE, out.width="105%", fig.align="center"}
knitr::include_graphics("Figs/king.pdf")
```

# Smart data: Audio recordings

```{r echo=FALSE, out.width="105%", fig.align="center"}
knitr::include_graphics("Figs/women.pdf")
```
