---
title: |
  | 
  |
  | Doing Political and Social Research 
  |
  |
author: |
  | Week 5
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Quiz 1 open on Moodle and due on 5 Nov

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/moodle.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/quiz.pdf")
```

# Revisiting the study of "British values"

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/british.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/asian.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/clashes.pdf")
```

---

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/weber.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/barbara.pdf")
```

---

```{r echo=FALSE, out.width="82%", fig.align="center"}
knitr::include_graphics("Figs/45.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/46.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/paper.pdf")
```

# Varieties of theory in social research

\begin{itemize}
  \item Explanatory forms of theorizing (theory as explanation)
  \vspace{0.6cm}
  \item Interpretive forms of theorizing (theory as interpretation)
  \vspace{0.6cm}
  \item Emancipatory forms of theorizing (theory as ideology)
\end{itemize}

# Varieties of theory in social research

\begin{itemize}
  \item \textbf{Explanatory forms of theorizing}: To identify and establish the fundamental processes and structures that ‘underlie’ and therefore explain a given topic
  \vspace{0.4cm}
  \item \textbf{Interpretive forms of theorizing}: To provide a deep reading and understanding of the circumstances in which people find themselves, working through the layers of "meaning" that constitute our experience of organizations
  \vspace{0.4cm}
  \item \textbf{Emancipatory forms of theorizing}: To inform practical actions so as to subvert such systems towards emancipation and potential reform
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/crt.pdf")
```

# CRT as an interpretative theory

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.5]{Figs/crenshaw}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  CRT is "'is \textbf{a way of seeing}, attending to, accounting for, tracing and analyzing \textbf{the ways} that race is produced," she said, '\textbf{the ways} that racial inequality is facilitated, and \textbf{the ways} that our history has created these inequalities that now can be almost effortlessly reproduced unless we attend to the existence of these inequalities.'"
  \endcol
\endcols

# CRT as an emancipatory theory

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/oiyang}
    \end{figure}
    \vspace{0.2cm}
  \endcol
  \begincol{.5\textwidth}
  "'The founders of CRT critiqued liberal ideologies, and that they called on research scholars to seek out and \textbf{understand the roots of why racial disparities are so persistent, and to systemically dismantle racism}.'"
  \endcol
\endcols

# Explanatory theory of inter-group inequality

\pause
\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/becker}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
    \item \textbf{Statistical discrimination} is based on stereotypes that occur when rational, information-seeking decision makers use aggregate group characteristics to evaluate individuals
  \vspace{0.45cm}
    \item \textbf{Taste-based discrimination} is based on prejudice or racial and gender bias
  \end{itemize}
  \endcol
\endcols
