---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | Welcome and Introduction
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Plan for today

\begin{itemize}
	\item \textbf{About Masterclass}
	\vspace{0.1cm}
	\begin{itemize}
		\item History and module objectives
		\item Weekly syllabus
		\item Assessment and requirements
		\item Module contacts
	\end{itemize}
	\vspace{0.3cm}
	\item \textbf{Introduction}: Tell us about yourself
	\vspace{0.3cm}
	\item \textbf{Exercise}: The mysterious case of "AI for Social Research"
\end{itemize}

# History and objectives

\begin{itemize}
\small
	\item \textbf{Masterclass in Social Research} is the final compulsory module for \textbf{MSc/PgDip Social Research} students
	\vspace{0.1cm}
	\begin{itemize}
		\item Masterclass was created at first to provide SR students with a variety of methods training
		\item While serving as a compulsory module for SR students, Masterclass is also open to PGR students
		\item The topics covered may change from year to year based on staff availability and new developments in social research
	\end{itemize}
	\vspace{0.2cm}
	\item Following recent program restructuring, Masterclass now serves two main objectives
	\vspace{0.1cm}
	\begin{itemize}
		\item Produce a new group of competent social researchers with different career objectives and academic ambitions
		\item Showcase cutting-edge social research methods while helping students reflect on some of the most common ones
	\end{itemize}
\end{itemize}

# Module organization

\begin{itemize}
\small
	\item "Master" lecture series (6-7:30pm)
	\vspace{0.1cm}
	\begin{itemize}
		\item Each week will feature a speaker who is an expert in the selected issue and method
		\item The speakers will decide how they will deliver the lectures, with the use of different pre-sessional and/or class activities
		\item \textbf{The lecture series is open to postgraduate research students and their participation will not be credited}
	\end{itemize}
	\vspace{0.2cm}
	\item "Pieces-of-the-craft" workshops (7:40-9pm)
	\begin{itemize}
		\item Each week will focus on a social research "skill" and the workshop activities will help students complete the required assessment
		\item Between Weeks 6-9, students will participate in the "Social Research Dissertation Fairs," where they share their research ideas and receive feedback
		%\item \textbf{The workshops unless otherwise are only open to students who are taking Masterclass for credits}
	\end{itemize}
	\vspace{0.1cm}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/masterclass_2026.pdf")
```

# Plan for Pieces-of-the-craft (PoC) Workshops

\begin{itemize}
	\item \textbf{Weeks 2-4}: "AI for Social Research"
	\vspace{0.1cm}
	\begin{itemize}
		\item Week 2: Introduction
		\item Week 3: AI and qualitative social research (interviews and discourse analysis)
		\item Week 4: AI and Quantitative Social Research (synthetic data and agentic data analytical pipelines)
	\end{itemize}	
	\vspace{0.3cm}
	\item \textbf{Week 5}: Communicating Your Research
	\vspace{0.3cm}
	\item \textbf{Weeks 6-9}: Social Research Dissertation Fairs
\end{itemize}

# Assessment (if you are taking Masterclass for credits)

\begin{itemize}
	\item Two \textbf{blog posts} (40\%): Write a blog post to summarize and reflect on \textbf{two of the lectures}; one lecture from Weeks 2-5 with the other one from Weeks 6-10
	\vspace{0.3cm}
	\item A \textbf{critical reflection paper} (60\%): (4,000 words max) to discuss how Masterclass helps  you advance/rethink social research by providing
	\vspace{0.1cm}
	\begin{itemize}
		\item a summary of your research project, 
		\item a discussion on how AI can change/help with your project and
		\item point-to-point responses to the questions and/or comments you receive during the presentation
	\end{itemize}
\end{itemize}

# Role of the moderator (i.e., Chao-yo)

\begin{itemize}
	\item The moderator will join the lecture series to
	\vspace{0.1cm}
	\begin{itemize}
	  \item Introduce the speaker
	  \item Moderate the Q\&A or discussion, if necessary
	\end{itemize}
	\vspace{0.3cm}
	\item The moderator will lead the workshops and provide feedback on your assignments 
	\vspace{0.3cm}
	\item The moderator is available to answer your questions about the module
	\vspace{0.1cm}
	\begin{itemize}
	  \item Students are encouraged to schedule 1:1 meetings during my office hours (Friday 2-4pm) or by appointment
	  \item All meetings take place via Teams by default -- email if you'd like to meet in person
	  %\item Read the module handbook for details contact policy
	\end{itemize}
\end{itemize}

# Notes for students who are auditing Masterclass

\begin{itemize}
    \item Attendance at the speaker series and the ``AI for Social Research'' workshops is selective and entirely at your discretion
    \vspace{0.3cm}
    \item Attendance is voluntary and will not be recorded
    \vspace{0.3cm}
    \item Please consult your supervisor beforehand to ensure that your participation contributes meaningfully to your doctoral research
    \vspace{0.3cm}
    \item You will be invited to contribute to a blog post at the end of the term
\end{itemize}

# Knowing each other

\begin{itemize}
	\item Please share your name, course, and subject area (e.g., political science) with us
	\vspace{0.5cm}
	\item What are the methods you have learned or used previously?
	\vspace{0.5cm}
	\item What project are you working on at this moment?
\end{itemize}

# Recap: Investigating the social world as a vocation

\begin{itemize}
  \item As a social researcher, we aim to produce "valid" knowledge about the social world
  \vspace{0.2cm}
  \item The production of valid knowledge involves a series of careful thinking about ontology/epistemology, theory-building, and methodology (or the theory-empirics alignment)
  \vspace{0.1cm}
  \begin{itemize}
    \item \textbf{Reveal nuances of the real world}: constructivism and interpretivism ; theory as approach; interpretation; qualitative; inductive/abductive
    \item \textbf{Search for the "ultimate" truth}: positivism; theory as paradigm; explanation; quantitative; deductive
  \end{itemize}
  \vspace{0.2cm}
  \item A new definition of "better" social research is emerging as boundaries across subject areas and research methods are constantly being contested and redefined
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/biases_done.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/subject_done.pdf")
```

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/openai.pdf")
```
\vspace{0.2cm}
\begin{center}
\url{https://openai.com/research}
\end{center}

# AI-4-Science: A paradigm shift? 

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/ai4sci}
    \end{figure}
  \endcol
  \hspace{0.2cm}
  \begincol{.5\textwidth}
  "AI for Science (AI4S) represents the \textbf{convergence} of artificial intelligence (AI) innovation in scientific research and AI-driven scientific discovery, demonstrating their deep integration, and the establishment of a transformative research paradigm."
  \endcol
\endcols

---

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.25]{Figs/ai4sci}
    \end{figure}
  \endcol
  \hspace{0.2cm}
  \begincol{.5\textwidth}
  \small
  "Traditional scientific discovery involves \textbf{generating and validating candidate hypotheses} from a large solution space, often characterized by low efficiency and challenges in identifying high-quality solutions. AI harnesses its \textbf{powerful data processing and analytical capabilities} to navigate solution spaces more efficiently, enabling the generation of high-quality candidate hypotheses."
  \endcol
\endcols

# AI and social sciences


```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ai_for_social.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pandora.pdf")
```
