---
title: |
  |
  |
  | Social Research Dissertation Autumn Workshop
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Today's workshop is for

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/groups.pdf")
```

# Plan for today

\begin{itemize}
  \item Why dissertation (and what is it)?
  \vspace{0.2cm}
  \item An overview of the dissertation process
  \vspace{0.2cm}
  \item Getting started: Dissertation proposal form
  \vspace{0.2cm}
  \begin{itemize}
    \item Personal information
    \item Title and subject area
    \item "Scholarly" literature
    \item Data and methods
    \item Choosing (naming) your potential supervisor
  \end{itemize}
  \vspace{0.2cm}
  \item Concluding remarks and Q\&A
\end{itemize}

# Why dissertation? And what is it?

\begin{itemize}
  \item An \textbf{independent}, \textbf{original} research project, guided by your supervisor(s)
  \vspace{0.3cm}
  \item Resembles a \textbf{peer-reviewed journal article} (12,000-15,000 words, excluding appendix)
  \vspace{0.3cm}
  \item A \textbf{rite of passage}
  \begin{itemize}
    \item from a student to a scholar
    \item from consuming to producing knowledge
    \item from an empirical "technician" to a well-informed social "researcher"
  \end{itemize}
\end{itemize}

# Dissertation process: An overview

\begin{itemize}
  \item Term 1: Prepare your \textbf{dissertation proposal form}; talk to the program director and/or your potential supervisor (if there is one)
  \vspace{0.2cm}
  \item Terms 2 and 3
    \begin{itemize}
      \item January: Submit your proposal form by \textbf{12 January 2025}
      \item February to early March: Supervisor allocation
      \item March to June/early July: Supervision in progress; complete ethics application and risk assessment, if necessary
      \item July-August: Supervision ends (typically); complete your dissertation
      \item September: Submit your dissertation by \textbf{15 September 2025} (tentative)
    \end{itemize}
  \vspace{0.2cm}
  \item After September 2025: Extension and deferral (TBA)
\end{itemize}

# Dissertation activities and events

\begin{itemize}
  \item \textbf{Term 1} (Autumn) Workshop by Dr Chao-Yo Cheng
  \vspace{0.3cm}
  \item \textbf{Term 2} (Spring) Workshop by Drs Laura Richards-Gray and Barry Maydom
  \begin{itemize}
    \item Literature review
    \item Time management
    \item Ethics and risk assessment
  \end{itemize}
  \vspace{0.3cm}
  \item \textbf{Term 3} (Summer) Social Research Dissertation Fairs, as part of \textbf{Masterclass}
  \vspace{0.3cm}
  \item Additional activities: Dissertation bootcamps (check with the Study Skills Team and UoL)
\end{itemize}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/form.pdf")
```
\vspace{0.2cm}
\begin{center}
You can find the dissertation proposal form on Moodle.
\end{center}

# Dissertation proposal form

\pause
\begin{small}
\begin{itemize}
  \item The proposal form is \textbf{NOT marked}; no written comments will be provided (but you can/should come to my office hours to talk about it)
  \vspace{0.3cm}
  \item The proposal form \textbf{helps us find you a supervisor} that best suits your research project
  \vspace{0.3cm}
  \item The proposal form \textbf{shows your tentative ideas} for the dissertation
  \vspace{0.3cm}
  \item The proposal form should be \textbf{short, succinct} and yet \textbf{sufficiently coherent and informative} (2-2.5 pages max)
  \vspace{0.3cm}
  \item Submit your proposal form by \textbf{12 January 2025} on Moodle
\end{itemize}
\end{small}

# Proposal form: Basic personal information

\begin{itemize}
  \item Name
  \item Email address
  \item MSc/MRes programme
  \item Full-time or part-time
\end{itemize}
\pause
\vspace{1.2cm}
\begin{center}
\textbf{The information you provide needs to be consistent with what we see on MyBirkbeck}
\end{center}

# Proposal form: Title and subject area

\begin{itemize}
  \item Title of proposed dissertation: Short v long? Creative v informative?
  \vspace{0.3cm}
  \pause
  \item Description of subject area: A brief statement to introduce your research (like a paper's abstract)
  \begin{itemize}
    \item What are the \textbf{research questions}?
    \item What is the \textbf{focus of previous research literature}?
    \item What is the \textbf{context and/or background}? What is the \textbf{scope condition}?
    \item What is \textbf{distinctive about your theory position or intellectual approach}?
    \item What are \textbf{the tentative methods and data sources}? 
    \item \textbf{[Not yet]} What are your benchmark findings? What "new facts" have you found? Or what key substantive conclusions do you draw? The value-added or originality of your work within this field? 
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/mason.pdf")
```
\begin{center}
\begin{scriptsize}
Mason, J. 2002. \textit{Qualitative researching} (2nd). London: Sage, p.18.
\end{scriptsize}
\end{center}

# Proposal form: "Scholarly literature"

\begin{itemize}
  \item List the relevant "scholarly" literature(s); arrange the literature(s) by topic/keyword if you'd like
  \pause
  \vspace{0.2cm}
  \item Identify the gap, such as
  \begin{itemize}
    \item Theoretical/empirical tensions or contradictions
    \item Different or alternative theoretical perspectives
    \item Different methodological approach (e.g., build on more inductive work to propose some hypothesis-testing deductive approach)
    \item Extension to new context (e.g., different region, time period)
    \item Untested conventional wisdom left untested or understudied (e.g., critical element not engaged with, need to consider other conditions)
  \end{itemize}
  \vspace{0.2cm}
  \item Use literature to think through your RQs and supervisor preferences
\end{itemize}

---

```{r echo=FALSE, out.width="94%", fig.align="center"}
knitr::include_graphics("Figs/say.pdf")
```

---

```{r echo=FALSE, out.width="94%", fig.align="center"}
knitr::include_graphics("Figs/annual.pdf")
```
\vspace{0.2cm}
\begin{center}
\url{https://www.annualreviews.org/}
\end{center}

# Proposal form: "Primary research material"

\begin{itemize}
  \item A suggestive research design where you
  \begin{itemize}
    \item describe the type(s) of evidence which will be collected and analyzed
    \item show that the proposed evidence can feasibly be collected given the time constraint
    \item outline potential method(s) that will be used to analyze the evidence
  \end{itemize}
  \vspace{0.2cm}
  \item It is okay to be tentative (we know things will change), but aim tob uild a coherent alignment between question, theory, and empirics; you can use your observations to \textit{evaluate} or \textit{challenge} a theory
  \vspace{0.2cm}
  \item The key task is to provide original analysis, which can be based on existing data
\end{itemize}

---

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/ssrc.pdf")
```
\begin{center}
\begin{small}
\url{https://www.ssrc.org/publications/the-art-of-writing-proposals/}
\end{small}
\end{center}

# Proposal form: Potential supervisor

Supervisor allocation is a \textbf{collaborative} process between the Program Director and you; you can/need to
\pause
\vspace{0.3cm}
\begin{itemize}
  \item Nominate one or more (permanent) academic staff members from
  \begin{itemize}
    \item School of Social Sciences
    \item Other subject areas, such as applied linguistics, criminology, and law in other Schools
  \end{itemize}
  \vspace{0.3cm}
  \item Show some knowledge of the work by your potential supervisor and provide explanations in the proposal form
  \vspace{0.3cm}
  \item Start informal conversations with me and/or your potential supervisors
\end{itemize}

# What supervisors do (and do not do)

\begin{itemize}
    \item Min \textbf{three} meetings (each about one hour) before summer break
    \vspace{0.1cm}
    \begin{itemize}
        \item It is your responsibility to arrange meetings and keep in good contact with your supervisors
        \item \textbf{Supervisors can}: Offer guidance and support (but not necessarily direction); advise you on approach, scope, questions to be asked, and the outline structure and research design and methods
        \item \textbf{Supervisors may not}: Comment on full drafts and meet you during the summer break
    \end{itemize}
    \vspace{0.3cm}
    \pause
    \item Read through the following documents for more information
    \vspace{0.1cm}
    \begin{itemize}
        \item "Guidelines for MRes and MSc Social Research Dissertations"
        \item "Policy on Supervised Dissertations for Taught Students"
    \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/staff.pdf")
```
\begin{center}
\begin{scriptsize}
\url{https://www.bbk.ac.uk/school/social-sciences/our-staff}
\end{scriptsize}
\end{center}

---

```{r echo=FALSE, out.width="92%", fig.align="center"}
knitr::include_graphics("Figs/tom.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/moodle.pdf")
```

# Course of action (until we meet again in Summmer 2025)

\begin{itemize}
  \item Revise workshop slides and recommended readings on Moodle
  \vspace{1cm}
  \item Start thinking about your dissertation (if you haven't) and talk to me and/or potential supervisor
  \vspace{1cm}
  \item Work on the proposal form and submit it by 12 January 2025
\end{itemize}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/contacts.pdf")
```
