---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | Week 1: Welcome and Introduction
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Plan for today

\begin{itemize}
	\item \textbf{About the Masterclass}
	\vspace{0.1cm}
	\begin{itemize}
		\item History and learning objectives
		\item Weekly syllabus
		\item Assessment and requirements
		\item Module contact
	\end{itemize}
	\vspace{0.3cm}
	\item \textbf{Introduction}: Tell us about yourself
	\vspace{0.3cm}
	\item \textbf{Exercise}: Find your research buddies
\end{itemize}

# History and objectives

\begin{itemize}
\small
	\item The \textbf{Masterclass in Social Research} is the final 30-credit compulsory module for \textbf{MSc Social Research} students
	\vspace{0.1cm}
	\begin{itemize}
		\item The Masterclass was created at the beginning of the Postgraduate Social Research Programmes to provide students at Birkbeck with a variety of methods training
		\item While serving as a compulsory module for MSc students, the Masterclass is also open to research postgraduate students
		\item The topics covered by the Masterclass may change from year to year based on staff availability and new developments in social research
	\end{itemize}
	\vspace{0.2cm}
	\item Following recent program restructuring, the Masterclass now serves two main objectives
	\vspace{0.1cm}
	\begin{itemize}
		\item Produce a new group of competent social researchers with different career and academic interests
		\item Showcase cutting-edge social research methods while helping students reflect on some of the most common methods
	\end{itemize}
\end{itemize}

# Module organization

\begin{itemize}
\small
	\item "Master" lecture series (6-7:30pm)
	\vspace{0.1cm}
	\begin{itemize}
		\item Each week will feature a speaker from an SSHP department who is an expert in the selected issue and method
		\item The speakers will decide how they will deliver the lectures, with the use of different pre-sessional and/or in-class activities
		\item \textbf{The lecture series is open to all taught and research postgraduate students and their participation will not be credited}
	\end{itemize}
	\vspace{0.2cm}
	\item "Pieces-of-the-craft" workshops (7:40-9pm)
	\begin{itemize}
		\item Each week will focus on a social research "skill" and the workshop activities will help students complete the required assessment.
		\item Between Weeks 8-11, students will participate in the "dissertation fairs" where they share their research ideas and receive feedback.
		\item \textbf{The workshops are only open to students who are taking the Masterclass for credits}.
	\end{itemize}
	\vspace{0.1cm}
\end{itemize}

# Weekly syllabus: Lecture series

\begin{small}
\begin{itemize}
	\item \textbf{Week 1}: Welcome and introduction
	\vspace{0.25cm}
	\item \textbf{Weeks 2-5}: Qualitative Social Research
	\begin{itemize}
		\item Oral history (Prof Jasmine Gideon)
		\item Ethnography (Dr Ben Gidley)
	  \item Discourse analysis (Dr Margarita Aragon)]
	  \item Historical research and archives (Dr Brodie Waddell)
	\end{itemize}
	\vspace{0.25cm}
	\item \textbf{Weeks 7-10}: Social Data Science 
	\begin{itemize}
		\item Survey (Dr Barry Maydom)
		\item Social network analysis (Dr Chao-Yo Cheng)
		\item Geographic data science (Dr Roberto Murcio)
		\item Text-as-data (Dr Laszlo Horvath)
	\end{itemize}
	\vspace{0.25cm}
	\item \textbf{Weeks 11}: Quali-quant methods for social media research (Drs Scott Rodgers and Rob Topinka)
\end{itemize}
\end{small}

# Weekly syllabus: Workshops and presentations (tenative)

\begin{itemize}
	\item \textbf{Week 2}: Academic blogging and podcasting
	\vspace{0.2cm}
	\item \textbf{Week 3}: Transparency and open social science 
	\vspace{0.2cm}
	\item \textbf{Week 4}: Using AI for social research
	\vspace{0.2cm}
	\item \textbf{Week 5}: Research talks for work in progress
	\vspace{0.2cm}
	\item \textbf{Week 7-10}: Dissertation fairs (student presentations)
	\vspace{0.2cm}
	\item \textbf{Week 11}: SR happy hours
\end{itemize}

# Assessment (if you are taking MSR for credits)

\begin{itemize}
	\item \textbf{Two blog posts} (40\%): Write a blog post to summarize and reflect on \textbf{two of the lectures}; one lecture from Weeks 2-5 with the other one from Weeks 7-11
	\vspace{0.2cm}
	\item \textbf{A critical reflection paper} (60\%)
	\begin{itemize}
		\item Write a critical reflection paper (3,500-4,000 words) to discuss how the Masterclass helps with you research
		\item Provide (1) a summary of your research project, (2) a discussion on how the lecture series help with your project and (3) a response to the questions/comments/critiques you receive during the presentation
	\end{itemize}
\end{itemize}

# Role of the moderator

\begin{itemize}
	\item The moderator will join the lecture series to
	\vspace{0.1cm}
	\begin{itemize}
	  \item Introduce the speaker
	  \item Moderate the Q\&A or discussion, if necessary
	\end{itemize}
	\vspace{0.3cm}
	\item The moderator will lead the workshops and provide feedback on your assignments 
	\vspace{0.3cm}
	\item The moderator is available to answer your questions
	\vspace{0.1cm}
	\begin{itemize}
	  \item Students are encouraged to schedule 1:1 meetings during my office hours (Friday 3-5pm) or by appointment
	  \item All meetings take place via Teams by default -- email if you'd like to meet in person
	  \item Read the module handbook for details contact policy
	\end{itemize}
	
\end{itemize}

# Plan for next week

\begin{itemize}
	\item Lecture series: Oral history
	\vspace{0.4cm}
	\item PoC workshops: Academic blogging
	\vspace{0.1cm}
	\begin{itemize}
		\item Before you come: Finish the LSE post and browse one of the examples
		\item In class: Use the lecture on surveys as a practice
	\end{itemize}
\end{itemize}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/subject_done.pdf")
```

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="102%", fig.align="center"}
knitr::include_graphics("Figs/biases_done.pdf")
```

# Knowing each other

\begin{itemize}
	\item Please share your name, course, and subject area (e.g., political science) with us
	\vspace{0.5cm}
	\item What are the methods you have learned or used previously?
	\vspace{0.5cm}
	\item What project are you working on at this moment?
\end{itemize}

# Exercise: Find your research buddies

\begin{itemize}
  \item Share your subject area with each other
  \vspace{0.3cm}
  \item Is your subject area a social "science?" 
  \vspace{0.3cm}
  \item Is your subject area "suffering" from any "bias" named by Grossmann?
  \vspace{0.1cm}
  \begin{itemize}
    \item Individual diversity: Demographic, geographical, and political
    \item Scientific practice: Methodological, disciplinary, and institutional
    \item Innate difficulties: Temporal, cognitive, and application
  \end{itemize}
  \vspace{0.3cm}
  \item If there is more than one subject area being named in your group, is any subject area "more scientific" than others? Why?
  \item Can you come up with a collaborative project?
\end{itemize}

# Recap: Investigating the social world as a vocation

\begin{itemize}
  \item As a social researcher, we aim to produce "valid" knowledge about the social world
  \vspace{0.2cm}
  \item The production of valid knowledge involves a series of careful thinking about ontology/epistemology, theory-building, and methodology (or the theory-empirics alignment)
  \vspace{0.1cm}
  \begin{itemize}
    \item \textbf{Reveal nuances of the real world}: constructivism and interpretivism ; theory as approach; interpretation; qualitative; inductive/abductive
    \item \textbf{Search for the "ultimate" truth}: positivism; theory as paradigm; explanation; quantitative; deductive
  \end{itemize}
  \vspace{0.2cm}
  \item A new definition of "better" social research is emerging as boundaries across subject areas and research methods are constantly being contested and redefined
\end{itemize}
