---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Weel 2: Organising}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week2.pdf")
```

# Today's plan

\begin{itemize}
  \item Logistics: Attendance, class duration and etc
  \vspace{0.2cm}
  \item Recap: "Doing Comparative Politics" (30 mins)
  \vspace{0.2cm}
  \item Refresher lecture and Q\&A (1-1.5 hr)
  \vspace{0.1cm}
  \begin{itemize}
    \item Regimes, states (and countries), governments (and governance)
    \item States, nations and nation-states
    \item How states vary: State capacity and economic development
  \end{itemize}
  \vspace{0.2cm}
  \item Reading exercise: "Democracy and Varieties of Nationalism" (1 hr)
\end{itemize}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/dcp.pdf")
```
\begin{center}
\vspace{0.1cm}
\tiny
\url{https://observer.co.uk/news/international/article/winners-in-the-fight-for-africas-democracies-are-getting-hard-to-find}
\end{center}

# Recap: "Doing Comparative Politics"

\begin{itemize}
  \item What is this about? Why is the issue/question this news article attempts to "discuss" or "explain?"
  \vspace{0.5cm}
  \item What are the "factors" highlighted by the author? How "convincing" are they? 
  \vspace{0.5cm}
  \item What is the comparative method used here?
\end{itemize}

# Question: The "strange" case of democracies in Africa

\begin{itemize}
  \item Many democracies in Sub-Saharan Africa have "reverted" back to authoritarianism, as illustrated by a rising number of
  \vspace{0.2cm}
  \begin{itemize}
    \item coups
    \vspace{0.1cm}
    \item rigged elections
    \vspace{0.1cm}
    \item opposition leaders facing changes or even being "jailed"
  \end{itemize}
\end{itemize}

# Report: "African insights 2024" (Afrobarometer 2025)

\begin{itemize}
\small
  \item Support for democracy remains strong but declining
  \vspace{0.1cm}
  \begin{itemize}
    \item 66\% prefer democracy; large majorities reject authoritarian alternatives
    \item Yet support dropped 7 percentage points over a decade (30 countries), with steep declines in South Africa (-29) and Mali (-23)
  \end{itemize}
  \vspace{0.2cm}
  \item Military rule gaining acceptance
  \vspace{0.1cm}
  \begin{itemize}
    \item Opposition to military rule weakened by 11 percentage points, especially in Mali (-40) and Burkina Faso (-36)
    \item 53\% would accept military takeover if leaders abuse power
  \end{itemize}
  \vspace{0.2cm}
  \item Mixed signals on support for democratic norms
  \begin{itemize}
    \item In favor of: accountability, rule of law, term limits, multiparty competition and media freedom
    \item Support for elections declined 8 points (though majority still favor)
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/afro1.pdf")
```
\begin{center}
\vspace{0.1cm}
\tiny
\url{https://www.afrobarometer.org/feature/flagship-report/}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/afro2.pdf")
```
\begin{center}
\vspace{0.1cm}
\tiny
\url{https://www.afrobarometer.org/feature/flagship-report/}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/afro3.pdf")
```
\begin{center}
\vspace{0.1cm}
\tiny
\url{https://www.afrobarometer.org/feature/flagship-report/}
\end{center}

# Why democracies in Africa are "fragile"

\begin{itemize}
  \item Worldwide democratic backsliding and the difficult "state" of global economy (inflation and cost of living crisis)
  \vspace{0.4cm}
  \item Weakening of the post-1945 international settlement (and the rise of alternatives)
  \vspace{0.4cm}
  \item African democracies are "young;" democractic institutions are "flimsy"
\end{itemize}

# "Refresher" lecture

TBD.