---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 7: Mobilising}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week7.pdf")
```

# Today's plan

\begin{itemize}
  \item Logistics: Start thinking about your final essay 
  \vspace{0.3cm}
  \item Recap from Week 6: Elections
  \begin{itemize}
    \item Electoral systems and politics
    \item Additional things to explore
    \item Extra: Elections in dictatorships
  \end{itemize}
  \vspace{0.3cm}
  \item Refresher lecture: Political parties and party systems
  \vspace{0.1cm}
  \begin{itemize}
    \item Political parties in "modern" democracies
    \item Where political parties come from: Institutional and societal explanations
    \item Extra: Political parties in authoritarian regimes (Ben Smith 2005)
    \item Reading exercise: "Getting Good Politicians into Politics"
  \end{itemize}
\end{itemize}

---

\begin{itemize}
\small
  \item Varieties of constitutional democracy
  \vspace{0.1cm}
  \begin{itemize}
    \item Executive system: presidential, parliamentary and semi-presidential
    \item Legislative system: uni- or bi-cameral
    \item \textbf{Electoral system}: majoritarian/plurality, PR and mixed-member (and the resulting party system)
    \item Central-local relations: federal, unitary and decentralized
  \end{itemize}
  \vspace{0.3cm}
  \item Democratic constitutional design is a response to "\textbf{Madison's Dilemma}"
  \vspace{0.1cm}
  \begin{itemize}
    \item The goal is to create a government strong enough to govern while limited enough to protect civil liberties
    \item Two visions of democracy: Majoritarian Model (power concentration) versus Consensus Model (power dispersion)
  \end{itemize}
\end{itemize}

# Recap: Electoral systems and their consequences

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/elecworld_tree.pdf")
```

---

\begin{itemize}
  \item \textbf{Ballot structure}: Vote for party or individual candidate? Perhaps both?
  \vspace{0.3cm}
  \item \textbf{Electoral formula}: How are votes translated into seats? Should we search for relative vs. absolute majority (i.e., how "permissive" should the formula be)? 
  \vspace{0.3cm}
  \item \textbf{District magnitude}: How many seats should be elected from a district? Single-member or multi-member?
\end{itemize}

# Electoral systems and their corresponding democratic visions

Permissive?

# Policy consequences of electoral systems: Income inequality

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/gini_system.pdf")
```

# Policy consequences of electoral systems: Gender gap

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/gender_system2.pdf")
```

# Other questions we ask about democratic elections

\begin{itemize}
\small
  \item Quality of elections: What are the metrics?
  \vspace{0.1cm}
  \begin{itemize}
    \item Candidate quality (e.g., education)
    \item Inclusiveness (e.g., voter registration, voter ID) 
    \item Integrity (e.g., vote buying and ballot box manipulation)
  \end{itemize}
  \vspace{0.2cm}
  \item Electoral violence -- why does it happen and how can we stop it?
  \vspace{0.2cm}
  \item Descriptive representation of marginalized social groups and discrimination
  \vspace{0.2cm}
  \item Why "good" elections do not lead to responsiveness governments?
\end{itemize}

---

Candidate quality.

---

Integrity -- Susan Hyde

---

Electoral violence

---

Representation and discrimination

---

Good elections and responsiveness govt

# Refresher lecture

# Why parties

Summary of John Aldrich.

# Parties in modern democracy

# Parties versus interest group

# Where parties come from

# Societal divide

---

Subaltern or ethnic/religious parties

# Duverger's law: "Plurality rule in single-member districts tends to produce a two-party system"

\begin{columns}
  \column{0.5\linewidth}
  \begin{itemize}
    \item Disproportionality: Largest two parties will get disproportionally more seats than other parties
    \vspace{0.4cm}
    \item Strategic voting: Voters do not wish to "waste" their votes on candidates with no realistic chance of winning
  \end{itemize}
  \column{0.46\linewidth}
  \centering
  \includegraphics[scale=0.5]{Figs/duverger}
\end{columns}

# Effective number of political parties

# Electoral disproportionality

\begin{equation}
  \text{Disproportionality} = \frac{1}{n}\sum_i^n |v_i - s_i|, 
\end{equation}
where $n$ refer to the number of parties contesting in the election. Also, $v_i$ and $s_i$ refer to the vote share and seat share of party $i$, respectively.
\vspace{0.2cm}
\begin{figure}
\centering
\includegraphics[scale=0.46]{Figs/dis}
\end{figure}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/uk2015.pdf")
```

# Reading exercise

TBD

# Extra: Parties in authoritarian regimes

Ben

---

China