---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 7: Mobilising}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week7.pdf")
```

# Today's plan

\begin{itemize}
  \item Recap from Week 6: Elections
  \begin{itemize}
    \item Electoral systems and politics
    \item Additional things to explore
    \item Extra: Elections in dictatorships
  \end{itemize}
  \vspace{0.4cm}
  \item Refresher lecture: Political parties and party systems
  \vspace{0.1cm}
  \begin{itemize}
    \item Political parties in "modern" democracies
    \item Where political parties come from: Institutional and societal explanations
    \item Extra: Political parties in authoritarian regimes (Ben Smith 2005)
    \item Reading exercise: "Getting Good Politicians into Politics"
  \end{itemize}
\end{itemize}

# Where we are now

\begin{itemize}
\small
  \item Varieties of constitutional democracy
  \vspace{0.1cm}
  \begin{itemize}
    \item Executive system: presidential, parliamentary and semi-presidential
    \item Legislative system: uni- or bi-cameral
    \item Electoral system: majoritarian/plurality, PR and mixed-member (and the resulting \textbf{party system})
    \item Central-local relations: federal, unitary and decentralized
  \end{itemize}
  \vspace{0.3cm}
  \item Democratic constitutional design is a response to "\textbf{Madison's Dilemma}"
  \vspace{0.1cm}
  \begin{itemize}
    \item The goal is to create a government strong enough to govern while limited enough to protect civil liberties
    \item Two visions of democracy: Majoritarian Model (power concentration) versus Consensus Model (power dispersion)
  \end{itemize}
\end{itemize}

# Recap: Electoral systems around the world

\begin{itemize}
\small
  \item \textbf{Ballot structure}: Vote for party or individual candidate? Perhaps both?
  \vspace{0.2cm}
  \item \textbf{Electoral formula}: How are votes translated into seats? Should we search for relative vs. absolute majority (i.e., how "permissive" should the formula be)? 
  \vspace{0.2cm}
  \item \textbf{District magnitude}: How many seats should be elected from a district? Single-member or multi-member?
\end{itemize}

# All of these determine how "\textit{permissive}" elections can be

\begin{itemize}
  \item For a candidate, it is easier to be elected when
  \vspace{0.3cm}
  \begin{itemize}
    \item Running for elections as part of the party list
    \vspace{0.2cm}
    \item Running for elections that do not require majority to win
    \vspace{0.2cm}
    \item Running in a district in which more than one candidate can win
  \end{itemize} 
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/elecworld_tree.pdf")
```

# Electoral systems and gender gap

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/gender_system2.pdf")
```

# Other questions we ask about democratic elections

\begin{itemize}
\small
  \item Quality of elections: What are the metrics?
  \vspace{0.1cm}
  \begin{itemize}
    \item Candidate quality (e.g., education)
    \item Inclusiveness (e.g., voter registration, voter ID) 
    \item Integrity (e.g., vote buying and ballot box manipulation)
  \end{itemize}
  \vspace{0.2cm}
  \item Electoral violence -- why does it happen and how can we stop it?
  \vspace{0.2cm}
  \item Descriptive representation of marginalized social groups and discrimination
  \vspace{0.2cm}
  \item Why "good" elections do not lead to responsiveness governments?
\end{itemize}

# Getting "good" politicians into politics

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/voxdev.pdf")
```

# Criminal politicians in India

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/crime_mv.pdf")
```

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/crime_cyc.pdf")
```

# Studying electoral integrity: The NELDA dataset

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/nelda_people.pdf")
```
\vspace{0.1cm}
\begin{center}
\url{https://nelda.co/}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ofosu.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/russia2.pdf")
```

---

```{r echo=FALSE, out.width="95%", fig.align="center"}
knitr::include_graphics("Figs/russia3.pdf")
```

# Electoral violence

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/violence.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/cameroon.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/japan1.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/japan2.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/violence_england.pdf")
```
\vspace{0.1cm}
\begin{center}
The Brentford Election, 1768
\end{center}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/violence_england2.pdf")
```
\vspace{0.1cm}
\begin{center}
The Election Riot in the Market Place, Nottingham 1865
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/mares.pdf")
```

# Elections and representation

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/india.pdf")
```

# Elections and representation

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/mala.pdf")
```

# Refresher lecture: Parties and party systems

\begin{itemize}
  \item Parties and political parties
  \vspace{0.1cm}
  \begin{itemize}
    \item Why parties and what they do (in contrast to interest groups)
    \item Where parties come from: Institutional and societal explanations
  \end{itemize}
  \vspace{0.3cm}
  \item Political parties in authoritarian regimes
  \vspace{0.3cm}
  \item Reading exercise
\end{itemize}

# Why parties and what they do (in contrast to interest groups)

\begin{itemize}
  \item Recruitment and socialization of the political elite
  \vspace{0.3cm}
  \item Mobilization of the masses
  \vspace{0.3cm}
  \item A link between the rulers and the ruled
\end{itemize}

# Where parties come from

\begin{itemize}
  \item Duverger's law: Electoral and party systems:  (Maurice Duverger; Gary Cox 1997)
  \vspace{0.5cm}
  \item Political parties and social "cleavages" (Seymour Martin Lipset and Stein Rokkan 1967)
\end{itemize}

# Duverger's law: "Plurality rule in single-member districts tends to produce a two-party system"

\begin{columns}
  \column{0.5\linewidth}
  \begin{itemize}
    \item Disproportionality: Largest two parties will get disproportionally more seats than other parties
    \vspace{0.4cm}
    \item Strategic voting: Voters do not wish to "waste" their votes on candidates with no realistic chance of winning
  \end{itemize}
  \column{0.46\linewidth}
  \centering
  \includegraphics[scale=0.5]{Figs/duverger}
\end{columns}

# Electoral disproportionality

\begin{equation}
  \text{Disproportionality} = \frac{1}{n}\sum_i^n |v_i - s_i|, 
\end{equation}
where $n$ refer to the number of parties contesting in the election. Also, $v_i$ and $s_i$ refer to the vote share and seat share of party $i$, respectively.
\vspace{0.2cm}
\begin{figure}
\centering
\includegraphics[scale=0.46]{Figs/dis}
\end{figure}

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/uk2024.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/cleavage.pdf")
```

# The "cleavage" theory of political parties

\begin{itemize}
  \item "\textbf{Party is organized opinion}" (Benjamin Disraeli) to reflect underlying social cleavages
  \vspace{0.1cm}
  \begin{itemize}
    \item Urban-rural cleavage
    \item Confessional (religious) cleavage
    \item Secular-clerical cleavage
    \item Class cleavage
    \item Post-material cleavage
    \item Ethno-linguistic cleavage
  \end{itemize}
  \vspace{0.2cm}
  \item Individuals as repertoire of identity attributes and strategic electoral mobilization: Cross-cutting or reinforcing 
\end{itemize}

# Ruling coalition by different ethnic parties in Malaysia

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/malaysia.pdf")
```

# "People's" parties in Europe

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/people.pdf")
```

# Green Party as "new" politics in post-industrial European socieities

\vspace{0.1cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/uk_green.pdf")
```

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/uk_green_win.pdf")
```

# Reading exercise

TBD

# Extra: Parties in authoritarian regimes

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/geddes.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/smith1.pdf")
```

---

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/smith2.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/smith_table.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/fewsmith.pdf")
```
