---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | Computational Thinking and Social Science
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# "A new kind of social science" with computational thinking

\begin{itemize}
	\item Computational thinking requires the integration of \textbf{social science} and \textbf{technological} imagination
	\vspace{0.7cm}
	\item Computational social science can involve one or more different \textbf{tasks}
	\vspace{0.7cm}
	\item The use of computational tools and thinking will help us improve the \textbf{speed} and \textbf{scale} of our research
\end{itemize}

# "A new kind of social science" with computational thinking

\begin{itemize}
	\item Computational thinking requires the integration of \textbf{social science} and \textbf{technological} imagination
	\vspace{0.1cm}
	\begin{itemize}
		\item Social science imagination: Ability to think about social phenomena using existing concepts and theories while identifying the interplay between individuals and structures
		\item Technological imagination: Ability to think and develop technologies to better serve research objectives
	\end{itemize}
	\vspace{0.7cm}
	\item Computational social science can involve one or more different \textbf{tasks}
	\vspace{0.7cm}
	\item The use of computational tools and thinking will help us improve the \textbf{speed} and \textbf{scale} of our research
\end{itemize}

# "A new kind of social science" with computational thinking

\begin{itemize}
	\item Computational thinking requires the integration of \textbf{social science} and \textbf{technological} imagination
	\vspace{0.7cm}
	\item Computational social science can involve one or more different \textbf{tasks}
	\vspace{0.1cm}
	\begin{itemize}
		\item Use novel data sources (e.g., big data and digital trace data)
		\item Use new computationally heavyweight analysis methods
		\item Build models to explain social patterns
		\item Identify and address new problems 
		\item Improve social science theory
	\end{itemize}
	\vspace{0.7cm}
	\item The use of computational tools and thinking will help us improve the \textbf{speed} and \textbf{scale} of our research
\end{itemize}

# "A new kind of social science" with computational thinking

\begin{itemize}
	\item Computational thinking requires the integration of \textbf{social science} and \textbf{technological} imagination
	\vspace{0.7cm}
	\item Computational social science can mean different things
	\vspace{0.7cm}
	\item The use of computational tools and thinking will help us improve the \textbf{speed} and \textbf{scale} of our research
	\begin{itemize}
		\item Skills to command and speak with computer
		\item Skills to ask interesting and important questions
		\item Skills to understand the possibilities computers provide for scholars
	\end{itemize}
\end{itemize}

# Thinking outside the box

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/neil.pdf")
```
\vspace{0.2cm}
\begin{center}
\textbf{\url{https://youtu.be/io6QdGcoWMU}}
\end{center}

# CIVICA Data Science seminar series

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/youtube}
    \end{figure}
    \vspace{0.2cm}
  \begin{tiny}
	\begin{center}
	\textbf{\url{https://youtube.com/playlist?list=PLj2A-_5PZ13KELSBSEVLxPcPQbOao5eSt}}
	\end{center}
	\end{tiny}
  \endcol
  \begincol{.5\textwidth}
  \begin{small}
  "The CIVICA Data Science seminar series is the \textbf{interdisciplinary} forum of the CIVICA European University of Social Sciences, and alliance of eight leading higher education and research institutions. We foster a community of researchers, practitioners, and policy makers working to address complex societal challenges through the use of \textbf{novel} methodological and \textbf{data-driven} approaches, drawing from machine learning and quantitative methods from biology, \textbf{sociology}, \textbf{political sciences}, \textbf{economics}, and physics among others disciplines."
  \end{small}
  \endcol
\endcols


---

\vspace{0.6cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/members.pdf")
```

# Exercise: Relating your research to "computers"

\begin{itemize}
	\item \textbf{Stage 1: What a computational social science project looks like}
	\vspace{0.1cm}
	\begin{itemize}
		\item Visit "CIVICA Data Science Seminar" -- find a previous talk related to your subject area and/or broader research interests and read the abstract.
		\item Visit "UK Data Services" -- find the method(s) mentioned in the talk/abstract and read the synopsis to get the intuition. 
	\end{itemize}
	\vspace{0.2cm}
	\item \textbf{Stage 2: Think through the following questions}
	\vspace{0.1cm}
	\begin{itemize}
		\item Why do you choose this talk? What is this talk about? What data and methods were used in the project?
		\item Can your research benefit from incorporating any computational elements, in terms of data and/or the methods?
		\item Can you come up with a research idea using similar methods and/or data?
		\item How can we create an "inclusive" computational social science research environment?
	\end{itemize}
\end{itemize}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/links.pdf")
```

# Next week: Using AI for/in your research

\vspace{0.3cm}
\begin{itemize}
	\item We will discuss the challenges and opportunities brought by \textbf{ChatGPT} (\url{https://openai.com/blog/chatgpt}).
	\vspace{1cm}
	\item We will practice how you can use ChatGPT to help with your research (and the caveats). 
\end{itemize}
