---
title: |
  |
  |
  | Doing Political and Social Research
  |
  |
author: |
  | Week 7
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Quiz 1

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/quiz1.pdf")
```

# Recap: Case studies

\begin{itemize}
  \item Evaluate the analytical strength/relevance of some specified arguments
  \vspace{0.4cm}
  \item Comparative studies v single-case study?
  \vspace{0.4cm}
  \item Additional things to consider
  \vspace{0.1cm}
  \begin{itemize}
    \item Alternative or rival theories?
    \item Alternative methods (interviews, ethnography and archival research)?
    \item Can you combine all of them?
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/treisman.pdf")
```

# General principles of archival research

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.5]{Figs/history}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{small}
  \begin{itemize}
    \item Be clear about the questions; think through the strategies (induction v deduction)
    \vspace{0.1cm}
    \item Be mindful of the underlying politics of information production, conservation, and declassification
    \vspace{0.1cm}
    \item Be transparent as you are trying to form your interpretations (as well as the selection of details to be reported)
    \vspace{0.1cm}
    \item Be patient and open-minded -- read (and document) as much information as possible; do not jump into conclusion
  \end{itemize}
  \end{small}
  \endcol
\endcols

---

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/iraq}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \textbf{RQ: Daily political life in dictatorships largely remains a black box}
  \vspace{0.4cm}
  \begin{small}
    \begin{itemize}
      \item What did the Ba'thist regime actually know about its citizens?
      \vspace{0.15cm}
      \item Why did it use blunt, seemingly suboptimal, forms of punishment against its population?
      \vspace{0.15cm}
      \item What explains variation in the types of compliance and resistance behaviors undertaken by Iraqis during Hussein’s dictatorship?
    \end{itemize}
  \end{small}  
  \endcol
\endcols

---

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/iraq}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
    \begin{small}
    \vspace{0.2cm}
    \textbf{Archives}
    \vspace{0.2cm}
    \begin{itemize}
      \item "The Iraqi Secret Police Files" (University of Colorado Boulder)
      \vspace{0.2cm}
      \item Miscellaneous tapes and associated transcripts (Hoover Institution)
    \end{itemize}
    \vspace{0.8cm}
    \textbf{Interviews} via the Iraq Memory Foundation
    \end{small}
  \endcol
\endcols

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/eic_cover.pdf")
```

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/eic_source.pdf")
```

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/eic_content.pdf")
```
\pause
\vspace{0.2cm}
\begin{center}
\small
Digitalization, deskwork, sampling and reading
\end{center}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/hoover.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.hoover.org/library-archives}
\end{scriptsize}
\end{center}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/bl.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.bl.uk/}
\end{scriptsize}
\end{center}

---

\vspace{0.5cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/loc.pdf")
```
\vspace{0.1cm}
\begin{center}
\begin{scriptsize}
\url{https://www.loc.gov/}
\end{scriptsize}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/vdem.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://v-dem.net/}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/freedom.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://freedomhouse.org/}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/dd.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://sites.google.com/site/joseantoniocheibub/datasets/dd}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/polity.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://www.systemicpeace.org/polityproject.html}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/gwf.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://sites.psu.edu/dictators/}
\end{center}
