---
title: |
  | 
  |
  | CP$^2$ Week 10: Control and Co-optation
  |
  |
author: |
  | Dr Chao-Yo Cheng
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Plan for today

\begin{itemize} 
  \item Assessment: Reminders and Q\&A
  \vspace{0.4cm}
  \item Lecture: Control and co-optation
  \vspace{0.1cm}
  \begin{itemize}
    \item Puzzle: "The dictator's dilemma"
    \item Coercive control as the means to prevent elite split and societal collective action
    \item Co-optation and representation (and yet without electoral accountability) as the means to stay responsive and give the mass a stake in the regime's survival 
  \end{itemize}
  \vspace{0.4cm}
  \item Discussion (also to be continued next week): The Covid-19 crisis in China
\end{itemize}

# End-of-term Dim Sum Party

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.3]{Figs/chinatown}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
    \item Date TBD, but likely after the learning journal is due
    \vspace{0.8cm}
    \item More details to out after the final lecture via email
    \vspace{0.8cm}
    \item Everyone is welcome -- let me know if you have any dietary restrictions
  \end{itemize}
  \endcol
\endcols

---

\vspace{0.4cm}
```{r echo=FALSE, out.width="45%", fig.align="center"}
knitr::include_graphics("Figs/doc.pdf")
```
\vspace{0.15cm}
\begin{center}
\small
"\textbf{I Want To Be A People's Representative}" by Zhitan Jia (2013)
\end{center}

# "The dictator's dilemma" (Wintrobe 1998)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/wintrobe}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \small
  \begin{itemize}
  \small
  \item Contrary to the popular belief, the use of repression can be counter-productive for dictators, usually dictatorship operates through political exchange (co-optation)
  \vspace{0.4cm}
  \item The dictators seek the loyal support of interest groups, especially among the military and other political and social forces having the strongest capacity to depose them
\end{itemize}
  \endcol
\endcols

# "The dictator's dilemma" (Wintrobe 1998)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.35]{Figs/wintrobe}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \small
  \begin{itemize}
  \small
  \item Co-optation can be done by redistribution and incorporating groups into the regime apparatus
  \vspace{0.5cm}
  \item democracies do not necessarily redistribute more than dictatorships; autocrats may distribute more to buy support
\end{itemize}
  \endcol
\endcols

# "The Dictator's Handbook" (Bdm et al 2011)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/bdm}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item The key to political survival is to properly manage the "3-D Politics": \textbf{Nominal selectorate}, {real selectorate} and {winning coalition}
  \vspace{0.4cm}
  \item Any political leader, if conditions permitted, can reshuffle the members of these groups
  \vspace{0.4cm}
  \item The relative size of real selectorate and winning coalition will determine the preference of redistribution using public goods or particularistic private benefits
\end{itemize}
  \endcol
\endcols

# Relative size of real selectorate and winning coalition is important

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/selectorate.pdf")
```

# "Spin Dictators" (Guriev and Treisman 2022)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/spin_auto}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item \textbf{Fear dictators} maintains power by violently repressing and punishing opposition, controlling all communication tools, denouncing the pluralist democracy and blocking cross-border flows of information and people
  \vspace{0.4cm}
  \item \textbf{Spin dictators}, in contrast, focus on shaping public opinion than directly repressing people, which appears to be a more efficient and less costly alternative
  \end{itemize}
  \endcol
\endcols

# "Spin Dictators" (Guriev and Treisman 2022)

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/spin_auto}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item Examples of "fear" dictators: Stalin in the former Soviet Union, Hitler in Nazi Germany and Mao in China
  \vspace{0.5cm}
  \item Examples of "spin" dictators: Victor Orbán in Hungary, Lee Hsien Loong in Singapore and Erdoğan in Turkey
  \vspace{0.5cm}
  \item Do you agree with the typology? Any other examples you can come up with?
  \end{itemize}
  \endcol
\endcols

# Placing the discussion in the Chinese context

\begin{itemize}
  \item Coercion and control: Preventing elite split and collective action
  \vspace{0.1cm}
  \begin{itemize}
    \item Elite purges
    \item Surveillance over civil society
  \end{itemize}
  \vspace{0.5cm}
  \item Co-optation and representation: Giving the mass a stake of the regime's survival
  \vspace{0.1cm}
  \begin{itemize}
    \item Natl and local legislatures
    \item Local elections
    \item Party membership
    \item Party-state employees
  \end{itemize}
\end{itemize}

# Coercion and control: Elite purges

\begin{itemize}
  \item Cultural revolution under Mao, see Liu, Shih and Zhang (2018)
  \vspace{1cm}
  \item Anti-corruption campaign under Xi (allegedly), see Lorentzen and Lu (2018)
\end{itemize}

# "The Fall of the Old Guards" (Liu, Shih and Zhang 2018)

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/res_liuetal.pdf")
```
\vspace{0.1cm}
\begin{small}
"The Cultural Revolution radically reoriented the composition of the elite selectorate from one favoring central agencies to one favoring local interests."
\end{small}

# Lorentzen and Lu 2018

\begincols
  \begincol{.5\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.36]{Figs/res_ll}
    \end{figure}
  \endcol
  \hspace{0.6cm}
  \begincol{.3\textwidth}
  \small
  \small
  "Almost no individuals connected to the Premier were arrested, but connections to the broader Politburo Standing Committee were not protective."
  \endcol
\endcols

# Coercion and control: Surveillance over civil society

\begin{itemize}
  \item Propaganda and media censorship (Week 9)
  \vspace{0.4cm}
  \item Strategic repression of "key" societal groups, such as
  \vspace{0.1cm}
  \begin{itemize}
    \item Women and LGBT groups
    \item Non-core ethnicities
    \item Religious activities
  \end{itemize}
  \vspace{0.4cm}
  \item Innovative
  \vspace{0.1cm}
  \begin{itemize}
    \item Outsourcing state coercion to "thugs" (Ong 2018)
    \item Societal atomization and technology-assisted "cellular" street-level governance (Zhou 2022)
  \end{itemize}
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/social.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/ong_both.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/zhou.pdf")
```

# Co-optation through natl and local legislatures

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.42]{Figs/manion}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item Representation in a Leninist party regime might take one of the following forms: Mandate, Lennist and Trustee
  \vspace{0.25cm}
  \item Chinese legislature as an institution of "authoritarian parochialism:" Deliever targeted public goods to geograpgic constituency through extra-legislative means
  \vspace{0.25cm}
  \item The People's Congress system provides a key solution to alleviate the classical "dictator's dilemma:" Information and loyalty
  \vspace{0.25cm}
  \item Responsive without accountable?
\end{itemize}
  \endcol
\endcols

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/res_manion.pdf")
```
\vspace{0.1cm}
\begin{small}
Survey data shows that Chinese local congressmen and women view themselves and act as "delegates."
\end{small}

# Co-optation through local elections

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/election}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item Currently, the Party allows for village elections as well as local elections for the People's Congress delegates
  \vspace{0.4cm}
  \item While elections are not uncommon among authoritarian regimes, few have national and executive elections
  \end{itemize}
  \endcol
\endcols

# Co-optation through local elections

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/election}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item In the case of village elections, they were introduced in the 1990s to tackle peasant grienvances when it comes to fees collection (Berstein and Lu 2003)
  \vspace{0.4cm}
  \item Broadly speaking, they provided an opportunity to monitor local officials and hold them accountable (Mayling Birney 2010)
  \end{itemize}
  \endcol
\endcols

# Co-optation through Party

\vspace{0.2cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.28]{Figs/party}
    \end{figure}
  \endcol
  \begincol{.5\textwidth}
  \begin{itemize}
  \small
  \item People who joined the party under Mao were more likely to report ideological reasons; those in the post-1989 cohort mostly joined for the career benefits, and not just to work for the party or the government (Dickson 2013)
  \vspace{0.3cm}
  \item Jiang Zemin's "Three Represents" opened the Party's door to entrepreneurs, and it appears as a good business opportunity especially when these entrepreneurs were "elected" into the People's Congress (Truex 2014)
  \vspace{0.3cm}
  \item The demographic composition of the Party Congress has changed a lot in the past two decades
  \end{itemize}
  \endcol
\endcols

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/party_elite.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/party_join.pdf")
```

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/hanzhang_both.pdf")
```

# Conclusion

\begin{itemize}
  \item Dictators rarely rule by force only, and they will try all means to induce compliance and loyalty
   \vspace{0.3cm}
  \item Authoritarian governance involves intense coalition building (within and outside the inner circle) and strikes the general equilibrium of coercion and co-optation
   \vspace{0.3cm}
  \item China is no exception, and yet a number of puzzles remain to be answered:
  \vspace{0.1cm}
  \begin{itemize}
    \item How can the Party sustain a system of responsiveness without complete electoral accountability?
    \item How can the Party strike the balance between the stick and carrot?
  \end{itemize}
\end{itemize}

---

\vspace{0.1cm}
\begin{center}
\textbf{Next week: Looking back at China's Covid-19 crisis}
\end{center}
\vspace{0.4cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/covid.pdf")
```
\vspace{0.3cm}
\begin{center}
See you next week!
\end{center}
