---
title: |
  |
  |
  | Masterclass in Social Research
  |
  |
author: |
  | Communicating Your Research
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

# Recap on Assessments

\begin{itemize}
	\item Two \textbf{blog posts} (each 20\%): each \textbf{1,000-word} post to summarize and reflect on two of the lectures (one from Weeks 2-5 and Reading Week; one from Weeks 6-10)
	\vspace{1.5cm}
	\item \textbf{Critical reflection paper} (60\%), a \textbf{4,000-word} (max) piece including
	\vspace{0.1cm}
	\begin{itemize}
		\item A summary of your dissertation idea/project (see \textbf{Little 2016})
		\item A discussion on how AI can help with your research project, and
		\item Brief responses to the questions and/or comments/critiques that emerge from your presentation
	\end{itemize}
\end{itemize}

---

\vspace{0.3cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/plats.pdf")
```

# Podcast for different subject areas

\vspace{0.3cm}
```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/blogs_crop.pdf")
```

# Recommended podcast: New Books Network

\vspace{0.3cm}
\begincols
  \begincol{.4\textwidth}
    \begin{figure}
    \centering
    \includegraphics[scale=0.5]{Figs/nbn_crop}
    \end{figure}
    \vspace{0.2cm}
	\begin{center}
	\textbf{\url{newbooksnetwork.com}}
	\end{center}
  \endcol
  \begincol{.5\textwidth}
  	\begin{itemize}
	\item A consortium of author-interview podcast channels:
	\vspace{0.3cm}
	\begin{itemize}
		\item Aiming to "raise the level of public discourse" by introducing scholars and other serious writers to a wide public via new media
		\vspace{0.3cm}
		\item Convering $100+$ subjects, disciplines, and genres
		\vspace{0.3cm}
		\item Publishing 50 to 75 episodes every week
		\vspace{0.3cm}
		\item Opening to suggestions or pitches
	\end{itemize}
	\end{itemize}
  \endcol
\endcols

# Short videos

\vspace{0.3cm}
```{r echo=FALSE, out.width="110%", fig.align="center"}
knitr::include_graphics("Figs/bbk3mt1_crop.pdf")
```
\vspace{0.5cm}
\begin{scriptsize}
\begin{center}
\textbf{\url{https://youtube.com/playlist?list=PL2Fy-5oxIlb7VLoWUC6j_eq9GJXzSjXfP}}
\end{center}
\end{scriptsize}

# Social media threads

\vspace{0.2cm}
```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/paul.pdf")
```
\vspace{0.1cm}
\begin{scriptsize}
\begin{center}
\textbf{\url{https://twitter.com/ProfPaulPoast/status/1630305966674640896?s=20}}
\end{center}
\end{scriptsize}

---

\begin{itemize}
	\item Start with an introduction for spoilers and/or highlights, put concise statements about the research idea (RQ, literature and context/background) and a quick discussion on methods (if the methods are innovative) and findings
	\vspace{0.4cm}
	\item Some useful tips:
	\vspace{0.2cm}
	\begin{itemize}
		\item Acknowledge those who inspire your research; give credits to your collaborators
		\item Include GIFs/memes or short audios/videos to attract attention
		\item Engage with the replies constructively to boost engagement
		\item Use Twitter Analytics and other tools (e.g., TweetDeck) to manage tweets
	\end{itemize}
\end{itemize}

# Blog posts

\begin{small}
\begin{itemize}
	\item \textbf{REF} (Research Excellence Framework) encourages researchers to use blogs to communicate their research with the public
	\vspace{0.2cm}
	\item Some key principles:
	\begin{itemize}
		\item Use titles for spoilers
		\item Decide how you want to hook/motivate readers -- narrate in a reflective manner (it can be reasonably personal)
		\item Present the core of the topics/projects clearly and simply -- use subject-specific vocabulary wisely
		\item Link additional materials if necessary (not required for this assessment)
		\item Build your research profile -- organizational position (within and outside Birkbeck), links to social media profiles (e.g., Linkedin), and published works and/or any other research activities (e.g., NGOs or think tanks)
	\end{itemize}
	\vspace{0.2cm}
	\item What blogs are available at Birkbeck: \textbf{\url{https://www.bbk.ac.uk/blogs}}
\end{itemize}
\end{small}

---

\vspace{0.2cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/bbk_psrp.pdf")
```
\vspace{0.1cm}
\begin{scriptsize}
\begin{center}
\textbf{\url{https://bbkpsrp.substack.com/}}
\end{center}
\end{scriptsize}

# Discussion: Composing a blog post on today's lecture

\begin{itemize}
	\item Summarize the key highlights of today's lecture to those who did not attend the lecture
	\vspace{0.5cm}
	\item Discuss what have you learned from the lecture, in relation to any relevant research experiences in the past
	\vspace{0.5cm}
	\item How can today's lecture help with or relate to your research?
	\vspace{0.5cm}
	\item What else would you like to learn about today's topic?
\end{itemize}

# Social Research Dissertation Fair

\begin{itemize}
	\item Starting from Week 7, in each workshop \textbf{we will have three to four students} present their dissertation/research projects
	\vspace{0.2cm}
	\item Each presentation will last for about \textbf{15 minutes}, including discussion and Q\&A moderated by the module convener (15 minutes max); we may \textbf{record your presentation} for internal use
	\vspace{0.2cm}
	\item The main objective is to give you the opportunity to \textbf{collect feedback that can help with your dissertation} in a supportive and intimate setting; your project can be in any stage
	\vspace{0.2cm}
	\item During Reading Week: Submit a working title, abstract and your bio to Chao-yo
	\item Weeks 7-9: Send in your \textbf{presentation slides} before the designated presentation data
	\item In your final reflection paper: Address the comments
\end{itemize}

# Key Principles for Your Presentation

\begin{itemize}
	\item \textbf{Structure and Planning}: the extent to which the presentation shows prior planning based on a clear structure and a logical sequencing of topics and ideas
	\vspace{0.4cm}
	\item \textbf{Content}: the extent to which the presentation shows evidence of knowledge of the topic and covers relevant ideas and debates
	\vspace{0.4cm}
	\item \textbf{Understanding}: the extent to which the presentation shows evidence of sufficient grasp of key issues to offer critical analysis and reflection
	\vspace{0.4cm}
	\item \textbf{Communication}: the extent to which the presentation is delivered in the way to maximize communication by interacting with audience, responding to queries, and keeping to time
\end{itemize}

# Presenting Your Research

\begin{itemize}
	\item Students are free to arrange their presentations as they consider suitable for their project, but the presentation should clearly address the following items:
	\vspace{0.3cm}
	\begin{itemize}
		\item Introduction -- explicitly discuss motivation and research question
		\vspace{0.2cm}
		\item Literature review -- focus on identify the gap and/or justify hypothesis (be brief)
		\vspace{0.2cm}
		\item Research design and methods -- make sure you fully explain and justify the methods of use
		\vspace{0.2cm}
		\item Preliminary or expected findings and discussion
		\vspace{0.2cm}
		\item Concluding remarks -- reflections on big picture ("so what?"), limitations and future directions; issues for which you need feedback from your peers
	\end{itemize}
\end{itemize}

# Responding to Comments and Critiques

\begin{itemize}
	\item At the end of your presentation, there will be a Q\&A session
	\vspace{0.4cm}
	\begin{itemize}
		\item Listen to the comments and questions carefully; the key is to collect as many suggestions as possible so no need respond to everything on the spot
		\vspace{0.2cm}
		\item Stay open-minded and do not take things personally; no need to be overly defensive
	\end{itemize}
	\vspace{0.5cm}
	\item Read Humphreys (2017) and Choi (2022) to learn more about how researchers discuss each other's work
\end{itemize}

# Responding to Comments and Critiques

\begin{itemize}
	\item After the presentation, compose a written response to the comments and/or critiques in your critical reflection paper, but you should at least
	\vspace{0.4cm}
	\begin{itemize}
		\item Provide a summary of the comments/critiques
		\vspace{0.2cm}
		\item Address comments selectively and yet carefully
		\vspace{0.2cm}
		\item Disagree with respect if you can (do explain though)
	\end{itemize}
	\vspace{0.5cm}
	\item Read Noble (2017) if necessary
\end{itemize}

# Pracrice: Writing an Abstract for Your Project

\begin{itemize}
	\item Your abstract is a brief statement to introduce your research (<200 words)
	\vspace{0.2cm}
	\begin{itemize}
		\item What is the main \textbf{research question}?
		\item What is the \textbf{gap} that your research trying to fill in based on prior literature? What is \textbf{distinctive to your own theory position or intellectual approach}?
		\item What is the \textbf{context and/or background}? What is the \textbf{scope condition} for general audience?
		\item What are \textbf{your methods and data sources}? 
		\item What are your \textbf{benchmark findings}? What "new facts" have you found (or expect to find)? Or what key conclusions do you draw? The value-added or originality of your work within this field?
	\end{itemize}
\end{itemize}
