---
title: |
  |
  |
  | Comparative Government
  |
  |
author: |
  | Dr Chao-Yo Cheng 
  |
output:
  beamer_presentation:
    includes:
      in_header: mystyle.tex
  slidy_presentation: default
  ioslides_presentation: default
fontsize: 10pt
---

\begin{center}
\textbf{Week 5: Leading}
\end{center}
\vspace{0.3cm}
```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/week5.pdf")
```

# Today's plan

\begin{itemize}
  \item Logistics: Seminar log due this Friday 
  \vspace{0.3cm}
  \item Recap from previous weeks
  \begin{itemize}
    \item Regime change and democratic transition
    \item Comparative authoritarianism and authoritarian governance
    \item Reading exercise: "US Descent Into Authoritarianism"
  \end{itemize}
  \vspace{0.3cm}
  \item Refresher lecture: Executive and legislature branches
  \vspace{0.1cm}
  \begin{itemize}
    \item Executive system: Presidental versus parliamentary democracies
    \item Unicameral and bicameral legislatures
    \item Recap: Supreme/constitutional courts and judicial review
  \end{itemize}
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/bad_bunny_cropped.pdf")
```

# Reading exercise summary: "Flawed" constitutions and democracies

\begin{itemize}
\small
  \item In addition to the technical details, "who" is behind the making of the constitution matters
  \vspace{0.2cm}
  \item Many democracies did not rewrite their constitutions when they went through the democratic transition; that is, regime change does not always lead to constitutional changes
  \vspace{0.2cm}
  \item Democratic constitutions can be "flawed" or "biased" when outgoing autocrats took part in or led the designing of the new constitution of the new democratic government by giving them unfair advantage (how?)
  \vspace{0.2cm}
  \item "Correcting" flawed constitutions requires patience and shrewd statecraft (how?)
  \vspace{0.2cm}
  \item Both history and leadership are pivotal for the practice of constitutions
\end{itemize}

---

```{r echo=FALSE, out.width="70%", fig.align="center"}
knitr::include_graphics("Figs/reading_trump.pdf")
```
\vspace{0.1cm}
\begin{center}
\tiny
\url{https://www.theguardian.com/us-news/ng-interactive/2025/mar/22/trump-administration-authoritarianism}
\end{center}

---

\begin{itemize}
  \item What are the indicators that the US might be rolling into "authoritarianism?"
  \vspace{0.3cm}
  \item Are there any indicators being left out?
  \vspace{0.3cm}
  \item Can you use the US anecdote to come up with the measurement of democratic backsliding? 
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/backslide.pdf")
```
\vspace{0.1cm}
\begin{center}
\scriptsize
\url{https://doi.org/10.1017/S104909652300063X}
\end{center}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/backslide_overall.pdf")
```

# Democracy as "polyarchy" (Robert Dahl 1971)

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/polyarchy_cropped.pdf")
```

# Recap: Democratic backsliding 

\begin{itemize}
  \item Competitiveness of elections, including incumbency (dis)advantage
  \vspace{0.3cm}
  \item Constraints on executive power
  \vspace{0.3cm}
  \item Freedom of press (and other civil liberties)
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/backslide_comp.pdf")
```
\vspace{0.2cm}
\begin{itemize}
\small
  \item Unified government control peaked at 55\% around 2010; while higher than the 1990s, it remains well below the pre-1990s high of 70\%
  \item Electoral Competitiveness for both executive and legislative indices shows a clear upward trend, indicating a global increase in multi-party competition and election contestability
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/backslide_exec.pdf")
```
\vspace{0.2cm}
\begin{itemize}
\small
  \item Post-Cold War, term limits, dismissal rules, and succession procedures surged; while term limits have recently plateaued, dismissal and succession rules continue a slight upward trend
  \item Between 2000 and 2018, 26\% of term-limited leaders (all in authoritarian or hybrid regimes) attempted to evade limits, with a 57\% success rate and no clear chronological trend
\end{itemize}

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/backslide_media.pdf")
```
\vspace{0.2cm}
\begin{itemize}
\small
  \item Jails for "anti-state" or professional reasons have risen steadily since 2000, marking a decade-long deterioration
  \item Conversely, fatalities peaked in the early 2000s and have decreased consistently since 2008
\end{itemize}

---

\begin{itemize}
\small
  \item Varieties of constitutional democracy
  \vspace{0.1cm}
  \begin{itemize}
    \item \textbf{Executive system}: presidential, parliamentary and semi-presidential
    \item \textbf{Legislative system}: uni- or bi-cameral
    \item Electoral system: majoritarian/plurality, PR and mixed-member (and the resulting party system)
    \item Central-local relations: federal, unitary and decentralized
  \end{itemize}
  \vspace{0.3cm}
  \item Democratic constitutional design is a response to "\textbf{Madison's Dilemma}"
  \vspace{0.1cm}
  \begin{itemize}
    \item the goal is to create a government strong enough to govern while limited enough to protect civil liberties
    \item Two visions of democracy: Majoritarian Model (power concentration) versus Consensus Model (power dispersion)
  \end{itemize}
\end{itemize}

# Refresher lecture

\begin{itemize}
  \item Executive structures
  \vspace{0.1cm}
  \begin{itemize}
    \item Presidential versus parliamentary democracies (and semi-presidentalism)
    \item "Perils of presidentialism"
  \end{itemize}
  \vspace{0.2cm}
  \item Unicameral and bicameral legislatures
  \vspace{0.1cm}
  \begin{itemize}
      \item Why bicamerialism?
      \item Different types of bicameralism: (a)symmetry and (incongrence) 
  \end{itemize}
  \vspace{0.2cm}
  \item Recap: Supreme/constitutional courts and judicial review (skim by yourself)
\end{itemize}

# Presidential versus parliamentary systems

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/type.pdf")
```

---

\begin{itemize}
  \item Is the government responsible for the legislative assembly?
  \vspace{0.1cm}
  \begin{itemize}
    \item Separation of origin
    \item Separation of survival
  \end{itemize}
  \vspace{0.4cm}
  \item Is there a "head of state" elected for a fixed term in office?
\end{itemize}

# Presidential versus parliamentary democracies from a delegation perspective

```{r echo=FALSE, out.width="75%", fig.align="center"}
knitr::include_graphics("Figs/type_delegate.pdf")
```
\vspace{0.2cm}
\pause
\begin{center}
\small
Having a president does not mean the country is a presidential democracy because "president" can mean different things in different countries!
\end{center}

# Presidential versus parliamentary democracies from a delegation perspective

```{r echo=FALSE, out.width="85%", fig.align="center"}
knitr::include_graphics("Figs/type_recap.pdf")
```
\vspace{0.2cm}
\begin{center}
\small
Having a president does not mean the country is a presidential democracy because "president" can mean different things in different countries!
\end{center}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pps_india.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/pps_germany.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/germany_fifa.pdf")
```
\vspace{0.1cm}
\begin{center}
Merkel and Gauck in Rio de Janeiro, FIFA World Cup 2014
\end{center}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/trend.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/trend_global.pdf")
```

# Presidential democracies are much less likely to survive

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/perils.pdf")
```

---

```{r echo=FALSE, out.width="80%", fig.align="center"}
knitr::include_graphics("Figs/coup.pdf")
```
\vspace{0.1cm}
\begin{center}
Rio de Janeiro, April 1, 1964
\end{center}

# Perils of presidentialism

\begin{itemize}
  \item The nature of executive-legislature relationship is different between parliamentary and presidential democracies
  \vspace{0.1cm}
  \begin{itemize}
    \item Presidential: Separation of origin and survival
    \item Parliamentary: Legislative responsibility
  \end{itemize}
  \vspace{0.3cm}
  \item What if the executive and legislature do not agree with each other?
  \vspace{0.1cm}
  \begin{itemize}
    \item Presidential: Locked in with temporal rigidity (and hence \textbf{policy} stability)
    \item Parliamentary: Vote of confidence by the govt \textbf{OR} vote of no confidence by the parliament
  \end{itemize}
\end{itemize}

---

\begin{itemize}
  \item Presidential: Extreme policy stability leads to gridlock and regime (or democratic) instability
  \vspace{0.5cm}
  \item Parliamentary: Extreme policy stability leads to government (or cabinet) instability
\end{itemize}

# Varieties of presidentialism

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/president.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/power1.pdf")
```

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/power2.pdf")
```

# Unicameral and bicameral legislatures in the world

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/legislature/trend.pdf")
```
\vspace{0.1cm}
\begin{center}
\small
Close to half of countries around the world have bicameral legislatures (in green), while more countries have unicameral legislatures (in blue). Data: Rice University and OpenStax.
\end{center}

# Why bicameralism

\begin{itemize}
  \item History and constitutional design
  \vspace{0.3cm}
  \item Different forms of political "representation"
  \begin{itemize}
    \item By constitueny (e.g., House of Representatives in Japan)
    \item By region (e.g., Senate in the US)
    \item By socioeconomic class (e.g., House of Lords in the UK)
  \end{itemize}
  \vspace{0.3cm}
  \item Malapportionment: When the distribution of political representation between constituencies does not match the constituency's relative population
\end{itemize}

---

```{r echo=FALSE, out.width="104%", fig.align="center"}
knitr::include_graphics("Figs/legislature/examples.pdf")
```
\vspace{0.3cm}
\begin{itemize}
  \small
  \item Japan: House of Representatives (lower) and Hose of Councilors (upper)
  \item Germany: Bundestag (lower) and Bundesrat (upper)
  \item India: Lok Sabha (lower) and Rajya Sabha (upper)
  \item Russia: State Duna (lower) and Federal Council (upper)
\end{itemize}

# (A)symmetry and (in)congrence bicameralism 

```{r echo=FALSE, out.width="50%", fig.align="center"}
knitr::include_graphics("Figs/legislature/bicameralism.pdf")
```
\vspace{0.1cm}
\begin{itemize}
  \item (A)Symmetry: Relative powers between two chambers/houses
  \item (In)Congruence: Relative preferences (and interests) between two chambers/houses
\end{itemize}

---

```{r echo=FALSE, out.width="90%", fig.align="center"}
knitr::include_graphics("Figs/legislature/ipu.pdf")
```
\vspace{0.1cm}
\begin{center}
\url{http://ipu.org/}
\end{center}

# Recap: Supreme/constitutional courts and judicial review

\begin{itemize}
  \item Judiciary in the constitution organizes relations between the executive and legislative branches of government:
  \vspace{0.2cm}
  \begin{itemize}
    \item \textbf{Parliamentary supremacy}: a principle according to which judges' decisions remain subordinate to decisions of the legislative majority (e.g., UK before 2009)
    \item \textbf{Judicial review}: the ability of a country's high court to invalidate laws the legislature has enacted by declaring them unconstitutional (e.g., USA, Germany and UK after 2009)
  \end{itemize}
  \vspace{0.3cm}
  \item What explains the adoption of either principle? What is the "politics" behind the operation of these principles?
\end{itemize}

---

```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/usa.pdf")
```

---

```{r echo=FALSE, out.width="60%", fig.align="center"}
knitr::include_graphics("Figs/germany.pdf")
```

---

\vspace{1.5cm}
"Domestic intelligence cannot push for a ban on the party -- that can only go through the \textbf{two houses of parliament}, \textbf{government} or \textbf{the constitutional court} -- but its latest decision could encourage others to start the process."

---

```{r echo=FALSE, out.width="100%", fig.align="center"}
knitr::include_graphics("Figs/uk.pdf")
```
